### 2019-09-27 CPP_analysis.Rmd
First commit

---
title: "CPP_analysis"
output:
  html_document: default
  pdf_document: default
---

```{r message=FALSE, include=FALSE}
install.packages("knitr")
install.packages("tidyverse") #1.2.1
install.packages("drc") #3.0.1
install.packages("stringr")
```

```{r message=FALSE}
library("drc") #3.0.1
library("knitr")
library("tidyverse") #1.2.1
library("stringr")
```

```{r datasheet}
all_data<-data.frame(read.csv("CPP_experiment_data_collated.csv"))
#Filter out empty/discarded oviposition papers 
blockAB<-dplyr::filter(all_data, TotalLarvaeEmbryos>0)
```

```{r yields}
#Calculate yield/fraction of initial count
blockAB$yield <- blockAB$TotalLarvaeEmbryos/blockAB$InitialEggs 
hist(blockAB$yield, main="Yield (final/initial count) per oviposition paper")
plot(blockAB$yield ~ blockAB$TotalLarvaeEmbryos)
```

```{r filter papers}
#removing papers with <60% or >100% initial count
blockAB <- filter(blockAB, yield<=1) %>% filter(yield>=0.6) 
```

```{r summarize}
DI_summary <- blockAB %>% 
  group_by(CageID) %>% 
  summarise(pop=first(Population), pp=mean(Photoperiod), lat=mean(Lat), block=first(Block), country=first(Country), h1=sum(H1Count), h2=sum(H2Count), emb=sum(EmbryoCount)) %>% 
  filter(h1+h2+emb>100)
DI_summary$total<-DI_summary$h1+DI_summary$h2+DI_summary$emb
DI_summary$DI<-DI_summary$emb/DI_summary$total
```

```{r totals_plot}
hist(DI_summary$total, main="Total number of viable eggs per cage")

```

```{r DI_plot}
#plot with DI at 8 h shown as TWO DIFFERENT points (corresponding to two cabinets)
ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~pop)

```

```{r incidence_lat}
DI_short<-filter(DI_summary, pp==8)
ggplot(DI_short, aes(x=lat, y=DI, color=country)) + geom_point() + theme_bw() +stat_smooth()
```

# Calculating CPP values

5-parameter generalized log-logistic model. Explanation on page 2: http://ftp.uni-bayreuth.de/math/statlib/R/CRAN/doc/vignettes/drc/drc.pdf
d and c are upper and lower limit, respectively
e is ED50
b is relative slope aroung e
f represents asymmetry in the model; f=1 if function is symmetrical

```{r CPP_calc_2019}
#From Kevin Emerson

# the following fits the models of the given function using data from the
# data.frame d separately for the various treatments using a 5 parameter
# logistic model
d.model.fit <- drm(DI ~ pp, data = DI_summary, curveid = pop, fct = LL.5())
#plot(d.model.fit, log = "")

# output the parameters of the model fits:
summary(d.model.fit)

# estimate the 50% intercept
ED(d.model.fit, 50)
cpp_values<-data.frame(ED(d.model.fit, 50))

```

#Recalculating CPP values for 2008 photoperiod response curves

```{r CPP_calc_2008}
DI_summary_2008<-data.frame(read.csv("CPP_2008.csv"))
d.model.fit.2008 <- drm(DI ~ PP, data = DI_summary_2008, curveid = POP, fct = LL.5())

#plot(d.model.fit.2008, log = "")
# output the parameters of the model fits:
summary(d.model.fit)
# estimate the 50% intercept
ED(d.model.fit.2008, 50)
cpp_values_2008<-data.frame(ED(d.model.fit.2008, 50))
```
#Combine and format CPP value tables from 1988, 2008 and 2018

```{r collate_cpp_values}
cpp_2019<-data.frame(ED(d.model.fit, 50))
cpp_2019<-rownames_to_column(cpp_2019, var="pop") 
cpp_2019$pop<-str_sub(cpp_2019$pop,3,5)
popdata_2019<-DI_summary %>% group_by(pop) %>% 
  summarize(lat=first(lat), country=first(country))
cpp_2019<-right_join(cpp_2019, popdata_2019, by="pop")
#drop PBS from this list for now; CPP estimate doesn't make sense
#cpp_2019<-filter(cpp_2019, pop!="PBS")
cpp_2019$year<-rep("2019", nrow(cpp_2019))

cpp_2008<-data.frame(ED(d.model.fit.2008, 50)) %>% rownames_to_column(var="pop")
cpp_2008$pop<-str_sub(cpp_2008$pop,3,5)
popdata_2008<-DI_summary_2008 %>% group_by(POP) %>% 
  summarize(lat=first(LAT), country=first(COUNTRY), year=first(YEAR)) 
colnames(popdata_2008)[1]<-"pop"
cpp_2008<-right_join(cpp_2008, popdata_2008, by="pop")

cpp_1988<-data.frame(read.csv("CPP_1988.csv")) 
cpp_1988$pop<-na_if(cpp_1988$pop, "NA")
cpp_1988$Std..Error<-na_if(cpp_1988$Std..Error, "NA")

cpp_2<-rbind(cpp_1988, cpp_2008, cpp_2019)

```

# CPP by latitude, country and year

```{r Japan_US_years}
ggplot(cpp_2, aes(x=lat, y=Estimate, fill=year, color=year)) + 
  scale_color_manual(values=c("white","gray","black")) + 
  scale_fill_manual(values=c("white","gray","black"))+ 
  geom_point(shape=21, color="black") + 
  geom_errorbar(aes(ymin=Estimate-Std..Error, ymax=Estimate+Std..Error), na.rm=TRUE) + 
  facet_grid(~country) + stat_smooth(method=lm, se=FALSE)
```
```{r }
ggplot(cpp_2, aes(x=lat, y=Estimate, color=country)) + 
  geom_point() + 
  geom_errorbar(aes(ymin=Estimate-Std..Error, ymax=Estimate+Std..Error), na.rm=TRUE) + 
  facet_grid(~year) + stat_smooth(method=lm, se=FALSE)
```

```{r incidence_lat_year}
DI_short_2008<-filter(DI_summary_2008, PP==8)
names(DI_short_2008)<-c("pop","country","lat","pp","year","DI")

DI_short_2018<-filter(DI_summary, pp==8) %>% group_by(pop) %>% 
  summarize(country=first(country), lat=first(lat), pp=first(pp), emb=sum(emb), total=sum(total))
DI_short_2018$DI<-DI_short_2018$emb/DI_short_2018$total
DI_short_2018<-DI_short_2018[,c(1:4,7)] 
DI_short_2018$year<-rep("2018",nrow(DI_short_2018))
DI_short_2018<-DI_short_2018[,c(1,2,3,4,6,5)]

DI_short_2<-rbind(DI_short_2008, DI_short_2018)

ggplot(DI_short_2, aes(x=lat, y=DI, color=year, label=pop)) + scale_color_manual(values=c("gray","black")) + 
  geom_point() + facet_grid(~country)

```

### 2019-10-02 CPP_analysis.Rmd
Renamed columns in different files for consistency; made maps

@@
 DI_summary_2008<-data.frame(read.csv("CPP_2008.csv"))
-d.model.fit.2008 <- drm(DI ~ PP, data = DI_summary_2008, curveid = POP, fct = LL.5())
+d.model.fit.2008 <- drm(DI ~ photoperiod, data = DI_summary_2008, curveid = pop, fct = LL.5())
 
@@
 cpp_2019<-right_join(cpp_2019, popdata_2019, by="pop")
-#drop PBS from this list for now; CPP estimate doesn't make sense
-#cpp_2019<-filter(cpp_2019, pop!="PBS")
 cpp_2019$year<-rep("2019", nrow(cpp_2019))
+colnames(cpp_2019)[colnames(cpp_2019)=="Std..Error"] <- "StdError"
 
@@
 cpp_2008$pop<-str_sub(cpp_2008$pop,3,5)
-popdata_2008<-DI_summary_2008 %>% group_by(POP) %>% 
-  summarize(lat=first(LAT), country=first(COUNTRY), year=first(YEAR)) 
-colnames(popdata_2008)[1]<-"pop"
+colnames(cpp_2008)[colnames(cpp_2008)=="Std..Error"] <- "StdError"
+popdata_2008<-DI_summary_2008 %>% group_by(pop) %>% 
+  summarize(country=first(country), lat=first(lat), year=first(year)) 
+#colnames(popdata_2008)[1]<-"pop"
 cpp_2008<-right_join(cpp_2008, popdata_2008, by="pop")
@@
 cpp_1988$pop<-na_if(cpp_1988$pop, "NA")
-cpp_1988$Std..Error<-na_if(cpp_1988$Std..Error, "NA")
+cpp_1988$StdError<-na_if(cpp_1988$StdError, "NA")
 
@@
   geom_point(shape=21, color="black") + 
-  geom_errorbar(aes(ymin=Estimate-Std..Error, ymax=Estimate+Std..Error), na.rm=TRUE) + 
+  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
   facet_grid(~country) + stat_smooth(method=lm, se=FALSE)
@@
   geom_point() + 
-  geom_errorbar(aes(ymin=Estimate-Std..Error, ymax=Estimate+Std..Error), na.rm=TRUE) + 
+  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
   facet_grid(~year) + stat_smooth(method=lm, se=FALSE)
@@
 ```{r incidence_lat_year}
-DI_short_2008<-filter(DI_summary_2008, PP==8)
-names(DI_short_2008)<-c("pop","country","lat","pp","year","DI")
+DI_short_2008<-filter(DI_summary_2008, photoperiod==8)
 
+#unnecessary renaming between pp/photoperiod - harmonize @ beginning?
 DI_short_2018<-filter(DI_summary, pp==8) %>% group_by(pop) %>% 
-  summarize(country=first(country), lat=first(lat), pp=first(pp), emb=sum(emb), total=sum(total))
+  summarize(country=first(country), lat=first(lat), photoperiod=first(pp), emb=sum(emb), total=sum(total))
 DI_short_2018$DI<-DI_short_2018$emb/DI_short_2018$total
@@
 DI_short_2018$year<-rep("2018",nrow(DI_short_2018))
-DI_short_2018<-DI_short_2018[,c(1,2,3,4,6,5)]
 
 DI_short_2<-rbind(DI_short_2008, DI_short_2018)
+write_csv(DI_short_2, "diapause_incidence_8h_2008_2018.csv")
 

### 2019-10-02 maps.R

@@ -1,0 +1,37 @@
+install.packages(c("maps", "mapdata"))
+devtools::install_github("dkahle/ggmap")
+install.packages("tidyverse")
+install.packages("scatterpie")
+install.packages("ggedit")
+
+library(ggplot2)
+library(ggmap)
+library(maps)
+library(mapdata)
+library(scatterpie)
+library(ggedit)
+
+usa<-map_data("usa")
+jp<-map_data("japan")
+pops<-data.frame(read.csv("mapping_pops_cpp_expt.csv"))
+
+us_map<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="black") + 
+  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 1) + xlim(c(-90, -65)) +
+  geom_label(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
+  theme_bw() 
+
+japan_map<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="black") + 
+  coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 1)  +
+  geom_label(data=filter(pops, country=="JP"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
+  theme_bw() + theme(axis.title.x=element_blank(), axis.title.y=element_blank())
+
+DI_data<-data.frame(read.csv("diapause_incidence_8h_2008_2018.csv"))
+DI_data$ND<-1-DI_data$DI
+DI_info<-full_join(DI_data, pops)
+DI_info<-rename(DI_info, Diapause=DI, Nondiapause=ND)
+
+remove_geom(us_map, "label") + geom_scatterpie(data=filter(DI_info, country=="US"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + 
+  facet_grid(~year) + theme(axis.title.x=element_blank(), axis.title.y=element_blank())
+
+remove_geom(japan_map, "label") + geom_scatterpie(data=filter(DI_info, country=="JP"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + 
[…]

### 2019-10-02 CPP_analysis.Rmd
Calculate dates

@@
 install.packages("stringr")
+install.packages("geosphere")
+install.packages("lubridate")
 ```
@@
 library("drc") #3.0.1
-library("knitr")
+library("knitr") #1.25
 library("tidyverse") #1.2.1
-library("stringr")
+library("stringr") #1.4.0
+library("geosphere") #1.5.10
+library("lubridate") #1.7.4
 ```
@@
 ```
+
+```{r day_function}
+findDate<-function(lat, pp, year) {
+#generate day lengths at latitude between July 1 and Dec 31
+x<-daylength(lat,182:365)
+#find closest day length in vector
+daynumber<-which.min(abs(x-pp))
+#convert day number to date
+origin<-as.character(paste(year, "-07-01",sep=""))
+return(as.character(as.Date(daynumber, origin=origin)))
+}
+
+```
+
+```{r}
+cpp_2$date<-mapply(findDate, lat=cpp_2$lat, pp=cpp_2$Estimate, year=cpp_2$year)
+cpp_2$date_again<-format(as.Date(cpp_2$date), "%mm-%dd")
+cpp_2$date_again<-as.Date(cpp_2$date_again, format="%mm-%dd")
+ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=date_again, color=year)) + geom_point() + scale_y_date(date_breaks = "weeks" , date_minor_breaks="days", date_labels = "%b-%d") + facet_grid(~country) + theme_bw()
+```

### 2019-10-03 CPP_analysis.Rmd
small adjustments to figures

--- 2019-10-02 CPP_analysis.Rmd
@@
 output:
+  word_document: default
+  pdf_document: default
   html_document: default
-  pdf_document: default
 ---
@@
 
-```{r CPP_calc_2019}
+```{r CPP_calc_2018}
 #From Kevin Emerson
@@
 ```{r collate_cpp_values}
-cpp_2019<-data.frame(ED(d.model.fit, 50))
-cpp_2019<-rownames_to_column(cpp_2019, var="pop") 
-cpp_2019$pop<-str_sub(cpp_2019$pop,3,5)
-popdata_2019<-DI_summary %>% group_by(pop) %>% 
+cpp_2018<-data.frame(ED(d.model.fit, 50))
+cpp_2018<-rownames_to_column(cpp_2018, var="pop") 
+cpp_2018$pop<-str_sub(cpp_2018$pop,3,5)
+popdata_2018<-DI_summary %>% group_by(pop) %>% 
   summarize(lat=first(lat), country=first(country))
-cpp_2019<-right_join(cpp_2019, popdata_2019, by="pop")
-cpp_2019$year<-rep("2019", nrow(cpp_2019))
-colnames(cpp_2019)[colnames(cpp_2019)=="Std..Error"] <- "StdError"
+cpp_2018<-right_join(cpp_2018, popdata_2018, by="pop")
+cpp_2018$year<-rep("2018", nrow(cpp_2018))
+colnames(cpp_2018)[colnames(cpp_2018)=="Std..Error"] <- "StdError"
 
@@
   summarize(country=first(country), lat=first(lat), year=first(year)) 
-#colnames(popdata_2008)[1]<-"pop"
 cpp_2008<-right_join(cpp_2008, popdata_2008, by="pop")
@@
 
-cpp_2<-rbind(cpp_1988, cpp_2008, cpp_2019)
+cpp_2<-rbind(cpp_1988, cpp_2008, cpp_2018)
 
@@
 ```
+
+```{r day_function}
+findDate<-function(lat, pp, year) {
+#generate day lengths at latitude between July 1 and Dec 31
+x<-daylength(lat,182:365)
+#find position of day length in the daylength vector that is closest to photoperiod queried
+daynumber<-which.min(abs(x-pp))
+#convert day number to date
+origin<-as.character(paste(year,"-07-01",sep=""))
+return(as.character(as.Date(daynumber, origin=origin)))
+}
+
+```
+
+```{r}
+cpp_2$date<-mapply(findDate, lat=cpp_2$lat, pp=cpp_2$Estimate, year=cpp_2$year)
+cpp_2$estimated_date<-format(as.Date(cpp_2$date), "%mm-%dd")
+cpp_2$estimated_date<-as.Date(cpp_2$estimated_date, format="%mm-%dd")
+ggplot(cpp_2, aes(x=lat, y=estimated_date, color=year)) + geom_point() + scale_y_date(date_breaks = "weeks" , date_minor_breaks="days", date_labels = "%b-%d") + facet_grid(~country) + theme_bw() + stat_smooth(method=lm, se=FALSE)
+```

### 2019-10-09 CPP_analysis.Rmd
Figures

@@
 output:
+  html_document: default
+  pdf_document: default
   word_document: default
-  pdf_document: default
-  html_document: default
 ---
@@
 #removing papers with <60% or >100% initial count
-blockAB <- filter(blockAB, yield<=1) %>% filter(yield>=0.6) 
+blockAB_unfiltered<-blockAB
+blockAB_filtered <- filter(blockAB, yield<=1) %>% filter(yield>=0.6) 
 ```
@@
 ```{r summarize}
+
+DI_summarize <- function(blockAB) {
 DI_summary <- blockAB %>% 
@@
 DI_summary$DI<-DI_summary$emb/DI_summary$total
+return(DI_summary)
+}
+
+```
+
+```{r}
+DI_summary<-DI_summarize(blockAB_filtered)
 ```
@@
 b is relative slope aroung e
-f represents asymmetry in the model; f=1 if function is symmetrical
+f represents asymmetry in the model; f=1 if  is symmetrical
 
@@
 
-# the following fits the models of the given function using data from the
+# the following fits the models of the given  using data from the
 # data.frame d separately for the various treatments using a 5 parameter
@@
 
-#Recalculating CPP values for 2008 photoperiod response curves
+# Recalculating CPP values for 2008 photoperiod response curves
 
@@
 ```
-#Combine and format CPP value tables from 1988, 2008 and 2018
+# Combine and format CPP value tables from 1988, 2008 and 2018
 
@@
 cpp_2<-rbind(cpp_1988, cpp_2008, cpp_2018) 
+write.csv(cpp_2, "cpp_calcs_filtered.csv")
 
@@
 ```{r }
-ggplot(cpp_2, aes(x=lat, y=Estimate, color=country)) + 
+ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, color=country)) + 
   geom_point() + 
@@
 
-```{r}
+```{r approx_dates, include=FALSE}
 cpp_2$date<-mapply(findDate, lat=cpp_2$lat, pp=cpp_2$Estimate, year=cpp_2$year)
@@
 cpp_2$estimated_date<-as.Date(cpp_2$estimated_date, format="%mm-%dd")
-ggplot(cpp_2, aes(x=lat, y=estimated_date, color=year)) + geom_point() + scale_y_date(date_breaks = "weeks" , date_minor_breaks="days", date_labels = "%b-%d") + facet_grid(~country) + theme_bw() + stat_smooth(method=lm, se=FALSE)
+date_plot<-ggplot(cpp_2, aes(x=lat, y=estimated_date, color=year)) + geom_point() + scale_y_date(date_breaks = "weeks" , date_minor_breaks="days", date_labels = "%b-%d") +  facet_grid(~country) + theme_bw() + stat_smooth(method=lm, se=FALSE)
 ```
+
+The approximate date of 50% diapause in Berlin, NJ in 2018 was `r cpp_2[cpp_2$pop=="BER" & cpp_2$year==2018,8] %>% format("%B %d")`. 
+```{r echo=FALSE}
+date_plot

### 2019-10-23 CPP_analysis.Rmd
more

@@
 ```
+```{r}
+cppmodel<-lm(Estimate~lat*country*year, data=filter(cpp_2, year!=1988))
+anova(cppmodel)
+```
+
 ```{r }
-ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, color=country)) + 
+ggplot(cpp_2, aes(x=lat, y=Estimate, color=country)) + 
   geom_point() + 
@@
   facet_grid(~year) + stat_smooth(method=lm, se=FALSE)
+```
+```{r}
+ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, color=country, shape=year)) + 
+  geom_point() +
+  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
+  stat_smooth(method=lm, se=FALSE, aes(linetype=year)) 
 ```

### 2019-10-24 CPP_analysis.Rmd
home

--- 2019-10-09 CPP_analysis.Rmd
@@
   group_by(CageID) %>% 
-  summarise(pop=first(Population), pp=mean(Photoperiod), lat=mean(Lat), block=first(Block), country=first(Country), h1=sum(H1Count), h2=sum(H2Count), emb=sum(EmbryoCount)) %>% 
+  summarise(pop=first(Population), pp=mean(Photoperiod), block=first(Block), lat=mean(Lat), block=first(Block), country=first(Country), h1=sum(H1Count), h2=sum(H2Count), emb=sum(EmbryoCount)) %>% 
   filter(h1+h2+emb>100)
@@
 popdata_2018<-DI_summary %>% group_by(pop) %>% 
-  summarize(lat=first(lat), country=first(country))
+  summarize(lat=first(lat), country=first(country), block=first(block))
 cpp_2018<-right_join(cpp_2018, popdata_2018, by="pop")
@@
 
+```
+
+```{r}
+ggplot(cpp_2018, aes(x=lat, y=Estimate, color=block)) + geom_point() + stat_smooth()
 ```

### 2019-10-25 MS.Rmd
Manuscript document draft created

---
title: "Ae albopictus CPP manuscript"
output:
  word_document: 
    fig_width: 3.5
    fig_height: 3
  html_document: default
---

```{r eval=FALSE, message=FALSE, include=FALSE}
install.packages("tidyverse") #1.2.1
install.packages("drc") #3.0.1
install.packages("stringr")
install.packages("geosphere")
install.packages("lubridate")
install.packages(c("maps", "mapdata"))
devtools::install_github("dkahle/ggmap")
install.packages("scatterpie")
install.packages("ggedit")
install.packages("cowplot")
```

```{r message=FALSE, echo=FALSE, include=FALSE}
library("drc") #3.0.1
library("knitr") #1.25
library("tidyverse") #1.2.1
library("stringr") #1.4.0
library("geosphere") #1.5.10
library("lubridate") #1.7.4
library("maps")
library("mapdata")
library("scatterpie")
library("ggedit")
library("cowplot")
```

```{r datasheet, echo=FALSE, include=FALSE}
all_data<-data.frame(read.csv("CPP_experiment_data_collated.csv"))
#Filter out empty/discarded oviposition papers 
blockAB<-dplyr::filter(all_data, TotalLarvaeEmbryos>0)
```

```{r yields, echo=FALSE, include=FALSE}
#Calculate yield/fraction of initial count
blockAB$yield <- blockAB$TotalLarvaeEmbryos/blockAB$InitialEggs 
hist(blockAB$yield, main="Yield (final/initial count) per oviposition paper")
plot(blockAB$yield ~ blockAB$TotalLarvaeEmbryos)
```

```{r filter papers, echo=FALSE, include=FALSE}
#removing papers with <60% or >100% initial count
blockAB_unfiltered<-blockAB
blockAB_filtered <- filter(blockAB, yield<=1) %>% filter(yield>=0.6) 
```

```{r summarize, echo=FALSE, include=FALSE}
DI_summarize <- function(blockAB) {
DI_summary <- blockAB %>% 
  group_by(CageID) %>% 
  summarise(pop=first(Population), pp=mean(Photoperiod), lat=mean(Lat), block=first(Block), country=first(Country), h1=sum(H1Count), h2=sum(H2Count), emb=sum(EmbryoCount)) %>% 
  filter(h1+h2+emb>100)
DI_summary$total<-DI_summary$h1+DI_summary$h2+DI_summary$emb
DI_summary$DI<-DI_summary$emb/DI_summary$total
return(DI_summary)
}
```

```{r, echo=FALSE, include=FALSE}
DI_summary<-DI_summarize(blockAB_filtered)
```

```{r totals_plot, echo=FALSE, include=FALSE}
hist(DI_summary$total, main="Total number of viable eggs per cage")

```

```{r DI_plot, echo=FALSE, include=FALSE}
#plot with DI at 8 h shown as TWO DIFFERENT points (corresponding to two cabinets)
ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~pop)

```

```{r incidence_lat, echo=FALSE, include=FALSE}
DI_short<-filter(DI_summary, pp==8)
ggplot(DI_short, aes(x=lat, y=DI, color=country)) + geom_point() + theme_bw() +stat_smooth()
```

```{r CPP_calc_2018, echo=FALSE, include=FALSE}
#From Kevin Emerson
# the following fits the models of the given  using data from the
# data.frame d separately for the various treatments using a 5 parameter
# logistic model
d.model.fit <- drm(DI ~ pp, data = DI_summary, curveid = pop, fct = LL.5())
#plot(d.model.fit, log = "")
# output the parameters of the model fits:
summary(d.model.fit)
# estimate the 50% intercept
ED(d.model.fit, 50)
cpp_values<-data.frame(ED(d.model.fit, 50))
```

```{r CPP_calc_2008, echo=FALSE, include=FALSE}
DI_summary_2008<-data.frame(read.csv("CPP_2008.csv"))
d.model.fit.2008 <- drm(DI ~ photoperiod, data = DI_summary_2008, curveid = pop, fct = LL.5())

#plot(d.model.fit.2008, log = "")
# output the parameters of the model fits:
summary(d.model.fit)
# estimate the 50% intercept
ED(d.model.fit.2008, 50)
cpp_values_2008<-data.frame(ED(d.model.fit.2008, 50))
```

```{r collate_cpp_values, echo=FALSE, include=FALSE}
cpp_2018<-data.frame(ED(d.model.fit, 50))
cpp_2018<-rownames_to_column(cpp_2018, var="pop") 
cpp_2018$pop<-str_sub(cpp_2018$pop,3,5)
popdata_2018<-DI_summary %>% group_by(pop) %>% 
  summarize(lat=first(lat), country=first(country))
cpp_2018<-right_join(cpp_2018, popdata_2018, by="pop")
cpp_2018$year<-rep("2018", nrow(cpp_2018))
colnames(cpp_2018)[colnames(cpp_2018)=="Std..Error"] <- "StdError"
cpp_2008<-data.frame(ED(d.model.fit.2008, 50)) %>% rownames_to_column(var="pop")
cpp_2008$pop<-str_sub(cpp_2008$pop,3,5)
colnames(cpp_2008)[colnames(cpp_2008)=="Std..Error"] <- "StdError"
popdata_2008<-DI_summary_2008 %>% group_by(pop) %>% 
  summarize(country=first(country), lat=first(lat), year=first(year)) 
cpp_2008<-right_join(cpp_2008, popdata_2008, by="pop")
cpp_1988<-data.frame(read.csv("CPP_1988.csv")) 
cpp_1988$pop<-na_if(cpp_1988$pop, "NA")
cpp_1988$StdError<-na_if(cpp_1988$StdError, "NA")
cpp_2<-rbind(cpp_1988, cpp_2008, cpp_2018) 
write.csv(cpp_2, "cpp_calcs_filtered.csv")
```

```{r Japan_US_years, echo=FALSE, include=FALSE}
ggplot(cpp_2, aes(x=lat, y=Estimate, fill=year, color=year)) + 
  scale_color_manual(values=c("white","gray","black")) + 
  scale_fill_manual(values=c("white","gray","black"))+ 
  geom_point(shape=21, color="black") + 
  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
  facet_grid(~country) + stat_smooth(method=lm, se=FALSE)
```

```{r cppmodel, echo=FALSE, include=FALSE}
cppmodel<-lm(Estimate~lat*country*year, data=filter(cpp_2, year!=1988))
anova(cppmodel)
```

```{r, echo=FALSE, include=FALSE}
ggplot(cpp_2, aes(x=lat, y=Estimate, color=country)) + 
  geom_point() + 
  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
  facet_grid(~year) + stat_smooth(method=lm, se=FALSE)
```

```{r, echo=FALSE, include=FALSE}
ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, color=country, shape=year)) + 
  geom_point() +
  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
  stat_smooth(method=lm, se=FALSE, aes(linetype=year)) 
```

```{r incidence_lat_year, echo=FALSE, include=FALSE}
DI_short_2008<-filter(DI_summary_2008, photoperiod==8)
DI_short_2018<-filter(DI_summary, pp==8) %>% group_by(pop) %>% 
  summarize(country=first(country), lat=first(lat), photoperiod=first(pp), emb=sum(emb), total=sum(total))
DI_short_2018$DI<-DI_short_2018$emb/DI_short_2018$total
DI_short_2018<-DI_short_2018[,c(1:4,7)] 
DI_short_2018$year<-rep("2018",nrow(DI_short_2018))
DI_short_2<-rbind(DI_short_2008, DI_short_2018)
write_csv(DI_short_2, "diapause_incidence_8h_2008_2018.csv")
incidence_lat_year_plot<-ggplot(DI_short_2, aes(x=lat, y=DI, color=year, label=pop)) + scale_color_manual(values=c("gray","black")) + 
  geom_point() + facet_grid(~country)

country_DI_2018<-ggplot(DI_short_2018, aes(x=lat, y=DI, color=country)) + scale_color_manual(values=c("red","blue")) + geom_point() + theme_bw() + stat_smooth() + theme(legend.position="none")
```

```{r dimodel, echo=FALSE, include=FALSE}
dimodel<-lm(DI~lat*country*year, data=DI_short_2)
anova(dimodel)
```

```{r day_function, echo=FALSE, include=FALSE}
findDate<-function(lat, pp, year) {
#generate day lengths at latitude between July 1 and Dec 31
x<-daylength(lat,182:365)
#find position of day length in the daylength vector that is closest to photoperiod queried
daynumber<-which.min(abs(x-pp))
#convert day number to date
origin<-as.character(paste(year,"-07-01",sep=""))
return(as.character(as.Date(daynumber, origin=origin)))
}
```

```{r approx_dates, include=FALSE, echo=FALSE, include=FALSE}
cpp_2$date<-mapply(findDate, lat=cpp_2$lat, pp=cpp_2$Estimate, year=cpp_2$year)
cpp_2$estimated_date<-format(as.Date(cpp_2$date), "%mm-%dd")
cpp_2$estimated_date<-as.Date(cpp_2$estimated_date, format="%mm-%dd")
date_plot<-ggplot(cpp_2, aes(x=lat, y=estimated_date, color=year)) + geom_point() + scale_y_date(date_breaks = "weeks" , date_minor_breaks="days", date_labels = "%b-%d") +  facet_grid(~country) + theme_bw() + stat_smooth(method=lm, se=FALSE)
```

```{r maps, echo=FALSE, include=FALSE}
usa<-map_data("usa")
jp<-map_data("japan")
pops<-data.frame(read.csv("mapping_pops_cpp_expt.csv"))

us_map<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="blue") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 1) + xlim(c(-90, -65)) +
  geom_label(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
  theme_bw()

japan_map<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="red") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 1)  +
  geom_label(data=filter(pops, country=="JP"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
  theme_bw() + theme(axis.title.x=element_blank(), axis.title.y=element_blank())

DI_data<-data.frame(read.csv("diapause_incidence_8h_2008_2018.csv"))
DI_data$ND<-1-DI_data$DI
DI_info<-full_join(DI_data, pops)
DI_info<-rename(DI_info, Diapause=DI, Nondiapause=ND)

us_DI_pies<-remove_geom(us_map, "label") + geom_scatterpie(data=filter(DI_info, country=="US"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + 
  facet_grid(~year) + theme(axis.title.x=element_blank(), axis.title.y=element_blank())

jp_DI_pies<-remove_geom(japan_map, "label") + geom_scatterpie(data=filter(DI_info, country=="JP"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + 
  facet_grid(~year) + theme(axis.title.x=element_blank(), axis.title.y=element_blank())

DI_pies_2018<-plot_grid(remove_geom(us_map, "label") + geom_scatterpie(data=filter(DI_info, country=="US", year=="2018"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none"), remove_geom(japan_map, "label") + geom_scatterpie(data=filter(DI_info, country=="JP", year=="2018"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none"), ncol=1, align="hv")
```

## Introduction

- Diapause is an important overwintering adaptation in many organisms
- This adaptation comes with trade-offs, but exactly what they are is unclear (winter phenotypes of diapausing organisms; life history strategies)
- Climate change may have unpredictable effects on growing seasons
- Invasive species provide natural experiments for rapid adaptation
- Aedes albopictus worldwide spread: diapause played a role 
- What is the continuing role of diapause in light of environmental change, continuing introductions/admixture,

## Methods

### Mosquito population sampling and rearing
Mosquito populations were sampled in the eastern USA and Japan in 2018, at latitudes ranging from (Fig 1). Sampling locations are described in Table 1. Locations in the US were largely identical to locations sampled in 2008. (More details about field sampling.) At least 50 male and 50 female mosquitoes were used to initiate laboratory populations, except for SEN, which had X females and Y males.

```{r table1, echo=FALSE}
kable(filter(pops, only_2008=="n")[,1:5], col.names=c("Population", "City", "Latitude", "Longitude", "Country"), caption="Table 1: 2018 Sampling Locations")
```

Populations were maintained in the laboratory for 3 or 4 generations prior to the experiment at 21C and 80% humidity under a 16:8h light:dark cycle. Larvae were reared in distilled water and fed a slurry of dog food and brine shrimp, while adults were fed organic raisins and blood-fed by human arm. Initial laboratory generations of larvae were reared in plastic storage containers, with one population per container and densities varying by population. Pupae were then transferred into one adult cage per population. Populations generally had at least 50 female and 50 male adults per generation, but the JAC population experienced visibly high mortality during the F2 generation. The larval generation immediately prior to the experiment was reared in 15 cm plastic petri dishes containing 30-40 larvae each, X dishes per population. Additional details of mosquito husbandry can be found in Supplementary File 1.

### Critical photoperiod determination
To determine the critical photoperiod of each population, the following general procedure was used: Starting from the pupal stage, mosquitoes were divided among 12 different photoperiods and allowed to produce eggs. The proportion of diapause eggs at each photoperiod was determined, and a dose-response curve was constructed with respect to hours of light and proportion of eggs in diapause. The critical photoperiod was calculated as the number of hours of light at which 50% of the eggs produced were in diapause, excluding the fraction of constitutively non-diapausing eggs (i.e. the fraction that did not enter diapause at 8 hours of light). The experimental design and most of the equipment and experimental details were the same as those used in Urbanski et al 2012; additional details on all steps can be found in Supplementary File 1.

Due to the large number of populations and photoperiods, the experiment was conducted in two blocks. Each block contained half of the populations from each country; populations across the full range of latitudes were interspersed between the blocks. At beginning of each block, eggs were hatched by submerging oviposition papers in water with larval food for two days. Larvae were distributed among X 15 cm petri dishes containing 90 ml of a mixture of distilled water and larval food (5 L of water to 10 mL of food slurry) at a density of 30-40 larvae per dish. Larvae were transferred to clean dishes containing fresh water/food mixture three times a week until pupation. Larvae were kept in the controlled temperature room at 21 degrees, 80 percent humidity and 16:8 hours light:dark until they produced pupae.

Pupae from each population were collected, pooled across dishes, and haphazardly distributed amongst thirteen 1.5 L adult cages, corresponding to 12 photoperiod cabinets and an additional short-day (8h light) incubator. The total number of pupae added to each cage ranged from X to Y; subsequent pupal and adult mortality was not monitored. Approximately one week after the last pupae were added, cages were blood-fed, and a dark oviposition cup half-filled with water and lined with unbleached paper towel was added to each cage. Oviposition papers were collected 5-7 days after blood-feeding and replaced every 2-3 days; a total of four oviposition papers were collected from each cage. Several cages were fed again and an additional oviposition paper was collected; see Supplementary File 1 for details. Oviposition papers were dried gently and stored in the photoperiod cabinets for 10 to 21 days until the diapause scoring procedure.

Oviposition papers were trimmed to a ~2 cm strip surrounding what was the waterline in the oviposition cup; this zone generally contained the vast majority of the eggs. On papers with few eggs, the paper was trimmed in such a way as to capture all eggs. On papers with very dense oviposition, the egg-containing strip was haphazardly subsampled further, to several hundred eggs as approximated by eye. Eggs were counted and egg papers were submerged in distilled water with several drops of larval food. Larvae that hatched were counted two days later and papers were re-dried. After 1-2 weeks, hatch was induced again and second installment hatched larvae were counted. Oviposition papers were then treated overnight with a bleach solution to clear the chorion of the remaining eggs. Eggs containing a pharate larva were scored as embryonated and in diapause. The total number of hatched larvae and embryonated eggs was considered the number of viable eggs. Oviposition papers in which viable eggs made up less than 60% or more than 100% of the eggs initially counted on the paper were excluded from further analysis. Analysis dates, subsampling and other detailed information about each oviposition paper can be found in Data Table 1.

For each cage, the counts of larvae and embryos across all oviposition papers were pooled to calculate the diapause fraction: number of embryonated eggs divided by total number of viable eggs. The median total number of viable eggs per cage was `r median(DI_summary$total)` (range `r min(DI_summary$total)`-`r max(DI_summary$total)`). For the 8-hour photoperiod, each population had data from two cages: one in a photoperiod cabinet and one in a separate incubator. These data were pooled to calculate the diapause incidence value for the population, but used as separate data points in photoperiod response curves.

### Statistical analysis
Critical photoperiod was calculated from photoperiod response curves as ED50 using a five-parameter generalized log-logistic model in the R (3.6.0) package drc (3.0.1). Effects of latitude, country and year (2008 vs. 2018) on CPP and on DI were analyzed with linear models. Additional packages used for calculations, analysis and figures included tidyverse (1.2.1), maps, mapdata, scatterpie…(update package list as figures come together).

## Results

Latitude-specific diapause incidence (the fraction of diapausing eggs produced under unambiguous short-day conditions) did not significantly differ between 2008 and 2018 in Japan or the US (Figure 1, Table 2). Diapause incidence was high (generally greater than 90%) at latitudes above ~32.5 degrees in both countries (Figure 2A). Substantial fractions of constitutively non-diapausing eggs were observed in populations south of South Carolina in the US and in Okinawa in Japan. Diapause incidence followed a latitudinal cline in the US, but a cline could not be observed in Japan due to absence of land to sample between Okinawa and Kagoshima. In the limited number of southern populations that could be compared between the US and Japan, latitude-specific diapause incidence appeared to be lower in the US (Figure 2B).

```{r fig1, fig.cap="Figure 1: Diapause incidence in Japan and US in 2008 and 2018", echo=FALSE}
incidence_lat_year_plot
```

```{r table2, echo=FALSE}
kable(anova(dimodel), digits=3, caption="Table 2")
```

```{r fig2, fig.cap="Figure 2: Diapause incidence (2018) by country. Black sections of pie charts represent diapausing fraction; white fractions represent non-diapausing. Blue points represent US diapause incidence; red points represent Japan.", echo=FALSE, message=FALSE}
plot_grid(DI_pies_2018, country_DI_2018, ncol=2, labels="AUTO")
```

As observed in

### 2019-10-25 MS.Rmd
MS edits

@@
 title: "Ae albopictus CPP manuscript"
+header-includes: #allows you to add in your own Latex packages
+- \usepackage{float} #use the 'float' package
+- \floatplacement{figure}{H} #make every figure with caption = h
+- \usepackage{caption}
+- \captionsetup[figure]{labelformat=empty}
+- \captionsetup[table]{labelformat=empty}
 output:
-  word_document: 
-    fig_width: 3.5
-    fig_height: 3
+  pdf_document: default
   html_document: default
+  word_document: default
 ---
@@
 #plot with DI at 8 h shown as TWO DIFFERENT points (corresponding to two cabinets)
-ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~pop)
+DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~pop)
 
@@
 ```{r Japan_US_years, echo=FALSE, include=FALSE}
-ggplot(cpp_2, aes(x=lat, y=Estimate, fill=year, color=year)) + 
+cpp_japan_us_years<-ggplot(cpp_2, aes(x=lat, y=Estimate, fill=year, color=year)) + 
   scale_color_manual(values=c("white","gray","black")) + 
@@
 ```{r, echo=FALSE, include=FALSE}
-ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, color=country, shape=year)) + 
-  geom_point() +
+cpp_08_18_shifts<-ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, color=country, shape=year)) + 
+  geom_point() + scale_color_manual(values=c("red", "blue")) +
   geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
-  stat_smooth(method=lm, se=FALSE, aes(linetype=year)) 
+  stat_smooth(method=lm, se=FALSE, aes(linetype=year)) + theme_bw() + theme(legend.position=c(.8,.25))
 ```
@@
 
-country_DI_2018<-ggplot(DI_short_2018, aes(x=lat, y=DI, color=country)) + scale_color_manual(values=c("red","blue")) + geom_point() + theme_bw() + stat_smooth() + theme(legend.position="none")
+country_DI_2018<-ggplot(DI_short_2018, aes(x=lat, y=DI, color=country)) + scale_color_manual(values=c("red","blue")) + geom_point() + theme_bw() + stat_smooth() + theme(legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "cm"))
 ```
@@
 
-DI_pies_2018<-plot_grid(remove_geom(us_map, "label") + geom_scatterpie(data=filter(DI_info, country=="US", year=="2018"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none"), remove_geom(japan_map, "label") + geom_scatterpie(data=filter(DI_info, country=="JP", year=="2018"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none"), ncol=1, align="hv")
+DI_pies_2018<-plot_grid(remove_geom(us_map, "label") + geom_scatterpie(data=filter(DI_info, country=="US", year=="2018"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "cm")), remove_geom(japan_map, "label") + geom_scatterpie(data=filter(DI_info, country=="JP", year=="2018"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "cm")), ncol=1, align="hv")
 ```
@@
 - Aedes albopictus worldwide spread: diapause played a role 
-- What is the continuing role of diapause in light of environmental change, continuing introductions/admixture,
+- What is the continuing role of diapause in light of environmental change, continuing introductions, competing selective pressures? 
+- We recapitulated 2008 CPP study
 
@@
 ### Mosquito population sampling and rearing
-Mosquito populations were sampled in the eastern USA and Japan in 2018, at latitudes ranging from (Fig 1).
+Mosquito populations were sampled in the eastern USA and Japan in 2018, at latitudes ranging from `r min(pops$latitude)` to `r max(pops$latitude)`.
 Sampling locations are described in Table 1.
 Locations in the US were largely identical to locations sampled in 2008.
-(More details about field sampling.) At least 50 male and 50 female mosquitoes were used to initiate laboratory populations, except for SEN, which had X females and Y males.
+[More details about field sampling.] At least 50 male and 50 female mosquitoes were used to initiate laboratory populations, except for SEN, which had X females and Y males.
 
@@
 
+Photoperiod response data from 2008 and calculated CPP values from 1988 were obtained from Urbanski et al 2012 data files (doi:10.5061/dryad.68277435).
+
 ### Statistical analysis
-Critical photoperiod was calculated from photoperiod response curves as ED50 using a five-parameter generalized log-logistic model in the R (3.6.0) package drc (3.0.1).
+Critical photoperiod was calculated from photoperiod response curves (Figure S1) as ED50 using a five-parameter generalized log-logistic model in the R (3.6.0) package drc (`r packageVersion("drc")`).
 Effects of latitude, country and year (2008 vs.
 2018) on CPP and on DI were analyzed with linear models.
-Additional packages used for calculations, analysis and figures included tidyverse (1.2.1), maps, mapdata, scatterpie…(update package list as figures come together).
+Additional packages used for calculations, analysis and figures included tidyverse (`r packageVersion("tidyverse")`), maps (`r packageVersion("maps")`), mapdata (`r packageVersion("mapdata")`), scatterpie (`r packageVersion("scatterpie")`)…[update package list as figures come together].
 
@@
 Latitude-specific diapause incidence (the fraction of diapausing eggs produced under unambiguous short-day conditions) did not significantly differ between 2008 and 2018 in Japan or the US (Figure 1, Table 2).
-Diapause incidence was high (generally greater than 90%) at latitudes above ~32.5 degrees in both countries (Figure 2A).
+Diapause incidence was high (generally greater than 90%) at latitudes north of ~32.5 degrees in both countries (Figure 2A).
 Substantial fractions of constitutively non-diapausing eggs were observed in populations south of South Carolina in the US and in Okinawa in Japan.
@@
 
-```{r fig1, fig.cap="Figure 1: Diapause incidence in Japan and US in 2008 and 2018", echo=FALSE}
+```{r fig1, fig.pos="H", fig.cap="Figure 1: Diapause incidence in Japan and US in 2008 and 2018", echo=FALSE}
 incidence_lat_year_plot
@@
 
-As observed in
+CPP values correlated strongly with latitude in the US and Japan.
+Latitudinal clines were nearly parallel to those observed in 2008 (Figure 3), as verified by the lack of a significant interaction term between year and latitude (Table 3).
+
+```{r fig3, fig.cap="Figure 3: Critical photoperiod (CPP) as a function of latitude in the US and Japan in 1988, 2008, and 2018.
+Error bars represent standard errors of CPP estimates.
+Errors around 1988 CPP estimates are absent because experiment and calculations were conducted using different methods and raw data were not available.", echo=FALSE, message=FALSE}
+cpp_japan_us_years
+```
+
+```{r table3, echo=FALSE}
+kable(anova(cppmodel), caption="Table 3: Effects of latitude, country and year on critical photoperiod in 2008 and 2018.")
+```
+
+However, the intercepts of the latitudinal clines significantly differed between both the US and Japan, and between 2008 and 2018 (Figure 4).
+Latitude-specific CPP was higher in Japan than the US, and higher in 2018 compared to 2008.
+This implies earlier entry into diapause at the same latitude in Japan than in the US and earlier entry into diapause in the same locations in 2018 than in 2008.
+
+```{r fig4, fig.cap="Figure 4: Shifts in critical photoperiod latitudinal clines. Lines represent linear regressions; dotted=2018, solid=2008.", echo=FALSE}
+cpp_08_18_shifts
+```
+
+## Supplementary Figures
+```{r figS1, fig.cap="Figure S1: Photoperiod response curves", echo=FALSE}
+DI_plot
+```

### 2019-10-25 MS.Rmd
Maps and methods

@@
 ---
-title: "Ae albopictus CPP manuscript"
+title: "Ae. albopictus CPP manuscript"
 header-includes: #allows you to add in your own Latex packages
@@
 output:
+  word_document: default
+  html_document: default
   pdf_document: default
-  html_document: default
-  word_document: default
 ---
@@
 install.packages("ggedit")
+install.packages("ggrepel")
 install.packages("cowplot")
@@
 library("ggedit")
+library("ggrepel")
 library("cowplot")
+```
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(dpi=300)
 ```
@@
 us_map<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="blue") + 
-  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 1) + xlim(c(-90, -65)) +
+  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 1) + xlim(c(-85, -65)) +
   geom_label(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
@@
 japan_map<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="red") + 
-  coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 1)  +
-  geom_label(data=filter(pops, country=="JP"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
+  coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 1)  + xlim(c(125, 147)) +
+  geom_label_repel(data=filter(pops, country=="JP", only_2008=="n"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
   theme_bw() + theme(axis.title.x=element_blank(), axis.title.y=element_blank())
@@
 
-us_DI_pies<-remove_geom(us_map, "label") + geom_scatterpie(data=filter(DI_info, country=="US"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + 
-  facet_grid(~year) + theme(axis.title.x=element_blank(), axis.title.y=element_blank())
+us_DI_pies<-remove_geom(us_map, "label") + geom_scatterpie(data=filter(DI_info, country=="US"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + facet_grid(~year) + theme(axis.title.x=element_blank(), axis.title.y=element_blank())
 
@@
 
-DI_pies_2018<-plot_grid(remove_geom(us_map, "label") + geom_scatterpie(data=filter(DI_info, country=="US", year=="2018"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "cm")), remove_geom(japan_map, "label") + geom_scatterpie(data=filter(DI_info, country=="JP", year=="2018"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "cm")), ncol=1, align="hv")
+DI_pies_2018<-plot_grid(remove_geom(us_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="US", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none"),
+
+remove_geom(japan_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="JP", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none"), ncol=2, align="hv")
 ```
@@
 - Diapause is an important overwintering adaptation in many organisms
-- This adaptation comes with trade-offs, but exactly what they are is unclear (winter phenotypes of diapausing organisms; life history strategies)
+- This adaptation comes with trade-offs, but exactly what they are is not well specified (discuss winter phenotypes of diapausing organisms; life history strategies)
 - Climate change may have unpredictable effects on growing seasons
@@
 Locations in the US were largely identical to locations sampled in 2008.
-[More details about field sampling.] At least 50 male and 50 female mosquitoes were used to initiate laboratory populations, except for SEN, which had X females and Y males.
+[More details about field sampling.] At least 50 male and 50 female mosquitoes were used to initiate laboratory populations, except for SEN, which had 37 females and 39 males initially.
 
@@
 Populations generally had at least 50 female and 50 male adults per generation, but the JAC population experienced visibly high mortality during the F2 generation.
-The larval generation immediately prior to the experiment was reared in 15 cm plastic petri dishes containing 30-40 larvae each, X dishes per population.
+The larval generation immediately prior to the experiment was reared in 15 cm plastic petri dishes containing 30-40 larvae each, 10 dishes per population.
 Additional details of mosquito husbandry can be found in Supplementary File 1.
@@
 At beginning of each block, eggs were hatched by submerging oviposition papers in water with larval food for two days.
-Larvae were distributed among X 15 cm petri dishes containing 90 ml of a mixture of distilled water and larval food (5 L of water to 10 mL of food slurry) at a density of 30-40 larvae per dish.
+Larvae were distributed among 40 15 cm petri dishes containing 90 ml of a mixture of distilled water and larval food (5 L of water to 10 mL of food slurry) at a density of 30-40 larvae per dish.
 Larvae were transferred to clean dishes containing fresh water/food mixture three times a week until pupation.
@@
 
-Pupae from each population were collected, pooled across dishes, and haphazardly distributed amongst thirteen 1.5 L adult cages, corresponding to 12 photoperiod cabinets and an additional short-day (8h light) incubator.
-The total number of pupae added to each cage ranged from X to Y; subsequent pupal and adult mortality was not monitored.
+Pupae from each population were collected, pooled across dishes, and split amongst thirteen 1.5 L adult cages, corresponding to 12 photoperiod cabinets and an additional short-day (8h light) incubator.
+The total number of pupae added to each cage ranged from 52 to 106; pupae were not sexed and subsequent pupal and adult mortality was not monitored.
 Approximately one week after the last pupae were added, cages were blood-fed, and a dark oviposition cup half-filled with water and lined with unbleached paper towel was added to each cage.
@@
 ```{r table2, echo=FALSE}
-kable(anova(dimodel), digits=3, caption="Table 2")
+kable(anova(dimodel), digits=3, caption="Table 2: Effects of latitude, country and year (2008 vs 2018) on diapause incidence.")
 ```
 
-```{r fig2, fig.cap="Figure 2: Diapause incidence (2018) by country.
-Black sections of pie charts represent diapausing fraction; white fractions represent non-diapausing.
-Blue points represent US diapause incidence; red points represent Japan.", echo=FALSE, message=FALSE}
-plot_grid(DI_pies_2018, country_DI_2018, ncol=2, labels="AUTO")
+```{r fig2A, fig.cap="Figure 2A: Diapause incidence (2018) by country. Dark sections of pie charts represent diapausing fraction; white fractions represent non-diapausing.", echo=FALSE, message=FALSE}
+DI_pies_2018
+```
+
+```{r fig.cap="Figure 2B: Diapause incidence (2018) by country. Blue points represent US diapause incidence; red points represent Japan.", echo=FALSE, message=FALSE}
+country_DI_2018
 ```
@@
 Latitudinal clines were nearly parallel to those observed in 2008 (Figure 3), as verified by the lack of a significant interaction term between year and latitude (Table 3).
+However, the intercepts of the latitudinal clines significantly differed between both the US and Japan, and between 2008 and 2018 (Figure 4).
+Latitude-specific CPP was higher in Japan than the US, and higher in 2018 compared to 2008.
+This implies earlier entry into diapause at the same latitude in Japan than in the US and earlier entry into diapause in the same locations in 2018 than in 2008.
 
@@
 
-However, the intercepts of the latitudinal clines significantly differed between both the US and Japan, and between 2008 and 2018 (Figure 4).
-Latitude-specific CPP was higher in Japan than the US, and higher in 2018 compared to 2008.
-This implies earlier entry into diapause at the same latitude in Japan than in the US and earlier entry into diapause in the same locations in 2018 than in 2008.
-
 ```{r fig4, fig.cap="Figure 4: Shifts in critical photoperiod latitudinal clines. Lines represent linear regressions; dotted=2018, solid=2008.", echo=FALSE}

### 2019-10-25 MS.Rmd
plots

@@
 
-DI_pies_2018<-plot_grid(remove_geom(us_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="US", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none"),
+DI_pies_2018<-plot_grid(remove_geom(us_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="US", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "null")),
 
-remove_geom(japan_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="JP", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none"), ncol=2, align="hv")
+remove_geom(japan_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="JP", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "null")), ncol=2, align="hv")
 ```

### 2019-10-30 MS.Rmd
Updating draft

@@
   pdf_document: default
+bibliography: MyLibrary.bib
 ---
@@
 ```{r message=FALSE, echo=FALSE, include=FALSE}
+library("papeR")
 library("drc") #3.0.1
@@
 blockAB_filtered <- filter(blockAB, yield<=1) %>% filter(yield>=0.6) 
+#test removing papers that had premature larval hatch on paper
+#badpapers<-c("333-1", "334-1", "335-1", "336-1", "337-1", "338-1", "339-1", "340-1", "341-1", "342-1", "343-1","366-1", "367-1", "368-1", "369-1", "370-1", "371-1", "372-1", "373-1", "374-1", "375-1", "376-1")
+#blockAB_filtered <- filter(blockAB, yield<=1) %>% filter(yield>=0.6) %>% filter(!PaperID %in% badpapers)
 ```
@@
   group_by(CageID) %>% 
-  summarise(pop=first(Population), pp=mean(Photoperiod), lat=mean(Lat), block=first(Block), country=first(Country), h1=sum(H1Count), h2=sum(H2Count), emb=sum(EmbryoCount)) %>% 
-  filter(h1+h2+emb>100)
+  summarise(pop=first(Population), pp=mean(Photoperiod), lat=mean(Lat), block=first(Block), country=first(Country), h1=sum(H1Count), h2=sum(H2Count), emb=sum(EmbryoCount)) %>%   filter(h1+h2+emb>100)
 DI_summary$total<-DI_summary$h1+DI_summary$h2+DI_summary$emb
@@
 #plot with DI at 8 h shown as TWO DIFFERENT points (corresponding to two cabinets)
-DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~pop)
+DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + stat_smooth(method="drm", se=FALSE) + facet_wrap(~pop)
 
@@
 cpp_values<-data.frame(ED(d.model.fit, 50))
+```
+
+```{r}
+
 ```
@@
 country_DI_2018<-ggplot(DI_short_2018, aes(x=lat, y=DI, color=country)) + scale_color_manual(values=c("red","blue")) + geom_point() + theme_bw() + stat_smooth() + theme(legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+country_DI<-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country)) + scale_color_manual(values=c("red","blue")) + geom_point() + theme_bw() + stat_smooth() + theme(legend.position=c(0.9,0.1)) + facet_grid(~year)
+
 ```
@@
 - Diapause is an important overwintering adaptation in many organisms
-- This adaptation comes with trade-offs, but exactly what they are is not well specified (discuss winter phenotypes of diapausing organisms; life history strategies)
+- Photoperiod provides a reliable signal of time of year and so is used by many organisms to regulate seasonal strategies
+- The optimal time of year to go into diapause might depend on a lot of different climactic factors and fitness considerations
 - Climate change may have unpredictable effects on growing seasons
@@
 - What is the continuing role of diapause in light of environmental change, continuing introductions, competing selective pressures? 
-- We recapitulated 2008 CPP study
+- We recapitulated 2008 CPP study using the same experimental design and methods
 
@@
 
-### Mosquito population sampling and rearing
-Mosquito populations were sampled in the eastern USA and Japan in 2018, at latitudes ranging from `r min(pops$latitude)` to `r max(pops$latitude)`.
+### Mosquito population sampling and pre-experiment colony maintenance
+Mosquito populations were sampled as larvae and pupae in the eastern USA and Japan in 2018, at latitudes ranging from `r min(pops$latitude)` to `r max(pops$latitude)`.
 Sampling locations are described in Table 1.
-Locations in the US were largely identical to locations sampled in 2008.
-[More details about field sampling.] At least 50 male and 50 female mosquitoes were used to initiate laboratory populations, except for SEN, which had 37 females and 39 males initially.
+Locations in the US were identical to locations sampled in 2008, except: .
+At least 50 male and 50 female mosquitoes were used to initiate laboratory populations, except for SEN, which had 37 females and 39 males initially.
 
@@
 
-Populations were maintained in the laboratory for 3 or 4 generations prior to the experiment at 21C and 80% humidity under a 16:8h light:dark cycle.
-Larvae were reared in distilled water and fed a slurry of dog food and brine shrimp, while adults were fed organic raisins and blood-fed by human arm.
-Initial laboratory generations of larvae were reared in plastic storage containers, with one population per container and densities varying by population.
+Populations were maintained in the laboratory for 3 or 4 generations prior to the experiment at 21C and 80% relative humidity under a 16:8h light:dark cycle, following previous studies [@armbrusterGeographicVariationLarval2006; @urbanskiRapidAdaptiveEvolution2012].
+Larvae were reared in distilled water and fed a slurry of dog food and brine shrimp.
+Adults were provided organic raisins as a carbohydrate source and blood meals from a human host.
+Initial laboratory generations of larvae were reared in plastic storage containers containing ~2.5 L of distilled water, with one population per container at initial densities of ~200-300.
 Pupae were then transferred into one adult cage per population.
 Populations generally had at least 50 female and 50 male adults per generation, but the JAC population experienced visibly high mortality during the F2 generation.
-The larval generation immediately prior to the experiment was reared in 15 cm plastic petri dishes containing 30-40 larvae each, 10 dishes per population.
+The larval generation immediately prior to the experiment was reared in 15 cm plastic petri dishes containing 30-40 larvae each, 10 dishes per population, to match the larval rearing conditions of the experimental generation, which followed the methods of Urbanski et al 2012.
 Additional details of mosquito husbandry can be found in Supplementary File 1.
@@
 ### Critical photoperiod determination
-To determine the critical photoperiod of each population, the following general procedure was used: Starting from the pupal stage, mosquitoes were divided among 12 different photoperiods and allowed to produce eggs.
+An overview of the procedure used to determine the critical photoperiod is as follows: Starting from the pupal stage, mosquitoes were divided among 12 different photoperiods (listed in Supplementary File 1) and allowed to produce eggs.
 The proportion of diapause eggs at each photoperiod was determined, and a dose-response curve was constructed with respect to hours of light and proportion of eggs in diapause.
 The critical photoperiod was calculated as the number of hours of light at which 50% of the eggs produced were in diapause, excluding the fraction of constitutively non-diapausing eggs (i.e.
-the fraction that did not enter diapause at 8 hours of light).
+the fraction that did not enter diapause under unambigous short-day conditions, 8 hours of light).
 The experimental design and most of the equipment and experimental details were the same as those used in Urbanski et al 2012; additional details on all steps can be found in Supplementary File 1.
@@
 Due to the large number of populations and photoperiods, the experiment was conducted in two blocks.
-Each block contained half of the populations from each country; populations across the full range of latitudes were interspersed between the blocks.
-At beginning of each block, eggs were hatched by submerging oviposition papers in water with larval food for two days.
+Each block contained half of the populations from each country; populations across the full range of latitudes from each country were interspersed between the blocks.
+At the beginning of each block, eggs were hatched by submerging oviposition papers in water with larval food for two days.
 Larvae were distributed among 40 15 cm petri dishes containing 90 ml of a mixture of distilled water and larval food (5 L of water to 10 mL of food slurry) at a density of 30-40 larvae per dish.
-Larvae were transferred to clean dishes containing fresh water/food mixture three times a week until pupation.
-Larvae were kept in the controlled temperature room at 21 degrees, 80 percent humidity and 16:8 hours light:dark until they produced pupae.
+Larvae were transferred to clean dishes containing fresh water/food mixture three times a week (MWF) until pupation.
+Larvae were kept in the controlled temperature room at 21C, 80% humidity and 16:8 hours light:dark until they produced pupae.
 
-Pupae from each population were collected, pooled across dishes, and split amongst thirteen 1.5 L adult cages, corresponding to 12 photoperiod cabinets and an additional short-day (8h light) incubator.
+Pupae from each population were collected, pooled across dishes, and split amongst thirteen 1.5 L adult cages, corresponding to 12 photoperiod cabinets and an additional short-day (8h light) incubator to provide additional data on unambiguous short-day diapause incidence.
 The total number of pupae added to each cage ranged from 52 to 106; pupae were not sexed and subsequent pupal and adult mortality was not monitored.
 Approximately one week after the last pupae were added, cages were blood-fed, and a dark oviposition cup half-filled with water and lined with unbleached paper towel was added to each cage.
-Oviposition papers were collected 5-7 days after blood-feeding and replaced every 2-3 days; a total of four oviposition papers were collected from each cage.
+Oviposition papers were collected 5-7 days after blood-feeding and replaced every MWF; a total of four oviposition papers were collected from each cage.
 Several cages were fed again and an additional oviposition paper was collected; see Supplementary File 1 for details.
-Oviposition papers were dried gently and stored in the photoperiod cabinets for 10 to 21 days until the diapause scoring procedure.
+Oviposition papers were dried gently and stored in the photoperiod cabinets from which they originated for 11 to 21 days until diapause determination.
 
 Oviposition papers were trimmed to a ~2 cm strip surrounding what was the waterline in the oviposition cup; this zone generally contained the vast majority of the eggs.
-On papers with few eggs, the paper was trimmed in such a way as to capture all eggs.
 On papers with very dense oviposition, the egg-containing strip was haphazardly subsampled further, to several hundred eggs as approximated by eye.
 Eggs were counted and egg papers were submerged in distilled water with several drops of larval food.
-Larvae that hatched were counted two days later and papers were re-dried.
-After 1-2 weeks, hatch was induced again and second installment hatched larvae were counted.
+Larvae that hatched were counted two days later and papers were re-dried and stored as before.
+After 1-2 weeks, hatch was induced again and larvae were counted.
 Oviposition papers were then treated overnight with a bleach solution to clear the chorion of the remaining eggs.
@@
 
-For each cage, the counts of larvae and embryos across all oviposition papers were pooled to calculate the diapause fraction: number of embryonated eggs divided by total number of viable eggs.
+For each cage, the counts of larvae and embryos across all oviposition papers were pooled to calculate the diapause incidence: number of embryonated eggs divided by total number of viable eggs (# embryonated / (# embryonated + # hatched)).
 The median total number of viable eggs per cage was `r median(DI_summary$total)` (range `r min(DI_summary$total)`-`r max(DI_summary$total)`).
@@
 
-Photoperiod response data from 2008 and calculated CPP values from 1988 were obtained from Urbanski et al 2012 data files (doi:10.5061/dryad.68277435).
+Photoperiod response data from 2008 and calculated CPP values from 1988 were obtained from the data files of Urbanski et al 2012 (doi:10.5061/dryad.68277435).
 
@@
 
-Latitude-specific diapause incidence (the fraction of diapausing eggs produced under unambiguous short-day conditions) did not significantly differ between 2008 and 2018 in Japan or the US (Figure 1, Table 2).
-Diapause incidence was high (generally greater than 90%) at latitudes north of ~32.5 degrees in both countries (Figure 2A).
+Latitude-specific diapause incidence (the fraction of diapausing eggs produced under unambiguous short-day conditions) did not significantly differ between 2008 and 2018 in Japan or the US (Figure 2, Table 2).
+Diapause incidence was high (generally greater than 90%) at latitudes north of ~32.5 degrees in both countries.
 Substantial fractions of constitutively non-diapausing eggs were observed in populations south of South Carolina in the US and in Okinawa in Japan.
 Diapause incidence followed a latitudinal cline in the US, but a cline could not be observed in Japan due to absence of land to sample between Okinawa and Kagoshima.
-In the limited number of southern populations that could be compared between the US and Japan, latitude-specific diapause incidence appeared to be lower in the US (Figure 2B).
+In the limited number of southern populations that could be compared between the US and Japan, latitude-specific diapause incidence appeared to be lower in the US.
 
-```{r fig1, fig.pos="H", fig.cap="Figure 1: Diapause incidence in Japan and US in 2008 and 2018", echo=FALSE}
+```{r fig1, fig.pos="H", fig.cap="Figure 1: Diapause incidence in Japan and US in 2008 and 2018", echo=FALSE, include=FALSE}
 incidence_lat_year_plot
@@
 ```{r table2, echo=FALSE}
-kable(anova(dimodel), digits=3, caption="Table 2: Effects of latitude, country and year (2008 vs 2018) on diapause incidence.")
+kable(prettify(anova(dimodel)), digits=3, caption="Table 2: Effects of latitude, country and year (2008 vs 2018) on diapause incidence.")
 ```
 
-```{r fig2A, fig.cap="Figure 2A: Diapause incidence (2018) by country. Dark sections of pie charts represent diapausing fraction; white fractions represent non-diapausing.", echo=FALSE, message=FALSE}
-DI_pies_2018
-```
+```{r fig.cap="Figure 2B: Diapause incidence by country and year.", echo=FALSE, message=FALSE}
+#country_DI
+ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point() + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year), se=FALSE)
 
-```{r fig.cap="Figure 2B: Diapause incidence (2018) by country. Blue points represent US diapause incidence; red points represent Japan.", echo=FALSE, message=FALSE}
-country_DI_2018
 ```
@@
 Latitudinal clines were nearly parallel to those observed in 2008 (Figure 3), as verified by the lack of a significant interaction term between year and latitude (Table 3).
-However, the intercepts of the latitudinal clines significantly differed between both the US and Japan, and between 2008 and 2018 (Figure 4).
+However, the intercepts of the latitudinal clines significantly differed between both the US and Japan, and between 2008 and 2018.
 Latitude-specific CPP was higher in Japan than the US, and higher in 2018 compared to 2008.
@@
 Error bars represent standard errors of CPP estimates.
-Errors around 1988 CPP estimates are absent because experiment and calculations were conducted using different methods and raw data were not available.", echo=FALSE, message=FALSE}
+Errors around 1988 CPP estimates are absent because experiment and calculations were conducted using different methods and raw data were not available.", echo=FALSE, message=FALSE, include=FALSE}
 cpp_japan_us_years
@@
 ```{r table3, echo=FALSE}
-kable(anova(cppmodel), caption="Table 3: Effects of latitude, country and year on critical photoperiod in 2008 and 2018.")
+kable(prettify(anova(cppmodel)), caption="Table 3: Effects of latitude, country and year on critical photoperiod in 2008 and 2018.")
 ```
@@
 ## Supplementary Figures
+
+```{r}
+pl<-drcplot(d.model.fit, log="")
+```
+
 ```{r figS1, fig.cap="Figure S1: Photoperiod response curves", echo=FALSE}
@@
 ```
+```{r figS2, fig.cap="Figure 2A: Diapause incidence (2018) by country. Dark sections of pie charts represent diapausing fraction; white fractions represent non-diapausing.", echo=FALSE, message=FALSE}
+DI_pies_2018
+```
+# References

### 2019-11-01 MS.Rmd
methods updates

@@
 install.packages("ggrepel")
+install.packages("papeR")
 install.packages("cowplot")
+
 ```

### 2019-11-01 MS.Rmd
edits

@@
 hist(DI_summary$total, main="Total number of viable eggs per cage")
-
 ```
@@
 #plot with DI at 8 h shown as TWO DIFFERENT points (corresponding to two cabinets)
-DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + stat_smooth(method="drm", se=FALSE) + facet_wrap(~pop)
+DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~pop)
 
@@
 cpp_08_18_shifts<-ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, color=country, shape=year)) + 
-  geom_point() + scale_color_manual(values=c("red", "blue")) +
-  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
-  stat_smooth(method=lm, se=FALSE, aes(linetype=year)) + theme_bw() + theme(legend.position=c(.8,.25))
+  geom_point(size=3) + scale_color_manual(values=c("red", "blue")) +
+  stat_smooth(method=lm, se=FALSE, aes(linetype=year)) + scale_linetype_manual(values=c("solid","dotted")) + theme_bw() + theme(legend.position=c(.8,.25))
+  # + geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) 
 ```
@@
 Locations in the US were identical to locations sampled in 2008, except: .
-At least 50 male and 50 female mosquitoes were used to initiate laboratory populations, except for SEN, which had 37 females and 39 males initially.
+Japanese population samples were sent to Georgetown as F1 eggs (except for TAK, for which F2 eggs were provided).
+At least 50 male and 50 female mosquitoes were used to initiate the laboratory populations at Georgetown, except for SEN, which had 37 females and 39 males.
 
@@
 Each block contained half of the populations from each country; populations across the full range of latitudes from each country were interspersed between the blocks.
-At the beginning of each block, eggs were hatched by submerging oviposition papers in water with larval food for two days.
+At the beginning of each block, eggs were hatched by submerging oviposition papers in distilled water with larval food for two days.
 Larvae were distributed among 40 15 cm petri dishes containing 90 ml of a mixture of distilled water and larval food (5 L of water to 10 mL of food slurry) at a density of 30-40 larvae per dish.
@@
 #country_DI
-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point() + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year), se=FALSE)
+ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year), se=FALSE) + scale_linetype_manual(values=c("solid","dotted"))
 
@@
 
-```{r}
-pl<-drcplot(d.model.fit, log="")
-```
-
 ```{r figS1, fig.cap="Figure S1: Photoperiod response curves", echo=FALSE}
@@
 ```
+
 ```{r figS2, fig.cap="Figure 2A: Diapause incidence (2018) by country. Dark sections of pie charts represent diapausing fraction; white fractions represent non-diapausing.", echo=FALSE, message=FALSE}
@@
 ```
+
+## Supplementary methods
+
+### Larval rearing
+Larvae were fed a slurry of 40 g frozen brine shrimp and 120 g dog food homogenized in 1 L of distilled water.
+Larvae were initially raised in plastic Sterilite storage containers containing 1.5 L of distilled water, for ease of bulk rearing.
+Every Monday-Wednesday-Friday (MWF), the water in each container was removed by straining through a net and replaced with fresh distilled water; each container was fed approximately 1.5 ml of larval food slurry.
+Containers were covered with a fine mesh and kept at 21C, 80% humidity, and 16:8 L:D photoperiod.
+Starting with the generation prior to the one used in the experiment, larvae were reared in 15 cm petri dishes, as in Urbanski et al 2012.
+The pre-experiment generation was fed ~0.5-0.6 ml larval food per dish, while the experimental generation had food/water concentrations as described in the main manuscript.
+Every MWF, larvae were transferred to clean petri dishes with a pipette and had fresh food/water added.
+When removing pupae substantially reduced the density of larvae, the remaining larvae were consolidated into fewer dishes.
+
+### Photoperiod cabinets 
+The custom-built photoperiod cabinets were the same ones used in Urbanski et al 2012, described in detail in the Appendix (https://www.journals.uchicago.edu/doi/suppl/10.1086/664709/suppl_file/53244apa.pdf).
+The light-dark cycles used were 8L : 16D, 12L : 12D, 12.5L : 11.5D, 12.75L : 11.25D, 13L : 11D, 13.25L : 10.75D, 13.5L : 10.5D, 13.75L : 10.25D, 14L : 10D, 14.25L : 9.75D, 14.5L : 9.5D, and 16L : 8D.
+Light and temperature inside cabinets were recorded for at least three full days for each cabinet in each block.
+The highest recorded temperature in any cabinet at any point was 22.65C, the lowest was 20.72C, and 70% of temperature readings were between 21C and 22C.
+All light:dark cycles were within 1 minute of desired photoperiod except for the separate short-day incubator (Cabinet 13 in Data Table 1), which was within 5 minutes of 8 hours of light (7:59 -8:05).
+Cages in photoperiod cabinets were monitored daily for desiccation and mold; water was added to cages and raisins were replaced as needed.
+
+### Blood meals, egg collection and storage
+Adult females were allowed to blood feed on a human host over a three-day period; feeding days were randomly assigned by photoperiod cabinet.
+Collected oviposition papers were stored in 10 cm petri dishes inside humidified containers in the photoperiod cabinet in which they originated.
+Two cages in the second block of the experiment produced less than 200 eggs after initial blood-feeding.
+Two cages in the second block of the experiment produced less than 200 eggs after initial blood meal, and the cages in the cabinets with 14.5 hours and 14.25 hours of light had large numbers of larvae that hatched on the wet oviposition paper on the first egg collection day due to incomplete drying and thus could not be accurately counted.
+These 24 cages were offered a second blood meal and an additional oviposition paper was collected a week later.
+Papers with eggs produced after a second blood meal have a Paper ID ending in “-5” in Data Table 1.
+
+### Diapause determination
+The diagnostic criteria of visible eye spots and abdominal segmentation were used to score bleached eggs positive for the presence of pharate larvae.
+On some oviposition papers, larvae “hatched” out of the eggs after the bleach solution was added; these were scored as embryonated diapause eggs because they appeared to be morphologically normal and did not hatch when induced with standard hatching cues twice.
+
+The total number of hatched larvae and embryos could differ from the number of eggs initially counted on the paper for the following reasons: 1) presence of inviable eggs or larvae (i.e., eggs that fail to hatch due to developmental abnormalities or larvae that die before being counted), 2) loss of eggs while handling oviposition papers, or 3) miscounting at any step of the scoring process.
+Based on the distribution of percent differences from the initial counts, we considered oviposition papers with over 100% or less than 60% of the initial count to be less reliable data points.
+Excluding these data points resulted in 5 cages out of 286 in the total data set having less than 100 remaining eggs, so they were excluded from the photoperiod response curves.
+These filtering steps and exclusions did not affect CPP estimates for any population except PBS; for this population, excluding the unreliable counts eliminated an outlier in the photoperiod response curve and brought the CPP estimate into the range of CPP values for the rest of the populations.
+
+```{r}
+plot(blockAB$yield ~ blockAB$TotalLarvaeEmbryos, xlab="Number of viable eggs on oviposition paper", ylab="Viable eggs/initial egg count on paper")
+```
+
 # References

### 2019-11-05 MS.Rmd
updated columns

@@
 #plot with DI at 8 h shown as TWO DIFFERENT points (corresponding to two cabinets)
+c("PBS", "FEM", "OAK", "JAC", "BRU", "ZIO", "FAY", "NVA", "WAV", "MAN", "BER", "NEW", "OKI", "TAN", "KAG", "YAT", "SAG", "SHI", "YAM", "HIR", "KYO", "TAK", "TOK", "UTS", "KAN", "KHO", "AIZ", "NIG", "SEN", "SAK")
+DI_summary$pop<-factor(DI_summary$pop, levels=c("PBS", "FEM", "OAK", "JAC", "BRU", "ZIO", "FAY", "NVA", "WAV", "MAN", "BER", "NEW", "OKI", "TAN", "KAG", "YAT", "SAG", "SHI", "YAM", "HIR", "KYO", "TAK", "TOK", "UTS", "KAN", "KHO", "AIZ", "NIG", "SEN", "SAK"))
 DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~pop)
@@
 
+### Larval collection in Japan
+
 ### Larval rearing

### 2019-11-06 MS.Rmd
added supp fig

@@
 Sampling locations are described in Table 1.
-Locations in the US were identical to locations sampled in 2008, except: .
+Locations in the US were identical to locations sampled in 2008.
 Japanese population samples were sent to Georgetown as F1 eggs (except for TAK, for which F2 eggs were provided).
-At least 50 male and 50 female mosquitoes were used to initiate the laboratory populations at Georgetown, except for SEN, which had 37 females and 39 males.
+At least 50 male and 50 female adult mosquitoes were used to initiate all populations in the laboratory at Georgetown, except for SEN, which had 37 females and 39 males.
 
@@
 The larval generation immediately prior to the experiment was reared in 15 cm plastic petri dishes containing 30-40 larvae each, 10 dishes per population, to match the larval rearing conditions of the experimental generation, which followed the methods of Urbanski et al 2012.
-Additional details of mosquito husbandry can be found in Supplementary File 1.
+Additional details of mosquito husbandry prior to the experiment can be found in Supplementary File 1.
 
@@
 the fraction that did not enter diapause under unambigous short-day conditions, 8 hours of light).
-The experimental design and most of the equipment and experimental details were the same as those used in Urbanski et al 2012; additional details on all steps can be found in Supplementary File 1.
+This experimental design and most of the equipment and experimental details were the same as those used in Urbanski et al 2012; additional details on all steps can be found in Supplementary File 1.
 
@@
 
-Pupae from each population were collected, pooled across dishes, and split amongst thirteen 1.5 L adult cages, corresponding to 12 photoperiod cabinets and an additional short-day (8h light) incubator to provide additional data on unambiguous short-day diapause incidence.
+Pupae from each population were collected, pooled across dishes, and split amongst thirteen 1.5 L adult cages, corresponding to twelve photoperiod cabinets and an additional short-day (8h light) incubator to provide additional data on unambiguous short-day diapause incidence.
 The total number of pupae added to each cage ranged from 52 to 106; pupae were not sexed and subsequent pupal and adult mortality was not monitored.
@@
 ```
+```{r}
+#comparing CPP values calculated with and without filtering out unreliable oviposition papers
+cpp_filtered<-read.csv("cpp_calcs_filtered.csv") %>% filter(year==2018)
+cpp_unfiltered<-read.csv("cpp_calcs_unfiltered.csv") %>% filter(year==2018)
+names(cpp_filtered)[5]<-"CPP_est_filt"
+names(cpp_unfiltered)[5]<-"CPP_est_unfilt"
+cpp_compare<-left_join(cpp_filtered,cpp_unfiltered,by="pop")
+ggplot(cpp_compare, aes(x=CPP_est_unfilt, y=CPP_est_filt)) + geom_point() + geom_abline(slope=1, intercept=0)
+
+```
 

### 2019-11-08 MS.Rmd
Added populations info

@@
 DI_summary$pop<-factor(DI_summary$pop, levels=c("PBS", "FEM", "OAK", "JAC", "BRU", "ZIO", "FAY", "NVA", "WAV", "MAN", "BER", "NEW", "OKI", "TAN", "KAG", "YAT", "SAG", "SHI", "YAM", "HIR", "KYO", "TAK", "TOK", "UTS", "KAN", "KHO", "AIZ", "NIG", "SEN", "SAK"))
-DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~pop)
+DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~pop) + theme_bw() + xlab("Photoperiod (hours)") + ylab("Dipause incidence")
 
@@
 cppmodel<-lm(Estimate~lat*country*year, data=filter(cpp_2, year!=1988))
-anova(cppmodel)
+cppmodel_result<-data.frame(anova(cppmodel))
 ```
@@
   geom_point(size=3) + scale_color_manual(values=c("red", "blue")) +
-  stat_smooth(method=lm, se=FALSE, aes(linetype=year)) + scale_linetype_manual(values=c("solid","dotted")) + theme_bw() + theme(legend.position=c(.8,.25))
+  stat_smooth(method=lm, se=FALSE, aes(linetype=year)) + scale_linetype_manual(values=c("dotted","solid")) + theme_bw() + theme(legend.position=c(.8,.25)) + xlab("Latitude") + ylab("Critical photoperiod")
   # + geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) 
@@
 dimodel<-lm(DI~lat*country*year, data=DI_short_2)
-anova(dimodel)
+dimodel_result<-data.frame(anova(dimodel))
 ```
@@
 ### Mosquito population sampling and pre-experiment colony maintenance
-Mosquito populations were sampled as larvae and pupae in the eastern USA and Japan in 2018, at latitudes ranging from `r min(pops$latitude)` to `r max(pops$latitude)`.
+Mosquito populations were sampled as larvae and pupae in the eastern USA and Japan in June-September 2018, at latitudes ranging from `r min(pops$latitude)` to `r max(pops$latitude)`.
 Sampling locations are described in Table 1.
 Locations in the US were identical to locations sampled in 2008.
-Japanese population samples were sent to Georgetown as F1 eggs (except for TAK, for which F2 eggs were provided).
+Japanese population samples were confirmed to be Ae.
+albopictus via morphological identification following [@tanakaRevisionAdultLarval1979] and sent to Georgetown as F1 eggs (except for TAK, for which F2 eggs were provided).
 At least 50 male and 50 female adult mosquitoes were used to initiate all populations in the laboratory at Georgetown, except for SEN, which had 37 females and 39 males.
@@
 ```{r table1, echo=FALSE}
-kable(filter(pops, only_2008=="n")[,1:5], col.names=c("Population", "City", "Latitude", "Longitude", "Country"), caption="Table 1: 2018 Sampling Locations")
+kable(filter(pops, only_2008=="n")[,1:6], col.names=c("Population", "City", "Latitude", "Longitude", "Country", "Life Stage Sampled"), caption="Table 1: 2018 Sampling Locations")
 ```
@@
 Adults were provided organic raisins as a carbohydrate source and blood meals from a human host.
-Initial laboratory generations of larvae were reared in plastic storage containers containing ~2.5 L of distilled water, with one population per container at initial densities of ~200-300.
+Initial laboratory generations of larvae were reared in plastic storage containers containing ~2.5 L of distilled water, with one population per container at initial densities of ~200-300, for ease of bulk rearing.
 Pupae were then transferred into one adult cage per population.
@@
 ### Critical photoperiod determination
-An overview of the procedure used to determine the critical photoperiod is as follows: Starting from the pupal stage, mosquitoes were divided among 12 different photoperiods (listed in Supplementary File 1) and allowed to produce eggs.
+An overview of the procedure used to determine the critical photoperiod (CPP) is as follows: Starting from the pupal stage, mosquitoes were divided among 12 different photoperiods (listed in Supplementary File 1) and allowed to produce eggs.
 The proportion of diapause eggs at each photoperiod was determined, and a dose-response curve was constructed with respect to hours of light and proportion of eggs in diapause.
@@
 Larvae were distributed among 40 15 cm petri dishes containing 90 ml of a mixture of distilled water and larval food (5 L of water to 10 mL of food slurry) at a density of 30-40 larvae per dish.
-Larvae were transferred to clean dishes containing fresh water/food mixture three times a week (MWF) until pupation.
+Larvae were transferred to clean dishes containing fresh water/food mixture every Monday-Wednesday-Friday (MWF) until pupation.
 Larvae were kept in the controlled temperature room at 21C, 80% humidity and 16:8 hours light:dark until they produced pupae.
@@
 The total number of pupae added to each cage ranged from 52 to 106; pupae were not sexed and subsequent pupal and adult mortality was not monitored.
-Approximately one week after the last pupae were added, cages were blood-fed, and a dark oviposition cup half-filled with water and lined with unbleached paper towel was added to each cage.
-Oviposition papers were collected 5-7 days after blood-feeding and replaced every MWF; a total of four oviposition papers were collected from each cage.
-Several cages were fed again and an additional oviposition paper was collected; see Supplementary File 1 for details.
+Approximately one week after the last pupae were added, cages were offered a blood meal, and a dark oviposition cup half-filled with water and lined with unbleached paper towel was added to each cage.
+Oviposition papers were collected 5-7 days after the blood meal and replaced every MWF; a total of four oviposition papers were collected from each cage.
+Several cages were fed again and an additional oviposition paper was collected; see supplemental methods for details.
 Oviposition papers were dried gently and stored in the photoperiod cabinets from which they originated for 11 to 21 days until diapause determination.
 
-Oviposition papers were trimmed to a ~2 cm strip surrounding what was the waterline in the oviposition cup; this zone generally contained the vast majority of the eggs.
-On papers with very dense oviposition, the egg-containing strip was haphazardly subsampled further, to several hundred eggs as approximated by eye.
-Eggs were counted and egg papers were submerged in distilled water with several drops of larval food.
+On each oviposition paper, eggs were counted (subsampling if oviposition was very high; see supplemental methods) and papers were submerged in distilled water with ~0.25 ml larval food.
 Larvae that hatched were counted two days later and papers were re-dried and stored as before.
 After 1-2 weeks, hatch was induced again and larvae were counted.
-Oviposition papers were then treated overnight with a bleach solution to clear the chorion of the remaining eggs.
-Eggs containing a pharate larva were scored as embryonated and in diapause.
+Oviposition papers were then treated overnight with a bleach solution to clear the chorion of the remaining eggs [@trpisNewBleachingDecalcifying1970] .
+Eggs were examined under a stereomicroscope, and those eggs containing a pharate larva were scored as embryonated and in diapause.
 The total number of hatched larvae and embryonated eggs was considered the number of viable eggs.
@@
 
-For each cage, the counts of larvae and embryos across all oviposition papers were pooled to calculate the diapause incidence: number of embryonated eggs divided by total number of viable eggs (# embryonated / (# embryonated + # hatched)).
+For each cage, the counts of larvae and embryos across all oviposition papers were pooled to calculate the diapause incidence, or the fraction of total viable eggs that are in diapause (# embryonated / (# embryonated + # hatched)).
 The median total number of viable eggs per cage was `r median(DI_summary$total)` (range `r min(DI_summary$total)`-`r max(DI_summary$total)`).
 For the 8-hour photoperiod, each population had data from two cages: one in a photoperiod cabinet and one in a separate incubator.
-These data were pooled to calculate the diapause incidence value for the population, but used as separate data points in photoperiod response curves.
+These data were pooled to calculate the value for diapause incidence at unambiguous short-day (DI) for the population, but used as separate data points in photoperiod response curves.
 
-Photoperiod response data from 2008 and calculated CPP values from 1988 were obtained from the data files of Urbanski et al 2012 (doi:10.5061/dryad.68277435).
+Photoperiod response data from 2008 were obtained from the data files of Urbanski et al 2012 (doi:10.5061/dryad.68277435).
 
@@
 2018) on CPP and on DI were analyzed with linear models.
-Additional packages used for calculations, analysis and figures included tidyverse (`r packageVersion("tidyverse")`), maps (`r packageVersion("maps")`), mapdata (`r packageVersion("mapdata")`), scatterpie (`r packageVersion("scatterpie")`)…[update package list as figures come together].
+Data manipulation and figure construction were carried out using packages in tidyverse (`r packageVersion("tidyverse")`).
 
@@
 
-Latitude-specific diapause incidence (the fraction of diapausing eggs produced under unambiguous short-day conditions) did not significantly differ between 2008 and 2018 in Japan or the US (Figure 2, Table 2).
-Diapause incidence was high (generally greater than 90%) at latitudes north of ~32.5 degrees in both countries.
-Substantial fractions of constitutively non-diapausing eggs were observed in populations south of South Carolina in the US and in Okinawa in Japan.
-Diapause incidence followed a latitudinal cline in the US, but a cline could not be observed in Japan due to absence of land to sample between Okinawa and Kagoshima.
-In the limited number of southern populations that could be compared between the US and Japan, latitude-specific diapause incidence appeared to be lower in the US.
+Latitude-specific DI (the fraction of diapausing eggs produced under unambiguous short-day conditions) did not significantly differ between 2008 and 2018 in Japan or the US (p=0.94, Figure 2, Table 2).
+DI was high (generally greater than 90%) at latitudes north of ~32.5 degrees in both countries.
+Substantial fractions of constitutively non-diapausing eggs were observed in populations south of South Carolina in the US and in Okinawa in Japan (latitude effect p=<0.001).
+Latitude-specific DI was lower in the US than in Japan (country effect p=0.026), though there were fewer sites sampled at lower latitudes in Japan than the US due to lack of land between Okinawa and Kagashima.
 
-```{r fig1, fig.pos="H", fig.cap="Figure 1: Diapause incidence in Japan and US in 2008 and 2018", echo=FALSE, include=FALSE}
+```{r fig.cap="Figure 1: Diapause incidence in Japan and US in 2008 and 2018", echo=FALSE, include=FALSE}
 incidence_lat_year_plot
@@
 
-```{r table2, echo=FALSE}
+```{r fig.cap="Figure 2B: Diapause incidence by country and year.
+Lines represent locally estimated scatterplot smoothing (LOESS).
+Solid lines, 2018 data; dotted lines, 2008 data.", echo=FALSE, message=FALSE}
+#country_DI
+ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year), se=FALSE) + scale_linetype_manual(values=c("dotted","solid")) + xlab("Latitude") + ylab("Short-day diapause incidence")
+
+```
+
+```{r echo=FALSE}
 kable(prettify(anova(dimodel)), digits=3, caption="Table 2: Effects of latitude, country and year (2008 vs 2018) on diapause incidence.")
@@
 
-```{r fig.cap="Figure 2B: Diapause incidence by country and year.", echo=FALSE, message=FALSE}
-#country_DI
-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year), se=FALSE) + scale_linetype_manual(values=c("solid","dotted"))
+CPP values correlated strongly with latitude in the US and Japan (p<0.001, Table 3).
+Latitudinal clines were nearly parallel to those observed in 2008 (Figure 3), as verified by the lack of a significant interaction term between year and latitude (p=0.17).
+However, the intercepts of the latitudinal clines significantly differed between both the US and Japan, and between 2008 and 2018 (p<0.001 for both country and year).
+Latitude-specific CPP was higher in Japan than the US, and higher in 2018 compared to 2008.
+This implies earlier entry into diapause at the same latitude in Japan than in the US, as well as earlier entry into diapause in the same locations in 2018 than in 2008.
 
-```
-
-CPP values correlated strongly with latitude in the US and Japan.
-Latitudinal clines were nearly parallel to those observed in 2008 (Figure 3), as verified by the lack of a significant interaction term between year and latitude (Table 3).
-However, the intercepts of the latitudinal clines significantly differed between both the US and Japan, and between 2008 and 2018.
-Latitude-specific CPP was higher in Japan than the US, and higher in 2018 compared to 2008.
-This implies earlier entry into diapause at the same latitude in Japan than in the US and earlier entry into diapause in the same locations in 2018 than in 2008.
-
-```{r fig3, fig.cap="Figure 3: Critical photoperiod (CPP) as a function of latitude in the US and Japan in 1988, 2008, and 2018.
+```{r fig.cap="Figure 3: Critical photoperiod (CPP) as a function of latitude in the US and Japan in 1988, 2008, and 2018.
 Error bars represent standard errors of CPP estimates.
@@
 cpp_japan_us_years
+```
+
+```{r fig.cap="Figure 4: Shifts in critical photoperiod latitudinal clines. Lines represent linear regressions. Solid lines, 2018; dotted lines, 2008.", echo=FALSE}
+cpp_08_18_shifts
 ```
@@
 
-```{r fig4, fig.cap="Figure 4: Shifts in critical photoperiod latitudinal clines. Lines represent linear regressions; dotted=2018, solid=2008.", echo=FALSE}
-cpp_08_18_shifts
-```
-
-## Supplementary Figures
-
-```{r figS1, fig.cap="Figure S1: Photoperiod response curves", echo=FALSE}
-DI_plot 
-```
-
-```{r figS2, fig.cap="Figure 2A: Diapause incidence (2018) by country. Dark sections of pie charts represent diapausing fraction; white fractions represent non-diapausing.", echo=FALSE, message=FALSE}
-DI_pies_2018
-```
+## Supplementary File 1
 
@@
 
-### Larval collection in Japan
-
 ### Larval rearing
-Larvae were fed a slurry of 40 g frozen brine shrimp and 120 g dog food homogenized in 1 L of distilled water.
-Larvae were initially raised in plastic Sterilite storage containers containing 1.5 L of distilled water, for ease of bulk rearing.
+Larvae were fed a slurry of 40 g frozen brine shrimp (San Francisco Bay Brand) and 120 g dog food (Nutro Feed Clean Ultra Small Breed Puppy formula) homogenized in 1 L of distilled water.
+Larvae were initially raised in plastic storage containers (Sterilite) containing 2.5 L of distilled water.
 Every Monday-Wednesday-Friday (MWF), the water in each container was removed by straining through a net and replaced with fresh distilled water; each container was fed approximately 1.5 ml of larval food slurry.
 Containers were covered with a fine mesh and kept at 21C, 80% humidity, and 16:8 L:D photoperiod.
-Starting with the generation prior to the one used in the experiment, larvae were reared in 15 cm petri dishes, as in Urbanski et al 2012.
+Starting with the generation prior to the one used in the experiment, larvae were reared in 15 cm petri dishes (Kord-Valmark), as in Urbanski et al 2012.
 The pre-experiment generation was fed ~0.5-0.6 ml larval food per dish, while the experimental generation had food/water concentrations as described in the main manuscript.
@@
 ### Blood meals, egg collection and storage
-Adult females were allowed to blood feed on a human host over a three-day period; feeding days were randomly assigned by photoperiod cabinet.
+Adult cages were fed from a human host over a three-day period; feeding days were randomly assigned by photoperiod cabinet.
 Collected oviposition papers were stored in 10 cm petri dishes inside humidified containers in the photoperiod cabinet in which they originated.
-Two cages in the second block of the experiment produced less than 200 eggs after initial blood-feeding.
-Two cages in the second block of the experiment produced less than 200 eggs after initial blood meal, and the cages in the cabinets with 14.5 hours and 14.25 hours of light had large numbers of larvae that hatched on the wet oviposition paper on the first egg collection day due to incomplete drying and thus could not be accurately counted.
+In the second block of the experiment, two cages produced less than 200 eggs after the initial blood meal, and the cages in the cabinets with 14.5 hours and 14.25 hours of light had large numbers of larvae that hatched on the wet oviposition paper on the first egg collection day due to incomplete drying and thus could not be accurately counted.
 These 24 cages were offered a second blood meal and an additional oviposition paper was collected a week later.
@@
 ### Diapause determination
-The diagnostic criteria of visible eye spots and abdominal segmentation were used to score bleached eggs positive for the presence of pharate larvae.
+Oviposition papers were trimmed to a ~2 cm strip surrounding the former location of the waterline in the oviposition cup; this zone generally contained the vast majority of the eggs.
+On oviposition papers with few eggs, the paper was trimmed in such a way as to capture all of them.
+On papers with very dense oviposition, the egg-containing strip was haphazardly subsampled further, to several hundred eggs as approximated by eye.
+
+After eggs were bleached, presence of pharate larvae was determined using the diagnostic criteria of visible eye spots and abdominal segmentation.
 On some oviposition papers, larvae “hatched” out of the eggs after the bleach solution was added; these were scored as embryonated diapause eggs because they appeared to be morphologically normal and did not hatch when induced with standard hatching cues twice.
 
-The total number of hatched larvae and embryos could differ from the number of eggs initially counted on the paper for the following reasons: 1) presence of inviable eggs or larvae (i.e., eggs that fail to hatch due to developmental abnormalities or larvae that die before being counted), 2) loss of eggs while handling oviposition papers, or 3) miscounting at any step of the scoring process.
-Based on the distribution of percent differences from the initial counts, we considered oviposition papers with over 100% or less than 60% of the initial count to be less reliable data points.
+The total number of hatched larvae and embryos could differ from the number of eggs initially counted on the paper for the following reasons: 1) presence of inviable eggs or larvae (i.e., eggs that fail to hatch due to developmental abnormalities or larvae that die before being counted), 2) loss of eggs while handling oviposition papers, or 3) inaccurate counting at any step of the scoring process.
+Based on the distribution of percent differences from the initial counts (Figure S1), we considered oviposition papers with over 100% or less than 60% of the initial count to be less reliable data points.
 Excluding these data points resulted in 5 cages out of 286 in the total data set having less than 100 remaining eggs, so they were excluded from the photoperiod response curves.
-These filtering steps and exclusions did not affect CPP estimates for any population except PBS; for this population, excluding the unreliable counts eliminated an outlier in the photoperiod response curve and brought the CPP estimate into the range of CPP values for the rest of the populations.
+These filtering steps and exclusions did not affect CPP estimates for any population except PBS; for this population, excluding the unreliable counts eliminated an outlier in the photoperiod response curve and brought the CPP estimate into the range of CPP values for the rest of the populations (Figure S2).
+The final photoperiod response curves are shown in Figure S3.
 
-```{r}
+```{r fig.cap="Figure S1: Discrepancies between initial egg counts and final calculated egg counts on oviposition papers.", echo=FALSE}
 plot(blockAB$yield ~ blockAB$TotalLarvaeEmbryos, xlab="Number of viable eggs on oviposition paper", ylab="Viable eggs/initial egg count on paper")
 ```
-```{r}
+
+```{r fig.cap="Figure S2.
+Critical photoperiod calculated with and without filtering oviposition papers based on criteria described.
+Line represents x=y.
+Outlier point corresponds to population PBS.", echo=FALSE}
 #comparing CPP values calculated with and without filtering out unreliable oviposition papers
@@
 cpp_compare<-left_join(cpp_filtered,cpp_unfiltered,by="pop")
-ggplot(cpp_compare, aes(x=CPP_est_unfilt, y=CPP_est_filt)) + geom_point() + geom_abline(slope=1, intercept=0)
+ggplot(cpp_compare, aes(x=CPP_est_unfilt, y=CPP_est_filt)) + geom_point() + geom_abline(slope=1, intercept=0) + xlab("CPP without filtering criteria") + ylab("CPP calculated with filtering criteria") + theme_bw()
 
@@
 
+```{r fig.cap="Figure S3: Photoperiod response curves", echo=FALSE}
+DI_plot 
+```
+
+```{r fig.cap="Supplemental figure TBD.
+Diapause incidence (2018) by country.
+Dark sections of pie charts represent diapausing fraction; white fractions represent non-diapausing.", echo=FALSE, message=FALSE}
+DI_pies_2018
+```
+
 # References

### 2019-11-11 MS.Rmd
poster figures

@@
 
+```{r figs_for_poster, include=FALSE, echo=FALSE}
+ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, color=country, shape=year)) + 
+     geom_point(size=3) + scale_color_manual(values=c("red", "blue")) +
+     stat_smooth(method=lm, se=FALSE, aes(linetype=year)) + scale_linetype_manual(values=c("dotted","solid")) + theme_bw() + theme(legend.position=c(.8,.25), text=element_text(size=24)) + xlab("Latitude") + ylab("Critical photoperiod")
+
+remove_geom(us_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="US", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(text=element_text(size=24), axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "null"))
+
+remove_geom(japan_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="JP", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(text=element_text(size=24), axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "null"))
+
+```
+
 # References

### 2020-01-15 GDDdata_input.Rmd
GDD and temperature data

---
title: "GDD Deadlines Japan/US 1998-2018"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

Naresh and Leslie compiled some additional climate parameters: monthly mean maximum and minimum temperatures in the US and Japan locations, 1998-2008, as well as calculating values we are calling GDD400 and GDD465: the latest days of the year on which there are 400 (or 465) growing degree days left in the year. These "deadline" values were selected based on values for temperature-dependent development rates in the literature. The climate data come from the Climate Prediction Center (CPC) and have a resolution of about 50 sq km. There are no data for Okinawa in this dataset so these will have to come from elsewhere

```{r include=FALSE}
library(tidyverse)
```

```{r}
#importing and reshaping table
gdd400_us <- read.csv("GDD400_US.csv", check.names=FALSE) %>% gather(year, gdd400, "1998":"2018", factor_key=FALSE)
gdd400_us$year<-as.numeric(gdd400_us$year)
ggplot(gdd400_us, aes(x=year, y=gdd400, color=POP)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018))

```
```{r}
gdd400_jp <- read.csv("GDD400_JP.csv", check.names=FALSE) %>% gather(year, gdd400, "1998":"2018", factor_key=FALSE)
gdd400_jp$year<-as.numeric(gdd400_jp$year)
ggplot(gdd400_jp, aes(x=year, y=gdd400, color=POP)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018))

```
```{r}
#adding latitudes
locations<-read.csv("pop_locations_2018.csv")
gdd400_us<-left_join(gdd400_us, select(locations, pop, latitude, country), by=c("POP"="pop"))
gdd400_jp<-left_join(gdd400_jp, select(locations, pop, latitude, country), by=c("POP"="pop"))
gdd400_all<-rbind(gdd400_us, gdd400_jp)
gdd400_all$POP<-as.factor(gdd400_all$POP)
gdd400_all<-mutate(gdd400_all, interval=case_when(year<=2017 & year>=2008 ~ 'y2008-2017',
                                                  year<2008 ~ 'y1998-2007'))
```

```{r}
ggplot(filter(gdd400_all, year>2014), aes(x=latitude, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)
```

```{r}
ggplot(filter(gdd400_all, year<2008), aes(x=latitude, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)

```
```{r}
ggplot(filter(gdd400_all, year>=2008), aes(x=latitude, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)

```

```{r}
gdd400_all$interval<-as.factor(gdd400_all$interval)
```

```{r}
ggplot(filter(gdd400_all, country=="US", interval!="NA"), aes(x=interval, y=gdd400)) + geom_point(position=position_jitter()) + geom_boxplot(alpha=0) + facet_wrap(~POP)
```

```{r}
ggplot(filter(gdd400_all, country=="JP", interval!="NA"), aes(x=interval, y=gdd400)) + geom_point(position=position_jitter()) + geom_boxplot(alpha=0) + facet_wrap(~POP)

```
```{r}
```

```{r}
 ggplot(filter(gdd400_all, country=="US", interval!="NA"), aes(x=interval, y=gdd400)) + geom_point(position=position_jitter()) + geom_boxplot(alpha=0) + facet_wrap(~POP)

```
```{r}
tmin_us<-read.csv("AvgTmin_US.csv", check.names=FALSE) %>% gather(year, Tmin, "1998":"2018", factor_key=FALSE)
#In order to look at winter temperatures, need to combine Nov/Dec with Jan/Feb of the following year. January and February are reassigned in this table to the previous year to create seasonal years
tmin_us$year<-as.numeric(tmin_us$year) 
tmin_us$seasonalyear<-if_else(tmin_us$MONTH %in% c("Jan","Feb"), tmin_us$year-1, tmin_us$year)
minwintertemp_us<-filter(tmin_us, MONTH %in% c("Nov","Dec","Jan","Feb")) %>% group_by(POP, seasonalyear) %>% summarise(wintermin=min(Tmin)) %>% filter(seasonalyear!=1997)
ggplot(minwintertemp_us, aes(x=seasonalyear, y=wintermin, color=POP)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018)) 
```
```{r}
tmin_jp<-read.csv("AvgTmin_JP.csv", check.names=FALSE) %>% gather(year, Tmin, "1998":"2018", factor_key=FALSE)
tmin_jp$year<-as.numeric(tmin_jp$year) 
tmin_jp$seasonalyear<-if_else(tmin_jp$MONTH %in% c("Jan","Feb"), tmin_jp$year-1, tmin_jp$year)
minwintertemp_jp<-filter(tmin_jp, MONTH %in% c("Nov","Dec","Jan","Feb")) %>% group_by(POP, seasonalyear) %>% summarise(wintermin=min(Tmin)) %>% filter(seasonalyear!=1997)
ggplot(minwintertemp_jp, aes(x=seasonalyear, y=wintermin, color=POP)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018))
```

```{r}
wintermin_all<-rbind(minwintertemp_us, minwintertemp_jp)
#gdd400_us<-left_join(gdd400_us, select(locations, pop, latitude, country), by=c("POP"="pop"))
#gdd400_jp<-left_join(gdd400_jp, select(locations, pop, latitude, country), by=c("POP"="pop"))
#gdd400_all<-rbind(gdd400_us, gdd400_jp)
#gdd400_all$POP<-as.factor(gdd400_all$POP)
#gdd400_all<-mutate(gdd400_all, interval=case_when(year<=2017 & year>=2008 ~ 'y2008-2017',
#                                                  year<2008 ~ 'y1998-2007'))
gdd400_all_new<-left_join(gdd400_all, wintermin_all, by=c("POP"="POP", "year"="seasonalyear"))
gdd400_all_new$POP<-as.factor(gdd400_all_new$POP)
```

```{r}
ggplot(filter(gdd400_all_new, interval=="y1998-2007"), aes(x=latitude, y=wintermin, color=country)) + geom_point() + facet_wrap(~year) 
```

```{r}
ggplot(filter(gdd400_all_new, interval=="y2008-2017"), aes(x=latitude, y=wintermin, color=country)) + geom_point() + facet_wrap(~year)

```

```{r}
ggplot(filter(gdd400_all_new, interval=="y1998-2007"), aes(x=wintermin, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)
#Generally: at any given winter coldness, Japan has an earlier deadline than US
```
```{r}
ggplot(filter(gdd400_all_new, interval=="y2008-2017"), aes(x=wintermin, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)
#Generally: at any given winter coldness, Japan has an earlier deadline than US

```

### 2020-01-21 GDDdata_input.Rmd
Updated GDD/temp values

@@
 The climate data come from the Climate Prediction Center (CPC) and have a resolution of about 50 sq km.
-There are no data for Okinawa in this dataset so these will have to come from elsewhere
+The data for Okinawa come from...?
 
@@
 
+# GDD deadline data
+
 ```{r}
@@
 gdd400_us$year<-as.numeric(gdd400_us$year)
-ggplot(gdd400_us, aes(x=year, y=gdd400, color=POP)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018))
+ggplot(gdd400_us, aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018))
 
@@
 gdd400_jp$year<-as.numeric(gdd400_jp$year)
-ggplot(gdd400_jp, aes(x=year, y=gdd400, color=POP)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018))
+ggplot(gdd400_jp, aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018))
 
@@
 #adding latitudes
-locations<-read.csv("pop_locations_2018.csv")
-gdd400_us<-left_join(gdd400_us, select(locations, pop, latitude, country), by=c("POP"="pop"))
-gdd400_jp<-left_join(gdd400_jp, select(locations, pop, latitude, country), by=c("POP"="pop"))
+locations<-read.csv("mapping_pops_cpp_expt.csv")
+gdd400_us<-left_join(gdd400_us, select(locations, pop, latitude, country, only_2008), by=c("pop"))
+gdd400_jp<-left_join(gdd400_jp, select(locations, pop, latitude, country, only_2008), by=c("pop"))
 gdd400_all<-rbind(gdd400_us, gdd400_jp)
-gdd400_all$POP<-as.factor(gdd400_all$POP)
+gdd400_all$pop<-as.factor(gdd400_all$pop)
 gdd400_all<-mutate(gdd400_all, interval=case_when(year<=2017 & year>=2008 ~ 'y2008-2017',
@@
 
+## GDD deadline by latitude in Japan and US
+
+NB: In the US, the same sites were sampled in 2008 and 2018. In Japan, some sites were different between 2008 and 2018. The following graphs show only the 2018 sites over all years.
+
 ```{r}
-ggplot(filter(gdd400_all, year>2014), aes(x=latitude, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)
+ggplot(filter(gdd400_all, year>2014, only_2008=="n"), aes(x=latitude, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)
 ```
@@
 ```{r}
-ggplot(filter(gdd400_all, year<2008), aes(x=latitude, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)
+ggplot(filter(gdd400_all, year<2008, only_2008=="n"), aes(x=latitude, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)
 
@@
 ```{r}
-ggplot(filter(gdd400_all, year>=2008), aes(x=latitude, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)
+ggplot(filter(gdd400_all, year>=2008, only_2008=="n"), aes(x=latitude, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)
 
@@
 
+## GDD deadlines 1998-2007 vs. 2008-2017
+
 ```{r}
-ggplot(filter(gdd400_all, country=="US", interval!="NA"), aes(x=interval, y=gdd400)) + geom_point(position=position_jitter()) + geom_boxplot(alpha=0) + facet_wrap(~POP)
+ggplot(filter(gdd400_all, country=="US", interval!="NA"), aes(x=interval, y=gdd400)) + geom_point(position=position_jitter()) + geom_boxplot(alpha=0) + facet_wrap(~pop)
 ```
@@
 ```{r}
-ggplot(filter(gdd400_all, country=="JP", interval!="NA"), aes(x=interval, y=gdd400)) + geom_point(position=position_jitter()) + geom_boxplot(alpha=0) + facet_wrap(~POP)
+ggplot(filter(gdd400_all, country=="JP", interval!="NA", only_2008=="n"), aes(x=interval, y=gdd400)) + geom_point(position=position_jitter()) + geom_boxplot(alpha=0) + facet_wrap(~pop)
 
 ```
+
+## GDD deadlines including 2008 Japan sites
+
 ```{r}
+ggplot(filter(gdd400_all, year>2014), aes(x=latitude, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)
 ```
 
-```{r}
- ggplot(filter(gdd400_all, country=="US", interval!="NA"), aes(x=interval, y=gdd400)) + geom_point(position=position_jitter()) + geom_boxplot(alpha=0) + facet_wrap(~POP)
+Notable: the point at ~30 degrees lat in Japan has a later GDD deadline than the other sites in Japan.
+This population, TAN (Tanegashima, located on a small island), was also an outlier in the 2008 CPP experiment, with a CPP value in the range of the US values:
 
-```
+![figures/CPP_year_country.png](figures/CPP_year_country.png)
+![Urbanski_08cpp.png](figures/Urbanski_08cpp.png)
+
+# Minimum average Tmin of winter months
+
 ```{r}
@@
 tmin_us$year<-as.numeric(tmin_us$year) 
-tmin_us$seasonalyear<-if_else(tmin_us$MONTH %in% c("Jan","Feb"), tmin_us$year-1, tmin_us$year)
-minwintertemp_us<-filter(tmin_us, MONTH %in% c("Nov","Dec","Jan","Feb")) %>% group_by(POP, seasonalyear) %>% summarise(wintermin=min(Tmin)) %>% filter(seasonalyear!=1997)
-ggplot(minwintertemp_us, aes(x=seasonalyear, y=wintermin, color=POP)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018)) 
+tmin_us$seasonalyear<-if_else(tmin_us$month %in% c("Jan","Feb"), tmin_us$year-1, tmin_us$year)
+minwintertemp_us<-filter(tmin_us, month %in% c("Nov","Dec","Jan","Feb")) %>% group_by(pop, seasonalyear) %>% summarise(wintermin=min(Tmin)) %>% filter(!(seasonalyear %in% c(1997,2018)))
+ggplot(minwintertemp_us, aes(x=seasonalyear, y=wintermin, color=pop)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018)) 
 ```
@@
 tmin_jp$year<-as.numeric(tmin_jp$year) 
-tmin_jp$seasonalyear<-if_else(tmin_jp$MONTH %in% c("Jan","Feb"), tmin_jp$year-1, tmin_jp$year)
-minwintertemp_jp<-filter(tmin_jp, MONTH %in% c("Nov","Dec","Jan","Feb")) %>% group_by(POP, seasonalyear) %>% summarise(wintermin=min(Tmin)) %>% filter(seasonalyear!=1997)
-ggplot(minwintertemp_jp, aes(x=seasonalyear, y=wintermin, color=POP)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018))
+tmin_jp$seasonalyear<-if_else(tmin_jp$month %in% c("Jan","Feb"), tmin_jp$year-1, tmin_jp$year)
+minwintertemp_jp<-filter(tmin_jp, month %in% c("Nov","Dec","Jan","Feb")) %>% group_by(pop, seasonalyear) %>% summarise(wintermin=min(Tmin)) %>% filter(!(seasonalyear %in% c(1997,2018)))
+ggplot(minwintertemp_jp, aes(x=seasonalyear, y=wintermin, color=pop)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018))
 ```
@@
 wintermin_all<-rbind(minwintertemp_us, minwintertemp_jp)
-#gdd400_us<-left_join(gdd400_us, select(locations, pop, latitude, country), by=c("POP"="pop"))
-#gdd400_jp<-left_join(gdd400_jp, select(locations, pop, latitude, country), by=c("POP"="pop"))
-#gdd400_all<-rbind(gdd400_us, gdd400_jp)
-#gdd400_all$POP<-as.factor(gdd400_all$POP)
-#gdd400_all<-mutate(gdd400_all, interval=case_when(year<=2017 & year>=2008 ~ 'y2008-2017',
-#                                                  year<2008 ~ 'y1998-2007'))
-gdd400_all_new<-left_join(gdd400_all, wintermin_all, by=c("POP"="POP", "year"="seasonalyear"))
-gdd400_all_new$POP<-as.factor(gdd400_all_new$POP)
+gdd400_all_new<-left_join(gdd400_all, wintermin_all, by=c("pop"="pop", "year"="seasonalyear"))
+gdd400_all_new$pop<-as.factor(gdd400_all_new$pop)
+
+```
+## Winter coldness by latitude in US and Japan
+
+Winter coldness looks similar in US and Japan.
+Previous data showed that *absolute* minumum temperature is colder in US than Japan (see Urbanski 2012 appendix).
+Note: these graphs include only the Japanese sites sampled in 2018.
+Also note: the winter prior to summer 2018 (i.e.
+Nov and Dec 2017 + Jan and Feb 2018) is labeled as year 2017 for this value
+
+```{r}
+ggplot(filter(gdd400_all_new, interval=="y1998-2007", only_2008=="n"), aes(x=latitude, y=wintermin, color=country)) + geom_point() + facet_wrap(~year) 
 ```
@@
 ```{r}
-ggplot(filter(gdd400_all_new, interval=="y1998-2007"), aes(x=latitude, y=wintermin, color=country)) + geom_point() + facet_wrap(~year) 
+ggplot(filter(gdd400_all_new, interval=="y2008-2017", only_2008=="n"), aes(x=latitude, y=wintermin, color=country)) + geom_point() + facet_wrap(~year)
+
 ```
 
+## Winter coldness 1998-2007 vs. 2008-2017
+
 ```{r}
-ggplot(filter(gdd400_all_new, interval=="y2008-2017"), aes(x=latitude, y=wintermin, color=country)) + geom_point() + facet_wrap(~year)
+ggplot(filter(gdd400_all_new, country=="US", interval!="NA"), aes(x=interval, y=wintermin)) + geom_point(position=position_jitter()) + geom_boxplot(alpha=0) + facet_wrap(~pop)
+```
+
+## Winter 2007 vs. 2017
+These are the winters immediately preceding the summer populations were collected for the study. ALL Japanese sites are included here.
+
+```{r}
+ggplot(filter(gdd400_all_new, year %in% c(2007,2017)), aes(x=latitude, y=wintermin, color=year)) + geom_point() + facet_grid(~country)
+
+```
+## Winter 2006 vs. 2016
+All Japanese sites included
+
+```{r}
+ggplot(filter(gdd400_all_new, year %in% c(2006,2016)), aes(x=latitude, y=wintermin, color=year)) + geom_point() + facet_grid(~country)
+
+```
+## Winter 2005 vs. 2015
+
+```{r}
+ggplot(filter(gdd400_all_new, year %in% c(2005,2015)), aes(x=latitude, y=wintermin, color=year)) + geom_point() + facet_grid(~country)
 
@@
 ```{r}
-ggplot(filter(gdd400_all_new, interval=="y1998-2007"), aes(x=wintermin, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)
-#Generally: at any given winter coldness, Japan has an earlier deadline than US
+#adding 2008 and 2018 CPP values to big table of yearly deadlines and winter temps
+cpp_year<-select(read.csv("cpp_calcs_filtered.csv"), pop, Estimate, year) %>% filter(year!="1988") %>% rename("CPP_YEAR"="year")
+gdd400_all_new<-left_join(gdd400_all_new, cpp_year, by="pop")
+gdd400_all_new$CPP_YEAR<-as.factor(gdd400_all_new2$CPP_YEAR)
+
+```
+
+```{r}
+ggplot(filter(gdd400_all_new, year %in% c(2007,2017)), aes(x=gdd400, y=Estimate, color=CPP_YEAR)) + scale_color_manual(values=c("gray", "black")) + geom_point() + facet_wrap(~year) + stat_smooth(method=lm)
 ```
 ```{r}
-ggplot(filter(gdd400_all_new, interval=="y2008-2017"), aes(x=wintermin, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)
-#Generally: at any given winter coldness, Japan has an earlier deadline than US
-
+ggplot(filter(gdd400_all_new, year %in% c(2007,2017)), aes(x=wintermin, y=Estimate, color=CPP_YEAR)) + scale_color_manual(values=c("gray", "black")) + geom_point() + facet_wrap(~year) + stat_smooth(method=lm)
 ```

### 2020-01-21 MS.Rmd
Updated GDD/temp values

@@
   group_by(CageID) %>% 
-  summarise(pop=first(Population), pp=mean(Photoperiod), lat=mean(Lat), block=first(Block), country=first(Country), h1=sum(H1Count), h2=sum(H2Count), emb=sum(EmbryoCount)) %>%   filter(h1+h2+emb>100)
+  dplyr::summarise(pop=first(Population), pp=mean(Photoperiod), lat=mean(Lat), block=first(Block), country=first(Country), h1=sum(H1Count), h2=sum(H2Count), emb=sum(EmbryoCount)) %>%   filter(h1+h2+emb>100)
 DI_summary$total<-DI_summary$h1+DI_summary$h2+DI_summary$emb
@@
 popdata_2018<-DI_summary %>% group_by(pop) %>% 
-  summarize(lat=first(lat), country=first(country))
+  dplyr::summarize(lat=first(lat), country=first(country))
 cpp_2018<-right_join(cpp_2018, popdata_2018, by="pop")
@@
 popdata_2008<-DI_summary_2008 %>% group_by(pop) %>% 
-  summarize(country=first(country), lat=first(lat), year=first(year)) 
+  dplyr::summarize(country=first(country), lat=first(lat), year=first(year)) 
 cpp_2008<-right_join(cpp_2008, popdata_2008, by="pop")
@@
 DI_short_2018<-filter(DI_summary, pp==8) %>% group_by(pop) %>% 
-  summarize(country=first(country), lat=first(lat), photoperiod=first(pp), emb=sum(emb), total=sum(total))
+  dplyr::summarize(country=first(country), lat=first(lat), photoperiod=first(pp), emb=sum(emb), total=sum(total))
 DI_short_2018$DI<-DI_short_2018$emb/DI_short_2018$total
@@
 ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, color=country, shape=year)) + 
-     geom_point(size=3) + scale_color_manual(values=c("red", "blue")) +
+     geom_point(size=6) + scale_color_manual(values=c("red", "blue")) +
      stat_smooth(method=lm, se=FALSE, aes(linetype=year)) + scale_linetype_manual(values=c("dotted","solid")) + theme_bw() + theme(legend.position=c(.8,.25), text=element_text(size=24)) + xlab("Latitude") + ylab("Critical photoperiod")

### 2020-01-24 GDDdata_input.Rmd
Ready for MS

@@
 output:
+  pdf_document: default
+  html_notebook: default
   html_document:
     df_print: paged
-  html_notebook: default
-  pdf_document: default
 ---
@@
 ```{r}
-ggplot(filter(gdd400_all, year>2014), aes(x=latitude, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)
+ggplot(filter(gdd400_all, year %in% c(2005, 2006, 2007, 2015, 2016, 2017)), aes(x=latitude, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)
 ```
@@
 tmin_us$seasonalyear<-if_else(tmin_us$month %in% c("Jan","Feb"), tmin_us$year-1, tmin_us$year)
-minwintertemp_us<-filter(tmin_us, month %in% c("Nov","Dec","Jan","Feb")) %>% group_by(pop, seasonalyear) %>% summarise(wintermin=min(Tmin)) %>% filter(!(seasonalyear %in% c(1997,2018)))
+minwintertemp_us<-filter(tmin_us, month %in% c("Nov","Dec","Jan","Feb")) %>% group_by(pop, seasonalyear) %>% summarise(wintermin=min(Tmin), wintermonths_sd=sd(Tmin)) %>% filter(!(seasonalyear %in% c(1997,2018)))
 ggplot(minwintertemp_us, aes(x=seasonalyear, y=wintermin, color=pop)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018)) 
 ```
+```{r}
+ggplot(minwintertemp_us, aes(x=seasonalyear, y=wintermonths_sd, color=pop)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018))
+
+```
+
 ```{r}
@@
 tmin_jp$seasonalyear<-if_else(tmin_jp$month %in% c("Jan","Feb"), tmin_jp$year-1, tmin_jp$year)
-minwintertemp_jp<-filter(tmin_jp, month %in% c("Nov","Dec","Jan","Feb")) %>% group_by(pop, seasonalyear) %>% summarise(wintermin=min(Tmin)) %>% filter(!(seasonalyear %in% c(1997,2018)))
+minwintertemp_jp<-filter(tmin_jp, month %in% c("Nov","Dec","Jan","Feb")) %>% group_by(pop, seasonalyear) %>% summarise(wintermin=min(Tmin), wintermonths_sd=sd(Tmin)) %>% filter(!(seasonalyear %in% c(1997,2018)))
 ggplot(minwintertemp_jp, aes(x=seasonalyear, y=wintermin, color=pop)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018))
+```
+```{r}
+ggplot(minwintertemp_jp, aes(x=seasonalyear, y=wintermonths_sd, color=pop)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018))
+
 ```
@@
 ```{r}
-ggplot(filter(gdd400_all_new, year %in% c(2007,2017)), aes(x=latitude, y=wintermin, color=year)) + geom_point() + facet_grid(~country)
+ggplot(filter(gdd400_all_new, year %in% c(2007,2017), only_2008=="n"), aes(x=latitude, y=wintermin, color=as.factor(year))) + geom_point() + facet_grid(~country) + scale_color_manual(values=c("blue","orange"))
 
@@
 ```{r}
-ggplot(filter(gdd400_all_new, year %in% c(2006,2016)), aes(x=latitude, y=wintermin, color=year)) + geom_point() + facet_grid(~country)
+ggplot(filter(gdd400_all_new, year %in% c(2006,2016), only_2008=="n"), aes(x=latitude, y=wintermin, color=as.factor(year))) + geom_point() + facet_grid(~country) + scale_color_manual(values=c("blue","orange"))
 
@@
 ```{r}
-ggplot(filter(gdd400_all_new, year %in% c(2005,2015)), aes(x=latitude, y=wintermin, color=year)) + geom_point() + facet_grid(~country)
+ggplot(filter(gdd400_all_new, year %in% c(2005,2015)), aes(x=latitude, y=wintermin, color=as.factor(year))) + geom_point() + facet_grid(~country) + scale_color_manual(values=c("blue","orange"))
 
@@
 gdd400_all_new<-left_join(gdd400_all_new, cpp_year, by="pop")
-gdd400_all_new$CPP_YEAR<-as.factor(gdd400_all_new2$CPP_YEAR)
+gdd400_all_new$CPP_YEAR<-as.factor(gdd400_all_new$CPP_YEAR)
+
+```
+
+# 2008 and 2018 CPP estimates vs GDD deadline
+
+```{r}
+
+ggplot(filter(gdd400_all_new, year==2007, CPP_YEAR==2018), aes(x=gdd400, y=Estimate, color=country)) + geom_point() + stat_smooth(method=lm) 
+```
+```{r}
+ggplot(filter(gdd400_all_new, year==2007, CPP_YEAR==2018), aes(x=latitude, y=Estimate, color=country)) + geom_point() + stat_smooth(method=lm)
 
@@
 ```
+
+# Variability (SD) in winter coldness in 1998-2007 vs. 2008-2017
+
 ```{r}
-ggplot(filter(gdd400_all_new, year %in% c(2007,2017)), aes(x=wintermin, y=Estimate, color=CPP_YEAR)) + scale_color_manual(values=c("gray", "black")) + geom_point() + facet_wrap(~year) + stat_smooth(method=lm)
+variation_summary<-filter(gdd400_all_new, interval!="NA", CPP_YEAR=="2018") %>% group_by(pop, interval) %>% dplyr::summarize(lat=mean(latitude), country=first(country), gdd400_sd=sd(gdd400), wintermin_sd=sd(wintermin), cpp_2018=mean(Estimate))
+ggplot(variation_summary, aes(x=lat, y=wintermin_sd, color=interval)) + geom_point() + facet_grid(~country)
 ```

### 2020-01-24 MS.Rmd
Ready for MS

@@
 install.packages("cowplot")
-
 ```
@@
 DI_summary$pop<-factor(DI_summary$pop, levels=c("PBS", "FEM", "OAK", "JAC", "BRU", "ZIO", "FAY", "NVA", "WAV", "MAN", "BER", "NEW", "OKI", "TAN", "KAG", "YAT", "SAG", "SHI", "YAM", "HIR", "KYO", "TAK", "TOK", "UTS", "KAN", "KHO", "AIZ", "NIG", "SEN", "SAK"))
-DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~pop) + theme_bw() + xlab("Photoperiod (hours)") + ylab("Dipause incidence")
+DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~fct_reorder(pop, lat)) + theme_bw() + xlab("Photoperiod (hours)") + ylab("Dipause incidence")
 
@@
 cpp_values<-data.frame(ED(d.model.fit, 50))
-```
-
-```{r}
-
 ```
@@
 
+```{r}
+```
+
 # References

### 2020-01-28 MS.Rmd
home

--- 2019-11-11 MS.Rmd
@@
 hist(DI_summary$total, main="Total number of viable eggs per cage")
+
 ```
@@
 #plot with DI at 8 h shown as TWO DIFFERENT points (corresponding to two cabinets)
-c("PBS", "FEM", "OAK", "JAC", "BRU", "ZIO", "FAY", "NVA", "WAV", "MAN", "BER", "NEW", "OKI", "TAN", "KAG", "YAT", "SAG", "SHI", "YAM", "HIR", "KYO", "TAK", "TOK", "UTS", "KAN", "KHO", "AIZ", "NIG", "SEN", "SAK")
-DI_summary$pop<-factor(DI_summary$pop, levels=c("PBS", "FEM", "OAK", "JAC", "BRU", "ZIO", "FAY", "NVA", "WAV", "MAN", "BER", "NEW", "OKI", "TAN", "KAG", "YAT", "SAG", "SHI", "YAM", "HIR", "KYO", "TAK", "TOK", "UTS", "KAN", "KHO", "AIZ", "NIG", "SEN", "SAK"))
-DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~pop) + theme_bw() + xlab("Photoperiod (hours)") + ylab("Dipause incidence")
+DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + stat_smooth(method="drm", se=FALSE) + facet_wrap(~pop)
 
@@
 cppmodel<-lm(Estimate~lat*country*year, data=filter(cpp_2, year!=1988))
-cppmodel_result<-data.frame(anova(cppmodel))
+anova(cppmodel)
 ```
@@
 cpp_08_18_shifts<-ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, color=country, shape=year)) + 
-  geom_point(size=3) + scale_color_manual(values=c("red", "blue")) +
-  stat_smooth(method=lm, se=FALSE, aes(linetype=year)) + scale_linetype_manual(values=c("dotted","solid")) + theme_bw() + theme(legend.position=c(.8,.25)) + xlab("Latitude") + ylab("Critical photoperiod")
-  # + geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) 
+  geom_point() + scale_color_manual(values=c("red", "blue")) +
+  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
+  stat_smooth(method=lm, se=FALSE, aes(linetype=year)) + theme_bw() + theme(legend.position=c(.8,.25))
 ```
@@
 dimodel<-lm(DI~lat*country*year, data=DI_short_2)
-dimodel_result<-data.frame(anova(dimodel))
+anova(dimodel)
 ```
@@
 ### Mosquito population sampling and pre-experiment colony maintenance
-Mosquito populations were sampled as larvae and pupae in the eastern USA and Japan in June-September 2018, at latitudes ranging from `r min(pops$latitude)` to `r max(pops$latitude)`.
+Mosquito populations were sampled as larvae and pupae in the eastern USA and Japan in 2018, at latitudes ranging from `r min(pops$latitude)` to `r max(pops$latitude)`.
 Sampling locations are described in Table 1.
-Locations in the US were identical to locations sampled in 2008.
-Japanese population samples were confirmed to be Ae.
-albopictus via morphological identification following [@tanakaRevisionAdultLarval1979] and sent to Georgetown as F1 eggs (except for TAK, for which F2 eggs were provided).
-At least 50 male and 50 female adult mosquitoes were used to initiate all populations in the laboratory at Georgetown, except for SEN, which had 37 females and 39 males.
+Locations in the US were identical to locations sampled in 2008, except: .
+At least 50 male and 50 female mosquitoes were used to initiate laboratory populations, except for SEN, which had 37 females and 39 males initially.
 
 ```{r table1, echo=FALSE}
-kable(filter(pops, only_2008=="n")[,1:6], col.names=c("Population", "City", "Latitude", "Longitude", "Country", "Life Stage Sampled"), caption="Table 1: 2018 Sampling Locations")
+kable(filter(pops, only_2008=="n")[,1:5], col.names=c("Population", "City", "Latitude", "Longitude", "Country"), caption="Table 1: 2018 Sampling Locations")
 ```
 
-Populations were maintained in the laboratory for 3 or 4 generations prior to the experiment at 21C and 80% relative humidity under a 16:8h light:dark cycle, following previous studies [@armbrusterGeographicVariationLarval2006; @urbanskiRapidAdaptiveEvolution2012].
+Populations were maintained in the laboratory at Georgetown University for 3 or 4 generations prior to the experiment at 21C and 80% relative humidity under a 16:8h light:dark cycle, following previous studies [@armbrusterGeographicVariationLarval2006; @urbanskiRapidAdaptiveEvolution2012].
 Larvae were reared in distilled water and fed a slurry of dog food and brine shrimp.
 Adults were provided organic raisins as a carbohydrate source and blood meals from a human host.
-Initial laboratory generations of larvae were reared in plastic storage containers containing ~2.5 L of distilled water, with one population per container at initial densities of ~200-300, for ease of bulk rearing.
+Initial laboratory generations of larvae were reared in plastic storage containers containing ~2.5 L of distilled water, with one population per container at initial densities of ~200-300.
 Pupae were then transferred into one adult cage per population.
-Populations generally had at least 50 female and 50 male adults per generation, but the JAC population experienced visibly high mortality during the F2 generation.
+Populations generally had at least 50 female and 50 male adults per generation, though the JAC population experienced visibly high mortality during the F2 generation.
 The larval generation immediately prior to the experiment was reared in 15 cm plastic petri dishes containing 30-40 larvae each, 10 dishes per population, to match the larval rearing conditions of the experimental generation, which followed the methods of Urbanski et al 2012.
-Additional details of mosquito husbandry prior to the experiment can be found in Supplementary File 1.
+Additional details of mosquito husbandry can be found in Supplementary File 1.
 
 ### Critical photoperiod determination
-An overview of the procedure used to determine the critical photoperiod (CPP) is as follows: Starting from the pupal stage, mosquitoes were divided among 12 different photoperiods (listed in Supplementary File 1) and allowed to produce eggs.
+An overview of the procedure used to determine the critical photoperiod is as follows: Starting from the pupal stage, mosquitoes were divided among 12 different photoperiods (listed in Supplementary File 1) and allowed to produce eggs.
 The proportion of diapause eggs at each photoperiod was determined, and a dose-response curve was constructed with respect to hours of light and proportion of eggs in diapause.
@@
 the fraction that did not enter diapause under unambigous short-day conditions, 8 hours of light).
-This experimental design and most of the equipment and experimental details were the same as those used in Urbanski et al 2012; additional details on all steps can be found in Supplementary File 1.
+The experimental design and most of the equipment and experimental details were the same as those used in Urbanski et al 2012; additional details on all steps can be found in Supplementary File 1.
 
@@
 Each block contained half of the populations from each country; populations across the full range of latitudes from each country were interspersed between the blocks.
-At the beginning of each block, eggs were hatched by submerging oviposition papers in distilled water with larval food for two days.
+At the beginning of each block, eggs were hatched by submerging oviposition papers in water with larval food for two days.
 Larvae were distributed among 40 15 cm petri dishes containing 90 ml of a mixture of distilled water and larval food (5 L of water to 10 mL of food slurry) at a density of 30-40 larvae per dish.
-Larvae were transferred to clean dishes containing fresh water/food mixture every Monday-Wednesday-Friday (MWF) until pupation.
+Larvae were transferred to clean dishes containing fresh water/food mixture three times a week (MWF) until pupation.
 Larvae were kept in the controlled temperature room at 21C, 80% humidity and 16:8 hours light:dark until they produced pupae.
 
-Pupae from each population were collected, pooled across dishes, and split amongst thirteen 1.5 L adult cages, corresponding to twelve photoperiod cabinets and an additional short-day (8h light) incubator to provide additional data on unambiguous short-day diapause incidence.
+Pupae from each population were collected, pooled across dishes, and split amongst 13 1.5 L adult cages, corresponding to 12 photoperiod cabinets and an additional short-day (8h light) incubator to provide additional data on unambiguous short-day diapause incidence.
 The total number of pupae added to each cage ranged from 52 to 106; pupae were not sexed and subsequent pupal and adult mortality was not monitored.
-Approximately one week after the last pupae were added, cages were offered a blood meal, and a dark oviposition cup half-filled with water and lined with unbleached paper towel was added to each cage.
-Oviposition papers were collected 5-7 days after the blood meal and replaced every MWF; a total of four oviposition papers were collected from each cage.
-Several cages were fed again and an additional oviposition paper was collected; see supplemental methods for details.
+Approximately one week after the last pupae were added, cages were provided a blood meal, and a dark oviposition cup half-filled with water and lined with unbleached paper towel was added to each cage.
+Oviposition papers were collected 5-7 days after blood-feeding and replaced every MWF; a total of four oviposition papers were collected from each cage.
+Several cages were provided a second blood meal and an additional oviposition paper was collected; see Supplementary File 1 for details.
 Oviposition papers were dried gently and stored in the photoperiod cabinets from which they originated for 11 to 21 days until diapause determination.
 
-On each oviposition paper, eggs were counted (subsampling if oviposition was very high; see supplemental methods) and papers were submerged in distilled water with ~0.25 ml larval food.
+Oviposition papers with substantial numbers of eggs were trimmed to a ~2 cm strip surrounding what was the waterline in the oviposition cup; this zone generally contained the vast majority of the eggs.
+On papers with very dense oviposition, the egg-containing strip was haphazardly subsampled, to several hundred eggs as approximated by eye.
+Eggs were counted and egg papers were submerged in distilled water with several drops of larval food.
 Larvae that hatched were counted two days later and papers were re-dried and stored as before.
 After 1-2 weeks, hatch was induced again and larvae were counted.
-Oviposition papers were then treated overnight with a bleach solution to clear the chorion of the remaining eggs [@trpisNewBleachingDecalcifying1970] .
-Eggs were examined under a stereomicroscope, and those eggs containing a pharate larva were scored as embryonated and in diapause.
+Oviposition papers were then treated overnight with a bleach solution to clear the chorion of the remaining eggs [@trpisNewBleachingDecalcifying1970].
+Eggs containing a pharate larva were scored as embryonated and in diapause.
 The total number of hatched larvae and embryonated eggs was considered the number of viable eggs.
 Oviposition papers in which viable eggs made up less than 60% or more than 100% of the eggs initially counted on the paper were excluded from further analysis.
+Viable eggs were summed across oviposition papers and cages that had less than a total of 100 viable eggs were excluded.
 Analysis dates, subsampling and other detailed information about each oviposition paper can be found in Data Table 1.
 
-For each cage, the counts of larvae and embryos across all oviposition papers were pooled to calculate the diapause incidence, or the fraction of total viable eggs that are in diapause (# embryonated / (# embryonated + # hatched)).
+For each cage, the counts of larvae and embryos across all oviposition papers were pooled to calculate the diapause incidence: number of embryonated eggs divided by total number of viable eggs (# embryonated / (# embryonated + # hatched)).
 The median total number of viable eggs per cage was `r median(DI_summary$total)` (range `r min(DI_summary$total)`-`r max(DI_summary$total)`).
 For the 8-hour photoperiod, each population had data from two cages: one in a photoperiod cabinet and one in a separate incubator.
-These data were pooled to calculate the value for diapause incidence at unambiguous short-day (DI) for the population, but used as separate data points in photoperiod response curves.
+These data were pooled to calculate the diapause incidence value for the population, but used as separate data points in photoperiod response curves.
 
@@
 2018) on CPP and on DI were analyzed with linear models.
-Data manipulation and figure construction were carried out using packages in tidyverse (`r packageVersion("tidyverse")`).
+Additional packages used for calculations, analysis and figures included tidyverse (`r packageVersion("tidyverse")`), maps (`r packageVersion("maps")`), mapdata (`r packageVersion("mapdata")`), scatterpie (`r packageVersion("scatterpie")`)…
 
@@
 
-Latitude-specific DI (the fraction of diapausing eggs produced under unambiguous short-day conditions) did not significantly differ between 2008 and 2018 in Japan or the US (p=0.94, Figure 2, Table 2).
-DI was high (generally greater than 90%) at latitudes north of ~32.5 degrees in both countries.
-Substantial fractions of constitutively non-diapausing eggs were observed in populations south of South Carolina in the US and in Okinawa in Japan (latitude effect p=<0.001).
-Latitude-specific DI was lower in the US than in Japan (country effect p=0.026), though there were fewer sites sampled at lower latitudes in Japan than the US due to lack of land between Okinawa and Kagashima.
+Latitude-specific diapause incidence (the fraction of diapausing eggs produced under unambiguous short-day conditions) did not significantly differ between 2008 and 2018 in Japan or the US (Figure 2, Table 2).
+Diapause incidence was high (generally greater than 90%) at latitudes north of ~32.5 degrees in both countries.
+Substantial fractions of constitutively non-diapausing eggs were observed in populations south of South Carolina in the US and in Okinawa in Japan.
+Diapause incidence followed a latitudinal cline, more pronounced in the US than in Japan because of the limited land to sample in Japan's southern latitudes.
+In the southern populations that could be compared between the US and Japan, latitude-specific diapause incidence was lower in the US.
 
-```{r fig.cap="Figure 1: Diapause incidence in Japan and US in 2008 and 2018", echo=FALSE, include=FALSE}
+```{r fig1, fig.pos="H", fig.cap="Figure 1: Diapause incidence in Japan and US in 2008 and 2018", echo=FALSE, include=FALSE}
 incidence_lat_year_plot
@@
 
-```{r fig.cap="Figure 2B: Diapause incidence by country and year.
-Lines represent locally estimated scatterplot smoothing (LOESS).
-Solid lines, 2018 data; dotted lines, 2008 data.", echo=FALSE, message=FALSE}
+```{r table2, echo=FALSE}
+kable(prettify(anova(dimodel)), digits=3, caption="Table 2: Effects of latitude, country and year (2008 vs 2018) on diapause incidence.")
+```
+
+```{r fig.cap="Figure 2B: Diapause incidence by country and year.", echo=FALSE, message=FALSE}
 #country_DI
-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year), se=FALSE) + scale_linetype_manual(values=c("dotted","solid")) + xlab("Latitude") + ylab("Short-day diapause incidence")
+ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point() + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year), se=FALSE)
 
@@
 
-```{r echo=FALSE}
-kable(prettify(anova(dimodel)), digits=3, caption="Table 2: Effects of latitude, country and year (2008 vs 2018) on diapause incidence.")
-```
+CPP values correlated strongly with latitude in the US and Japan.
+Latitudinal clines were nearly parallel to those observed in 2008 (Figure 3), as verified by the lack of a significant interaction term between year and latitude (Table 3).
+However, the intercepts of the latitudinal clines significantly differed between both the US and Japan, and between 2008 and 2018.
+Latitude-specific CPP was higher in Japan than the US, and higher in 2018 compared to 2008.
+This implies earlier entry into diapause at the same latitude in Japan than in the US and earlier entry into diapause in the same locations in 2018 than in 2008.
 
-CPP values correlated strongly with latitude in the US and Japan (p<0.001, Table 3).
-Latitudinal clines were nearly parallel to those observed in 2008 (Figure 3), as verified by the lack of a significant interaction term between year and latitude (p=0.17).
-However, the intercepts of the latitudinal clines significantly differed between both the US and Japan, and between 2008 and 2018 (p<0.001 for both country and year).
-Latitude-specific CPP was higher in Japan than the US, and higher in 2018 compared to 2008.
-This implies earlier entry into diapause at the same latitude in Japan than in the US, as well as earlier entry into diapause in the same locations in 2018 than in 2008.
-
-```{r fig.cap="Figure 3: Critical photoperiod (CPP) as a function of latitude in the US and Japan in 1988, 2008, and 2018.
-Error bars represent standard errors of CPP estimates.
-Errors around 1988 CPP estimates are absent because experiment and calculations were conducted using different methods and raw data were not available.", echo=FALSE, message=FALSE, include=FALSE}
+```{r fig3, fig.cap="Figure 3: Critical photoperiod (CPP) as a function of latitude in the US and Japan in 1988, 2008, and 2018.
+Error bars represent standard errors of CPP estimates.", echo=FALSE, message=FALSE, include=FALSE}
 cpp_japan_us_years
-```
-
-```{r fig.cap="Figure 4: Shifts in critical photoperiod latitudinal clines. Lines represent linear regressions. Solid lines, 2018; dotted lines, 2008.", echo=FALSE}
-cpp_08_18_shifts
 ```
@@
 
-## Supplementary File 1
-
-## Supplementary methods
-
-### Larval rearing
-Larvae were fed a slurry of 40 g frozen brine shrimp (San Francisco Bay Brand) and 120 g dog food (Nutro Feed Clean Ultra Small Breed Puppy formula) homogenized in 1 L of distilled water.
-Larvae were initially raised in plastic storage containers (Sterilite) containing 2.5 L of distilled water.
-Every Monday-Wednesday-Friday (MWF), the water in each container was removed by straining through a net and replaced with fresh distilled water; each container was fed approximately 1.5 ml of larval food slurry.
-Containers were covered with a fine mesh and kept at 21C, 80% humidity, and 16:8 L:D photoperiod.
-Starting with the generation prior to the one used in the experiment, larvae were reared in 15 cm petri dishes (Kord-Valmark), as in Urbanski et al 2012.
-The pre-experiment generation was fed ~0.5-0.6 ml larval food per dish, while the experimental generation had food/water concentrations as described in the main manuscript.
-Every MWF, larvae were transferred to clean petri dishes with a pipette and had fresh food/water added.
-When removing pupae substantially reduced the density of larvae, the remaining larvae were consolidated into fewer dishes.
-
-### Photoperiod cabinets 
-The custom-built photoperiod cabinets were the same ones used in Urbanski et al 2012, described in detail in the Appendix (https://www.journals.uchicago.edu/doi/suppl/10.1086/664709/suppl_file/53244apa.pdf).
-The light-dark cycles used were 8L : 16D, 12L : 12D, 12.5L : 11.5D, 12.75L : 11.25D, 13L : 11D, 13.25L : 10.75D, 13.5L : 10.5D, 13.75L : 10.25D, 14L : 10D, 14.25L : 9.75D, 14.5L : 9.5D, and 16L : 8D.
-Light and temperature inside cabinets were recorded for at least three full days for each cabinet in each block.
-The highest recorded temperature in any cabinet at any point was 22.65C, the lowest was 20.72C, and 70% of temperature readings were between 21C and 22C.
-All light:dark cycles were within 1 minute of desired photoperiod except for the separate short-day incubator (Cabinet 13 in Data Table 1), which was within 5 minutes of 8 hours of light (7:59 -8:05).
-Cages in photoperiod cabinets were monitored daily for desiccation and mold; water was added to cages and raisins were replaced as needed.
-
-### Blood meals, egg collection and storage
-Adult cages were fed from a human host over a three-day period; feeding days were randomly assigned by photoperiod cabinet.
-Collected oviposition papers were stored in 10 cm petri dishes inside humidified containers in the photoperiod cabinet in which they originated.
-In the second block of the experiment, two cages produced less than 200 eggs after the initial blood meal, and the cages in the cabinets with 14.5 hours and 14.25 hours of light had large numbers of larvae that hatched on the wet oviposition paper on the first egg collection day due to incomplete drying and thus could not be accurately counted.
-These 24 cages were offered a second blood meal and an additional oviposition paper was collected a week later.
-Papers with eggs produced after a second blood meal have a Paper ID ending in “-5” in Data Table 1.
-
-### Diapause determination
-Oviposition papers were trimmed to a ~2 cm strip surrounding the former location of the waterline in the oviposition cup; this zone generally contained the vast majority of the eggs.
-On oviposition papers with few eggs, the paper was trimmed in such a way as to capture all of them.
-On papers with very dense oviposition, the egg-containing strip was haphazardly subsampled further, to several hundred eggs as approximated by eye.
-
-After eggs were bleached, presence of pharate larvae was determined using the diagnostic criteria of visible eye spots and abdominal segmentation.
-On some oviposition papers, larvae “hatched” out of the eggs after the bleach solution was added; these were scored as embryonated diapause eggs because they appeared to be morphologically normal and did not hatch when induced with standard hatching cues twice.
-
-The total number of hatched larvae and embryos could differ from the number of eggs initially counted on the paper for the following reasons: 1) presence of inviable eggs or larvae (i.e., eggs that fail to hatch due to developmental abnormalities or larvae that die before being counted), 2) loss of eggs while handling oviposition papers, or 3) inaccurate counting at any step of the scoring process.
-Based on the distribution of percent differences from the initial counts (Figure S1), we considered oviposition papers with over 100% or less than 60% of the initial count to be less reliable data points.
-Excluding these data points resulted in 5 cages out of 286 in the total data set having less than 100 remaining eggs, so they were excluded from the photoperiod response curves.
-These filtering steps and exclusions did not affect CPP estimates for any population except PBS; for this population, excluding the unreliable counts eliminated an outlier in the photoperiod response curve and brought the CPP estimate into the range of CPP values for the rest of the populations (Figure S2).
-The final photoperiod response curves are shown in Figure S3.
-
-```{r fig.cap="Figure S1: Discrepancies between initial egg counts and final calculated egg counts on oviposition papers.", echo=FALSE}
-plot(blockAB$yield ~ blockAB$TotalLarvaeEmbryos, xlab="Number of viable eggs on oviposition paper", ylab="Viable eggs/initial egg count on paper")
+```{r fig4, fig.cap="Figure 4: Shifts in critical photoperiod latitudinal clines. Lines represent linear regressions; dotted=2018, solid=2008.", echo=FALSE}
+cpp_08_18_shifts
 ```
 
-```{r fig.cap="Figure S2.
-Critical photoperiod calculated with and without filtering oviposition papers based on criteria described.
-Line represents x=y.
-Outlier point corresponds to population PBS.", echo=FALSE}
-#comparing CPP values calculated with and without filtering out unreliable oviposition papers
-cpp_filtered<-read.csv("cpp_calcs_filtered.csv") %>% filter(year==2018)
-cpp_unfiltered<-read.csv("cpp_calcs_unfiltered.csv") %>% filter(year==2018)
-names(cpp_filtered)[5]<-"CPP_est_filt"
-names(cpp_unfiltered)[5]<-"CPP_est_unfilt"
-cpp_compare<-left_join(cpp_filtered,cpp_unfiltered,by="pop")
-ggplot(cpp_compare, aes(x=CPP_est_unfilt, y=CPP_est_filt)) + geom_point() + geom_abline(slope=1, intercept=0) + xlab("CPP without filtering criteria") + ylab("CPP calculated with filtering criteria") + theme_bw()
+## Supplementary Figures
 
-```
-
-```{r fig.cap="Figure S3: Photoperiod response curves", echo=FALSE}
+```{r figS1, fig.cap="Figure S1: Photoperiod response curves", echo=FALSE}
 DI_plot 
 ```
-
-```{r fig.cap="Supplemental figure TBD.
-Diapause incidence (2018) by country.
-Dark sections of pie charts represent diapausing fraction; white fractions represent non-diapausing.", echo=FALSE, message=FALSE}
+```{r figS2, fig.cap="Figure 2A: Diapause incidence (2018) by country. Dark sections of pie charts represent diapausing fraction; white fractions represent non-diapausing.", echo=FALSE, message=FALSE}
 DI_pies_2018
 ```
-
-```{r figs_for_poster, include=FALSE, echo=FALSE}
-ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, color=country, shape=year)) + 
-     geom_point(size=3) + scale_color_manual(values=c("red", "blue")) +
-     stat_smooth(method=lm, se=FALSE, aes(linetype=year)) + scale_linetype_manual(values=c("dotted","solid")) + theme_bw() + theme(legend.position=c(.8,.25), text=element_text(size=24)) + xlab("Latitude") + ylab("Critical photoperiod")
-
-remove_geom(us_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="US", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(text=element_text(size=24), axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "null"))
-
-remove_geom(japan_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="JP", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(text=element_text(size=24), axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "null"))
-
-```
-
 # References

### 2020-01-28 MS.Rmd
update

--- 2020-01-24 MS.Rmd
@@
+Skip to content
+Pull requests
+Issues
+Marketplace
+Explore
+@amusheg
+amusheg /
+albopictus-cpp
+
+1
+0
+
+    0
+
+Code
+Issues 0
+Pull requests 0
+Actions
+Projects 0
+Wiki
+Security
+Insights
+Settings
+albopictus-cpp/MS.Rmd
+Alexandra Mushegian Ready for MS b6284d3 4 days ago
+386 lines (303 sloc) 30.8 KB
 ---
@@
   geom_point() + facet_grid(~country) 
-
 country_DI_2018<-ggplot(DI_short_2018, aes(x=lat, y=DI, color=country)) + scale_color_manual(values=c("red","blue")) + geom_point() + theme_bw() + stat_smooth() + theme(legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "cm"))
-
 country_DI<-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country)) + scale_color_manual(values=c("red","blue")) + geom_point() + theme_bw() + stat_smooth() + theme(legend.position=c(0.9,0.1)) + facet_grid(~year)
-
 ```
@@
 pops<-data.frame(read.csv("mapping_pops_cpp_expt.csv"))
-
 us_map<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="blue") + 
@@
   theme_bw() 
-
 japan_map<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="red") + 
@@
   theme_bw() + theme(axis.title.x=element_blank(), axis.title.y=element_blank())
-
 DI_data<-data.frame(read.csv("diapause_incidence_8h_2008_2018.csv"))
@@
 DI_info<-rename(DI_info, Diapause=DI, Nondiapause=ND)
-
 us_DI_pies<-remove_geom(us_map, "label") + geom_scatterpie(data=filter(DI_info, country=="US"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + facet_grid(~year) + theme(axis.title.x=element_blank(), axis.title.y=element_blank())
-
 jp_DI_pies<-remove_geom(japan_map, "label") + geom_scatterpie(data=filter(DI_info, country=="JP"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + 
   facet_grid(~year) + theme(axis.title.x=element_blank(), axis.title.y=element_blank())
-
 DI_pies_2018<-plot_grid(remove_geom(us_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="US", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "null")),
@@
 ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year), se=FALSE) + scale_linetype_manual(values=c("dotted","solid")) + xlab("Latitude") + ylab("Short-day diapause incidence")
-
 ```
@@
 ggplot(cpp_compare, aes(x=CPP_est_unfilt, y=CPP_est_filt)) + geom_point() + geom_abline(slope=1, intercept=0) + xlab("CPP without filtering criteria") + ylab("CPP calculated with filtering criteria") + theme_bw()
-
 ```
@@
      stat_smooth(method=lm, se=FALSE, aes(linetype=year)) + scale_linetype_manual(values=c("dotted","solid")) + theme_bw() + theme(legend.position=c(.8,.25), text=element_text(size=24)) + xlab("Latitude") + ylab("Critical photoperiod")
-
 remove_geom(us_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="US", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(text=element_text(size=24), axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "null"))
-
 remove_geom(japan_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="JP", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(text=element_text(size=24), axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "null"))
-
-```
-
-```{r}
 ```

### 2020-01-29 MS.Rmd
bw figures

@@
-Skip to content
-Pull requests
-Issues
-Marketplace
-Explore
-@amusheg
-amusheg /
-albopictus-cpp
-
-1
-0
-
-    0
-
-Code
-Issues 0
-Pull requests 0
-Actions
-Projects 0
-Wiki
-Security
-Insights
-Settings
-albopictus-cpp/MS.Rmd
-Alexandra Mushegian Ready for MS b6284d3 4 days ago
-386 lines (303 sloc) 30.8 KB
 ---
@@
 ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year), se=FALSE) + scale_linetype_manual(values=c("dotted","solid")) + xlab("Latitude") + ylab("Short-day diapause incidence")
+
+ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("black","gray")) + 
+scale_shape_manual(values=c(21,24)) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year, color=country), se=FALSE) + scale_color_manual(values=c("black","gray")) + scale_linetype_manual(values=c("dotted","solid")) + xlab("Latitude") + ylab("Short-day diapause incidence")
+
 ```

### 2020-01-29 GDDdata_input.Rmd
bw figures

@@
 library(tidyverse)
+library(drc)
+library(cowplot)
 ```
@@
 locations<-read.csv("mapping_pops_cpp_expt.csv")
-gdd400_us<-left_join(gdd400_us, select(locations, pop, latitude, country, only_2008), by=c("pop"))
-gdd400_jp<-left_join(gdd400_jp, select(locations, pop, latitude, country, only_2008), by=c("pop"))
+gdd400_us<-left_join(gdd400_us, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
+gdd400_jp<-left_join(gdd400_jp, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
 gdd400_all<-rbind(gdd400_us, gdd400_jp)
@@
 ggplot(filter(gdd400_all, year>=2008, only_2008=="n"), aes(x=latitude, y=gdd400, color=country)) + geom_point() + facet_wrap(~year)
+
+```
+```{r}
+ggplot(filter(gdd400_all, only_2008=="n", year!=2018), aes(x=latitude, y=gdd400, color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") + geom_point() + facet_wrap(~year) + xlab("Latitude") + ylab("GDD Deadline (ordinal day)") + theme_bw()
 
@@
 tmin_us$seasonalyear<-if_else(tmin_us$month %in% c("Jan","Feb"), tmin_us$year-1, tmin_us$year)
-minwintertemp_us<-filter(tmin_us, month %in% c("Nov","Dec","Jan","Feb")) %>% group_by(pop, seasonalyear) %>% summarise(wintermin=min(Tmin), wintermonths_sd=sd(Tmin)) %>% filter(!(seasonalyear %in% c(1997,2018)))
+minwintertemp_us<-filter(tmin_us, month %in% c("Nov","Dec","Jan","Feb")) %>% group_by(pop, seasonalyear) %>% dplyr::summarise(wintermin=min(Tmin), wintermonths_sd=sd(Tmin)) %>% filter(!(seasonalyear %in% c(1997,2018)))
 ggplot(minwintertemp_us, aes(x=seasonalyear, y=wintermin, color=pop)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018)) 
@@
 tmin_jp$seasonalyear<-if_else(tmin_jp$month %in% c("Jan","Feb"), tmin_jp$year-1, tmin_jp$year)
-minwintertemp_jp<-filter(tmin_jp, month %in% c("Nov","Dec","Jan","Feb")) %>% group_by(pop, seasonalyear) %>% summarise(wintermin=min(Tmin), wintermonths_sd=sd(Tmin)) %>% filter(!(seasonalyear %in% c(1997,2018)))
+minwintertemp_jp<-filter(tmin_jp, month %in% c("Nov","Dec","Jan","Feb")) %>% group_by(pop, seasonalyear) %>% dplyr::summarise(wintermin=min(Tmin), wintermonths_sd=sd(Tmin)) %>% filter(!(seasonalyear %in% c(1997,2018)))
 ggplot(minwintertemp_jp, aes(x=seasonalyear, y=wintermin, color=pop)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018))
@@
 #adding 2008 and 2018 CPP values to big table of yearly deadlines and winter temps
-cpp_year<-select(read.csv("cpp_calcs_filtered.csv"), pop, Estimate, year) %>% filter(year!="1988") %>% rename("CPP_YEAR"="year")
+cpp_year<-dplyr::select(read.csv("cpp_calcs_filtered.csv"), pop, Estimate, year) %>% filter(year!="1988") %>% rename("CPP_YEAR"="year")
 gdd400_all_new<-left_join(gdd400_all_new, cpp_year, by="pop")
@@
 
-ggplot(filter(gdd400_all_new, year==2007, CPP_YEAR==2018), aes(x=gdd400, y=Estimate, color=country)) + geom_point() + stat_smooth(method=lm) 
+gdd_08 <- ggplot(filter(gdd400_all_new, year==2007, CPP_YEAR==2008), aes(x=gdd400, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=3, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
+  theme_bw() + theme(legend.position="none", text=element_text(size=14)) +
+  xlab("GDD Deadline (ordinal day), 2007") + ylab("Critical photoperiod (h), 2008")
+
+gdd_18 <- ggplot(filter(gdd400_all_new, year==2017, CPP_YEAR==2018), aes(x=gdd400, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=3, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
+theme_bw() + theme(legend.position=c(.8,.8), text=element_text(size=14)) +
+xlab("GDD Deadline (ordinal day), 2017") + ylab("Critical photoperiod (h), 2018")
+
+plot_grid(gdd_08, gdd_18, labels="AUTO")
+
 ```
+
 ```{r}
@@
 
-```{r}
+```{r, include=FALSE}
 variation_summary<-filter(gdd400_all_new, interval!="NA", CPP_YEAR=="2018") %>% group_by(pop, interval) %>% dplyr::summarize(lat=mean(latitude), country=first(country), gdd400_sd=sd(gdd400), wintermin_sd=sd(wintermin), cpp_2018=mean(Estimate))
@@
 ```
+
+# 2015 Diapause Incidence Field Trial
+
+```{r}
+field_data<-read.csv("DI_2015_FIELD.csv")
+field_data$WeekDate <- reorder(field_data$WeekDate, field_data$Week)
+ggplot(field_data, aes(x=WeekDate, y=DI, fill=Eggs)) + geom_point(size=3, shape=21) + scale_fill_manual(values=c("white","grey"), name="Mosquito Population") + theme_bw() + 
+theme(legend.position=c(.2,.8), axis.text.x = element_text(angle = 90, hjust = 1)) +
+xlab("Week") + ylab("Diapause incidence")
+
+field.model.fit <- drm(DI ~ Week, data = field_data, curveid = Eggs, fct = LL.5())
+summary(field.model.fit)
+# estimate the 50% intercept
+cpp_field<-data.frame(ED(field.model.fit, 50))
+
+```

### 2020-01-29 MS.Rmd
bw figures

--- 2020-01-28 MS.Rmd
@@
   # + geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE)
+
+ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, fill=country, shape=year)) + 
+  geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") +
+  stat_smooth(method=lm, se=FALSE, aes(linetype=year, color=country)) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + 
+  scale_color_manual(values=c("black","gray"), name="Country") +
+  guides(fill=guide_legend(override.aes=list(shape=22)), linetype=guide_legend(override.aes=list(color="black")),
+  color=guide_legend(override.aes=list(linetype=0)), color=guide_legend(override.aes=list(linetype=0)), linetype=guide_legend(legend.key.width=unit(3,"line"))) +
+  theme_bw() + theme(legend.position=c(.8,.25), text=element_text(size=14)) + xlab("Latitude") + ylab("Critical photoperiod (h)")
+
+ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, fill=country)) + 
+geom_point(size=3, shape=21) + scale_fill_manual(values=c("black", "gray"), name="Country") + stat_smooth(method=lm, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
+  facet_grid(~year) +
+  guides(fill=guide_legend(override.aes=list(shape=21))) +
+  theme_bw() + theme(legend.position=c(.8,.25), text=element_text(size=14)) + 
+  xlab("Latitude") + ylab("Critical photoperiod (h)")
+
 ```
@@
 remove_geom(japan_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="JP", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "null")), ncol=2, align="hv")
+
+#bw locations map
+us_map_bw<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="black") + 
+  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 1) + xlim(c(-87, -65)) + ylim(c(24,48)) +
+  geom_label(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) + theme_bw() + xlab("") + ylab("Latitude")
+
+japan_map_bw<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="black") + 
+  coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 1)  + xlim(c(125, 147)) + ylim(c(24,48)) +
+  geom_label_repel(data=filter(pops, country=="JP", only_2008=="n"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) + theme_bw()  + xlab("") + ylab("")
+
+plot_grid(us_map_bw, japan_map_bw)
+
 ```
@@
 #country_DI
-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year), se=FALSE) + scale_linetype_manual(values=c("dotted","solid")) + xlab("Latitude") + ylab("Short-day diapause incidence")
+country_DI_color<-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year), se=FALSE) + scale_linetype_manual(values=c("dotted","solid")) + xlab("Latitude") + ylab("Short-day diapause incidence")
+
+country_DI_bw<-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("black","gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3), text=element_text(size=14)) + guides(fill = guide_legend(override.aes=list(shape=22))) + xlab("Latitude") + ylab("Short-day diapause incidence")
+
 ```
@@
 plot(blockAB$yield ~ blockAB$TotalLarvaeEmbryos, xlab="Number of viable eggs on oviposition paper", ylab="Viable eggs/initial egg count on paper")
+ggplot(blockAB, aes(x=TotalLarvaeEmbryos, y=yield)) + geom_point() + xlab("Number of viable eggs on oviposition paper") + ylab("Viable eggs/initial egg count on paper") + theme_bw()
 ```

### 2020-01-29 MS.Rmd
curves

--- 2020-01-29 MS.Rmd
@@
 DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~fct_reorder(pop, lat)) + theme_bw() + xlab("Photoperiod (hours)") + ylab("Dipause incidence") 
-
 ```
@@
 cpp_values<-data.frame(ED(d.model.fit, 50))
+
+pm<-drc:::predict.drc(d.model.fit, type="response")
 ```
@@
 
+```{r}
+#Dose response curve plot from Jen
+cpp<-read.table("CPP.csv", header=TRUE, sep=",")
+cpp
+
+fm1<-cbind(hatched, nothatched)~photoperiod
+cppmodel<-glm(fm1, data=cpp, family=binomial())
+
+summary(cppmodel)
+
+fitteddata1<-predict(cppmodel,type="response")
+fitteddata1
+
+plot(cpp$photoperiod, fitteddata1, ylab="Probability of hatching",
+     xlab="photoperiod", ylim=c(0,1))
+     lines(cpp$photoperiod,fitteddata1, lty=1)
+y<-cpp$hatched/(cpp$hatched+cpp$nothatched)
+text(cpp$photoperiod,y)
+hours<-
+-(cppmodel$coefficients["(Intercept)"])/cppmodel$coefficients["photoperiod"]
+hours
+
+```
+
 # References

### 2020-01-31 GDDdata_input.Rmd
lab meeting

@@
 gdd400_jp$year<-as.numeric(gdd400_jp$year)
-ggplot(gdd400_jp, aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + geom_vline(xintercept=c(2008,2018))
 
@@
 ```{r}
+#model comparison of gdd-cpp relationship, with and without country as factor
+data_18<-filter(gdd400_all_new, year==2007, CPP_YEAR==2018)
+model1<-lm(Estimate~gdd400, data=data_18)
+model2<-lm(Estimate~gdd400*country, data=data_18)
+anova(model1, model2)
+
+model3<-lm(Estimate~latitude, data=data_18)
+model4<-lm(Estimate~latitude*country, data=data_18)
+anova(model3, model4)
+```
+
+```{r}
 ggplot(filter(gdd400_all_new, year %in% c(2007,2017)), aes(x=gdd400, y=Estimate, color=CPP_YEAR)) + scale_color_manual(values=c("gray", "black")) + geom_point() + facet_wrap(~year) + stat_smooth(method=lm)

### 2020-01-31 MS.Rmd
lab meeting

--- 2020-01-29 MS.Rmd
@@
 #country_DI
+
 country_DI_color<-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year), se=FALSE) + scale_linetype_manual(values=c("dotted","solid")) + xlab("Latitude") + ylab("Short-day diapause incidence")
@@
 country_DI_bw<-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("black","gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3), text=element_text(size=14)) + guides(fill = guide_legend(override.aes=list(shape=22))) + xlab("Latitude") + ylab("Short-day diapause incidence")
+
+ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year), se=FALSE) + scale_linetype_manual(values=c("dotted","solid")) + xlab("Latitude") + ylab("Short-day diapause incidence")
+
+ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("black","gray")) + 
+scale_shape_manual(values=c(21,24)) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year, color=country), se=FALSE) + scale_color_manual(values=c("black","gray")) + scale_linetype_manual(values=c("dotted","solid")) + xlab("Latitude") + ylab("Short-day diapause incidence")
 
@@
 
+```{r}
+#Dose response curve plot from Jen
+#cpp<-read.table("CPP.csv", header=TRUE, sep=",")
+#cpp
+
+#fm1<-cbind(hatched, nothatched)~photoperiod
+#cppmodel<-glm(fm1, data=cpp, family=binomial())
+
+#summary(cppmodel)
+
+#fitteddata1<-predict(cppmodel,type="response")
+#fitteddata1
+
+#plot(cpp$photoperiod, fitteddata1, ylab="Probability of hatching",
+ #    xlab="photoperiod", ylim=c(0,1))
+  #   lines(cpp$photoperiod,fitteddata1, lty=1)
+#y<-cpp$hatched/(cpp$hatched+cpp$nothatched)
+#text(cpp$photoperiod,y)
+#hours<-(cppmodel$coefficients["(Intercept)"])/cppmodel$coefficients["photoperiod"]
+#hours
+
+```
+
 # References

### 2020-02-05 GDDdata_input.Rmd
updated figures

@@
 ```
+
 ```{r}
@@
 ```{r}
-ggplot(filter(gdd400_all, only_2008=="n", year!=2018), aes(x=latitude, y=gdd400, color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") + geom_point() + facet_wrap(~year) + xlab("Latitude") + ylab("GDD Deadline (ordinal day)") + theme_bw()
+ggplot(filter(gdd400_all, only_2008=="n", year!=2018), aes(x=latitude, y=gdd400, fill=country)) + scale_fill_manual(values=c("black", "gray"), name="Country") + geom_point(shape=21) + facet_wrap(~year) + xlab("Latitude") + ylab("GDD Deadline (ordinal day)") + theme_bw() + theme(legend.position="bottom")
 
@@
 theme(legend.position=c(.2,.8), axis.text.x = element_text(angle = 90, hjust = 1)) +
-xlab("Week") + ylab("Diapause incidence")
+xlab("Week") + ylab("Diapause incidence") + geom_vline(aes(xintercept=10.39), linetype="dotted", color="grey") + geom_vline(aes(xintercept=9.99), linetype="dotted", color="black")
 

### 2020-02-05 MS.Rmd
updated figures

@@
 DI_summary$pop<-factor(DI_summary$pop, levels=c("PBS", "FEM", "OAK", "JAC", "BRU", "ZIO", "FAY", "NVA", "WAV", "MAN", "BER", "NEW", "OKI", "TAN", "KAG", "YAT", "SAG", "SHI", "YAM", "HIR", "KYO", "TAK", "TOK", "UTS", "KAN", "KHO", "AIZ", "NIG", "SEN", "SAK"))
-DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~fct_reorder(pop, lat)) + theme_bw() + xlab("Photoperiod (hours)") + ylab("Dipause incidence") 
+DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop, fill=country)) + geom_point(shape=21) + facet_wrap(~fct_reorder(pop, lat)) + theme_bw() + xlab("Photoperiod (hours)") + ylab("Dipause incidence") + scale_fill_manual(values=c("black","gray")) + theme(legend.position="NULL")
 ```
@@
 
-country_DI_bw<-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("black","gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3), text=element_text(size=14)) + guides(fill = guide_legend(override.aes=list(shape=22))) + xlab("Latitude") + ylab("Short-day diapause incidence")
+country_DI_bw<-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("black","gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3), text=element_text(size=14)) + guides(fill = guide_legend(override.aes=list(shape=22))) + xlab("Latitude") + ylab("Short-day diapause incidence") + stat_smooth(aes(linetype=year, color=country), se=FALSE) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + scale_color_manual(values=c("black","gray"), name="Country")
 
-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year), se=FALSE) + scale_linetype_manual(values=c("dotted","solid")) + xlab("Latitude") + ylab("Short-day diapause incidence")
+ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year, color=country), se=FALSE) + scale_linetype_manual(values=c("dotted","solid")) + xlab("Latitude") + ylab("Short-day diapause incidence")
 

### 2020-02-05 MS.Rmd
Removed redundant figures

@@
 blockAB_filtered <- filter(blockAB, yield<=1) %>% filter(yield>=0.6) 
-#test removing papers that had premature larval hatch on paper
-#badpapers<-c("333-1", "334-1", "335-1", "336-1", "337-1", "338-1", "339-1", "340-1", "341-1", "342-1", "343-1","366-1", "367-1", "368-1", "369-1", "370-1", "371-1", "372-1", "373-1", "374-1", "375-1", "376-1")
-#blockAB_filtered <- filter(blockAB, yield<=1) %>% filter(yield>=0.6) %>% filter(!PaperID %in% badpapers)
 ```
@@
 DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop, fill=country)) + geom_point(shape=21) + facet_wrap(~fct_reorder(pop, lat)) + theme_bw() + xlab("Photoperiod (hours)") + ylab("Dipause incidence") + scale_fill_manual(values=c("black","gray")) + theme(legend.position="NULL")
-```
-
-```{r incidence_lat, echo=FALSE, include=FALSE}
-DI_short<-filter(DI_summary, pp==8)
-ggplot(DI_short, aes(x=lat, y=DI, color=country)) + geom_point() + theme_bw() +stat_smooth()
 ```
@@
 
-ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, fill=country, shape=year)) + 
+cpp_08_18_shifts_bw<-ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, fill=country, shape=year)) + 
   geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") +
@@
   theme_bw() + theme(legend.position=c(.8,.25), text=element_text(size=14)) + xlab("Latitude") + ylab("Critical photoperiod (h)")
-
-ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, fill=country)) + 
-geom_point(size=3, shape=21) + scale_fill_manual(values=c("black", "gray"), name="Country") + stat_smooth(method=lm, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
-  facet_grid(~year) +
-  guides(fill=guide_legend(override.aes=list(shape=21))) +
-  theme_bw() + theme(legend.position=c(.8,.25), text=element_text(size=14)) + 
-  xlab("Latitude") + ylab("Critical photoperiod (h)")
-
 ```
@@
 write_csv(DI_short_2, "diapause_incidence_8h_2008_2018.csv")
-incidence_lat_year_plot<-ggplot(DI_short_2, aes(x=lat, y=DI, color=year, label=pop)) + scale_color_manual(values=c("gray","black")) + 
-  geom_point() + facet_grid(~country) 
-country_DI_2018<-ggplot(DI_short_2018, aes(x=lat, y=DI, color=country)) + scale_color_manual(values=c("red","blue")) + geom_point() + theme_bw() + stat_smooth() + theme(legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "cm"))
-country_DI<-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country)) + scale_color_manual(values=c("red","blue")) + geom_point() + theme_bw() + stat_smooth() + theme(legend.position=c(0.9,0.1)) + facet_grid(~year)
 ```
@@
 dimodel_result<-data.frame(anova(dimodel))
-```
-
-```{r day_function, echo=FALSE, include=FALSE}
-findDate<-function(lat, pp, year) {
-#generate day lengths at latitude between July 1 and Dec 31
-x<-daylength(lat,182:365)
-#find position of day length in the daylength vector that is closest to photoperiod queried
-daynumber<-which.min(abs(x-pp))
-#convert day number to date
-origin<-as.character(paste(year,"-07-01",sep=""))
-return(as.character(as.Date(daynumber, origin=origin)))
-}
-```
-
-```{r approx_dates, include=FALSE, echo=FALSE, include=FALSE}
-cpp_2$date<-mapply(findDate, lat=cpp_2$lat, pp=cpp_2$Estimate, year=cpp_2$year)
-cpp_2$estimated_date<-format(as.Date(cpp_2$date), "%mm-%dd")
-cpp_2$estimated_date<-as.Date(cpp_2$estimated_date, format="%mm-%dd")
-date_plot<-ggplot(cpp_2, aes(x=lat, y=estimated_date, color=year)) + geom_point() + scale_y_date(date_breaks = "weeks" , date_minor_breaks="days", date_labels = "%b-%d") +  facet_grid(~country) + theme_bw() + stat_smooth(method=lm, se=FALSE)
 ```
@@
 
-plot_grid(us_map_bw, japan_map_bw)
+map_panels_bw<-plot_grid(japan_map_bw, us_map_bw)
 
@@
 
-```{r fig.cap="Figure 1: Diapause incidence in Japan and US in 2008 and 2018", echo=FALSE, include=FALSE}
-incidence_lat_year_plot
-```
-
 ```{r fig.cap="Figure 2B: Diapause incidence by country and year.
@@
 country_DI_bw<-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("black","gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3), text=element_text(size=14)) + guides(fill = guide_legend(override.aes=list(shape=22))) + xlab("Latitude") + ylab("Short-day diapause incidence") + stat_smooth(aes(linetype=year, color=country), se=FALSE) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + scale_color_manual(values=c("black","gray"), name="Country")
-
-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, color=country, shape=year)) + scale_color_manual(values=c("red","blue")) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year, color=country), se=FALSE) + scale_linetype_manual(values=c("dotted","solid")) + xlab("Latitude") + ylab("Short-day diapause incidence")
-
-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("black","gray")) + 
-scale_shape_manual(values=c(21,24)) + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3)) + stat_smooth(aes(linetype=year, color=country), se=FALSE) + scale_color_manual(values=c("black","gray")) + scale_linetype_manual(values=c("dotted","solid")) + xlab("Latitude") + ylab("Short-day diapause incidence")
 
@@
 ```{r fig.cap="Figure 4: Shifts in critical photoperiod latitudinal clines. Lines represent linear regressions. Solid lines, 2018; dotted lines, 2008.", echo=FALSE}
-cpp_08_18_shifts
+cpp_08_18_shifts_bw
 ```
@@
 
-```{r fig.cap="Figure S1: Discrepancies between initial egg counts and final calculated egg counts on oviposition papers.", echo=FALSE}
-plot(blockAB$yield ~ blockAB$TotalLarvaeEmbryos, xlab="Number of viable eggs on oviposition paper", ylab="Viable eggs/initial egg count on paper")
-ggplot(blockAB, aes(x=TotalLarvaeEmbryos, y=yield)) + geom_point() + xlab("Number of viable eggs on oviposition paper") + ylab("Viable eggs/initial egg count on paper") + theme_bw()
-```
+```{r, echo=FALSE}
+yield_plot<-ggplot(blockAB, aes(x=TotalLarvaeEmbryos, y=yield)) + geom_point(alpha=0.5) + xlab("Number of viable eggs on oviposition paper") + ylab("Viable eggs/initial egg count on paper") + theme_bw()
 
-```{r fig.cap="Figure S2.
-Critical photoperiod calculated with and without filtering oviposition papers based on criteria described.
-Line represents x=y.
-Outlier point corresponds to population PBS.", echo=FALSE}
 #comparing CPP values calculated with and without filtering out unreliable oviposition papers
@@
 cpp_compare<-left_join(cpp_filtered,cpp_unfiltered,by="pop")
-ggplot(cpp_compare, aes(x=CPP_est_unfilt, y=CPP_est_filt)) + geom_point() + geom_abline(slope=1, intercept=0) + xlab("CPP without filtering criteria") + ylab("CPP calculated with filtering criteria") + theme_bw()
+filter_compare_plot<-ggplot(cpp_compare, aes(x=CPP_est_unfilt, y=CPP_est_filt)) + geom_point() + geom_abline(slope=1, intercept=0) + xlab("CPP without filtering criteria") + ylab("CPP calculated with filtering criteria") + theme_bw()
 ```
 
-```{r fig.cap="Figure S3: Photoperiod response curves", echo=FALSE}
-DI_plot 
+```{r, echo=FALSE}
+quality_control_fig<-plot_grid(yield_plot, filter_compare_plot, labels="AUTO")
+quality_control_fig
 ```
 
-```{r fig.cap="Supplemental figure TBD.
-Diapause incidence (2018) by country.
-Dark sections of pie charts represent diapausing fraction; white fractions represent non-diapausing.", echo=FALSE, message=FALSE}
-DI_pies_2018
+```{r, echo=FALSE}
+DI_plot 
 ```
@@
 
-```{r}
-#Dose response curve plot from Jen
-#cpp<-read.table("CPP.csv", header=TRUE, sep=",")
-#cpp
-
-#fm1<-cbind(hatched, nothatched)~photoperiod
-#cppmodel<-glm(fm1, data=cpp, family=binomial())
-
-#summary(cppmodel)
-
-#fitteddata1<-predict(cppmodel,type="response")
-#fitteddata1
-
-#plot(cpp$photoperiod, fitteddata1, ylab="Probability of hatching",
- #    xlab="photoperiod", ylim=c(0,1))
-  #   lines(cpp$photoperiod,fitteddata1, lty=1)
-#y<-cpp$hatched/(cpp$hatched+cpp$nothatched)
-#text(cpp$photoperiod,y)
-#hours<-(cppmodel$coefficients["(Intercept)"])/cppmodel$coefficients["photoperiod"]
-#hours
-
-```
-
 # References

### 2020-02-06 MS.Rmd
Added GDD calculations

@@
 map_panels_bw<-plot_grid(japan_map_bw, us_map_bw)
+```
 
+```{r, echo=FALSE}
+#GDD deadline data
+#importing and reshaping table
+gdd400_us <- read.csv("GDD400_US.csv", check.names=FALSE) %>% gather(year, gdd400, "1998":"2018", factor_key=FALSE)
+gdd400_us$year<-as.numeric(gdd400_us$year)
+gdd400_jp <- read.csv("GDD400_JP.csv", check.names=FALSE) %>% gather(year, gdd400, "1998":"2018", factor_key=FALSE)
+gdd400_jp$year<-as.numeric(gdd400_jp$year)
+#adding latitudes
+locations<-read.csv("mapping_pops_cpp_expt.csv")
+gdd400_us<-left_join(gdd400_us, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
+gdd400_jp<-left_join(gdd400_jp, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
+gdd400_all<-rbind(gdd400_us, gdd400_jp)
+gdd400_all$pop<-as.factor(gdd400_all$pop)
+gdd400_all<-mutate(gdd400_all, interval=case_when(year<=2017 & year>=2008 ~ 'y2008-2017',
+                                                  year<2008 ~ 'y1998-2007'))
+```
+
+```{r, echo=FALSE}
+gdd_deadline_plots<-ggplot(filter(gdd400_all, only_2008=="n", year!=2018), aes(x=latitude, y=gdd400, fill=country)) + scale_fill_manual(values=c("black", "gray"), name="Country") + geom_point(shape=21) + facet_wrap(~year) + xlab("Latitude") + ylab("GDD Deadline (ordinal day)") + theme_bw() + theme(legend.position="bottom")
+```
+
+```{r}
+#adding 2008 and 2018 CPP values 
+cpp_year<-dplyr::select(read.csv("cpp_calcs_filtered.csv"), pop, Estimate, year) %>% filter(year!="1988") %>% rename("CPP_YEAR"="year")
+gdd400_all_new<-left_join(gdd400_all, cpp_year, by="pop")
+gdd400_all_new$CPP_YEAR<-as.factor(gdd400_all_new$CPP_YEAR)
+
+```
+
+```{r}
+#plotting GDD deadlines and CPP
+gdd_08 <- ggplot(filter(gdd400_all_new, year==2007, CPP_YEAR==2008), aes(x=gdd400, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=3, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
+  theme_bw() + theme(legend.position="none", text=element_text(size=14)) +
+  xlab("GDD Deadline (ordinal day), 2007") + ylab("Critical photoperiod (h), 2008")
+
+gdd_18 <- ggplot(filter(gdd400_all_new, year==2017, CPP_YEAR==2018), aes(x=gdd400, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=3, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
+theme_bw() + theme(legend.position=c(.8,.8), text=element_text(size=14)) +
+xlab("GDD Deadline (ordinal day), 2017") + ylab("Critical photoperiod (h), 2018")
+
+gdd_cpp_plot<-plot_grid(gdd_08, gdd_18, labels="AUTO")
+```
+
+```{r}
+#2015 Diapause Incidence Field Trial
+field_data<-read.csv("DI_2015_FIELD.csv")
+field_data$WeekDate <- reorder(field_data$WeekDate, field_data$Week)
+ggplot(field_data, aes(x=WeekDate, y=DI, fill=Eggs)) + geom_point(size=3, shape=21) + scale_fill_manual(values=c("white","grey"), name="Mosquito Population") + theme_bw() + 
+theme(legend.position=c(.2,.8), axis.text.x = element_text(angle = 90, hjust = 1)) +
+xlab("Week") + ylab("Diapause incidence") + geom_vline(aes(xintercept=10.39), linetype="dotted", color="grey") + geom_vline(aes(xintercept=9.99), linetype="dotted", color="black")
+
+field.model.fit <- drm(DI ~ Week, data = field_data, curveid = Eggs, fct = LL.5())
+summary(field.model.fit)
+# estimate the 50% intercept
+cpp_field<-data.frame(ED(field.model.fit, 50))
+
+```
+
+```{r}
+data_18<-gdd400_all_new %>% filter(CPP_YEAR==2018) %>% filter(year==2017)
+model1<-lm(Estimate~gdd400*country, data=data_18)
+anova(model1)
+
+data_08<-gdd400_all_new %>% filter(CPP_YEAR==2008) %>% filter(year==2007)
+model2<-lm(Estimate~gdd400*country, data=data_08)
+anova(model2)
 ```

### 2020-02-06 MS.Rmd
field data update

@@
 field_data$WeekDate <- reorder(field_data$WeekDate, field_data$Week)
-ggplot(field_data, aes(x=WeekDate, y=DI, fill=Eggs)) + geom_point(size=3, shape=21) + scale_fill_manual(values=c("white","grey"), name="Mosquito Population") + theme_bw() + 
-theme(legend.position=c(.2,.8), axis.text.x = element_text(angle = 90, hjust = 1)) +
-xlab("Week") + ylab("Diapause incidence") + geom_vline(aes(xintercept=10.39), linetype="dotted", color="grey") + geom_vline(aes(xintercept=9.99), linetype="dotted", color="black")
+ggplot(filter(field_data, Eggs=="Field"), aes(x=WeekDate, y=DI)) + geom_point(size=3, shape=21, fill="gray") + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+xlab("Week") + ylab("Diapause incidence") + geom_vline(aes(xintercept=10.4), linetype="dotted")
 
-field.model.fit <- drm(DI ~ Week, data = field_data, curveid = Eggs, fct = LL.5())
+field.model.fit <- drm(DI ~ Week, data = filter(field_data, Eggs=="Field"), fct = LL.5())
 summary(field.model.fit)

### 2020-02-10 MS.Rmd
GDD methods

@@
 ```{r, echo=FALSE}
+#GDD deadline data
+#importing and reshaping table
+gdd465_us <- read.csv("GDD465_US.csv", check.names=FALSE) %>% gather(year, gdd465, "1998":"2018", factor_key=FALSE)
+gdd465_us$year<-as.numeric(gdd465_us$year)
+gdd465_jp <- read.csv("GDD465_JP.csv", check.names=FALSE) %>% gather(year, gdd465, "1998":"2018", factor_key=FALSE)
+gdd465_jp$year<-as.numeric(gdd465_jp$year)
+gdd465_all<-rbind(gdd465_us, gdd465_jp)
+gdd465_all$pop<-as.factor(gdd465_all$pop)
+gdd_all<-left_join(gdd400_all, gdd465_all)
+gdd_all$gdd<-(gdd_all$gdd400 + gdd_all$gdd465)/2
+```
+
+```{r, echo=FALSE}
 gdd_deadline_plots<-ggplot(filter(gdd400_all, only_2008=="n", year!=2018), aes(x=latitude, y=gdd400, fill=country)) + scale_fill_manual(values=c("black", "gray"), name="Country") + geom_point(shape=21) + facet_wrap(~year) + xlab("Latitude") + ylab("GDD Deadline (ordinal day)") + theme_bw() + theme(legend.position="bottom")
+
+ggplot(filter(gdd_all, only_2008=="n", year!=2018), aes(x=latitude, y=gdd, fill=country)) + scale_fill_manual(values=c("black", "gray"), name="Country") + geom_point(shape=21) + facet_wrap(~year) + xlab("Latitude") + ylab("GDD Deadline (ordinal day)") + theme_bw() + theme(legend.position="bottom")
+
 ```
@@
 gdd400_all_new$CPP_YEAR<-as.factor(gdd400_all_new$CPP_YEAR)
+
+gdd_all_new<-left_join(gdd_all, cpp_year, by="pop")
+gdd_all_new$CPP_YEAR<-as.factor(gdd_all_new$CPP_YEAR)
+
+```
+
+```{r echo=FALSE}
+#GDD deadlines averaged across all years
+gdd_summaries<-gdd_all_new %>% filter(interval!="NA") %>% group_by(pop) %>% dplyr::summarise(gdd400_mean=mean(gdd400), gdd400_sd=sd(gdd400), gdd465_mean=mean(gdd465), gdd465_sd=sd(gdd465), latitude=mean(latitude), country=first(country), only_2008=first(only_2008))
+
+lat_gdd_average<-ggplot(filter(gdd_summaries, only_2008=="n"), aes(x=latitude, y=gdd400_mean, fill=country)) + geom_point(shape=21) + scale_fill_manual(values=c("black", "gray")) + geom_errorbar(aes(ymin=gdd400_mean-gdd400_sd, ymax=gdd400_mean+gdd400_sd, color=country)) + scale_color_manual(values=c("black","gray")) + theme_bw() + theme(legend.position="bottom") + xlab("Latitude (N)") + ylab("GDD Deadline (ordinal day)")
+
+```
+
+```{r}
+m1<-lm(cpp~gdd400_mean*country, data=filter(interval_summaries, interval=="y2008-2017", CPP_YEAR==2018))
+m2<-lm(cpp~gdd400_mean*country, data=filter(interval_summaries, interval=="y1998-2007", CPP_YEAR==2008))
 
@@
 gdd_cpp_plot<-plot_grid(gdd_08, gdd_18, labels="AUTO")
+
+ggplot(filter(gdd_all_new, year==2007, CPP_YEAR==2008), aes(x=gdd, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=3, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
+  theme_bw() + theme(legend.position="none", text=element_text(size=14)) +
+  xlab("GDD Deadline (ordinal day), 2007") + ylab("Critical photoperiod (h), 2008")
+
+ggplot(filter(gdd_all_new, year==2017, CPP_YEAR==2018), aes(x=gdd, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=3, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
+theme_bw() + theme(legend.position=c(.8,.8), text=element_text(size=14)) +
+xlab("GDD Deadline (ordinal day), 2017") + ylab("Critical photoperiod (h), 2018")
+
+```
+```{r}
+#plotting CPP by GDD deadlines averaged over intervals
+interval_summaries<-gdd_all_new %>% filter(interval!="NA") %>% group_by(pop, interval, CPP_YEAR) %>% dplyr::summarise(gdd400_mean=mean(gdd400), gdd400_sd=sd(gdd400), cpp=mean(Estimate), latitude=mean(latitude), country=first(country), only_2008=first(only_2008))
+
+ggplot(filter(interval_summaries, interval=="y2008-2017", CPP_YEAR==2018), aes(x=gdd400_mean, y=cpp, fill=country)) + geom_point(shape=21) + scale_fill_manual(values=c("black","gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") + theme_bw() + theme(legend.position="bottom") + xlab("Mean GDD deadline, 2008-2017 (ordinal day)") + ylab("Critical photoperiod, 2018 (h)")
+
+ggplot(filter(interval_summaries, interval=="y1998-2007", CPP_YEAR==2008), aes(x=gdd400_mean, y=cpp, fill=country)) + geom_point(shape=21) + scale_fill_manual(values=c("black","gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") + theme_bw() + theme(legend.position="bottom") + xlab("Mean GDD deadline, 1998-2007 (ordinal day)") + ylab("Critical photoperiod, 2008 (h)")
+
 ```
@@
 field_data$WeekDate <- reorder(field_data$WeekDate, field_data$Week)
-ggplot(filter(field_data, Eggs=="Field"), aes(x=WeekDate, y=DI)) + geom_point(size=3, shape=21, fill="gray") + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+field_data_plot<-ggplot(filter(field_data, Eggs=="Field"), aes(x=WeekDate, y=DI)) + geom_point(size=3, shape=21, fill="gray") + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
 xlab("Week") + ylab("Diapause incidence") + geom_vline(aes(xintercept=10.4), linetype="dotted")
@@
 anova(model2)
+
+data_18<-gdd_all_new %>% filter(CPP_YEAR==2018) %>% filter(year==2017)
+model1<-lm(Estimate~gdd*country, data=data_18)
+anova(model1)
+
 ```
@@
 
+### Growing degree day models
+We reasoned that the optimal time for entering diapause depended on the amount of time required to complete a full generation under particular climatic conditions.
+Insect phenology studies often estimate generation times using growing degree day (GDD) models.
+Based on a survey of literature of Aedes albopictus growth rates under various temperature conditions, we estimated that a mosquito generation - from the egg stage through ovipostion - required 400-465 GDD.
+We then used climate records over the past 20 years from our sampling locations to calculate a yearly GDD-based "deadline" for diapause at each location: the date after which eggs laid would not have enough growing degree days left in the season to complete another generation.
+Further details about how these dates were calculated can be found in Supplementary File X.
+
+Although field data about Aedes albopictus diapause egg production are scarce, we were able to use field observations collected in 2015 at site in Washington, DC, USA.
+Briefly, oviposition cups were placed outdoors and Aedes albopictus eggs collected weekly were scored for diapause incidence using the same methods described here.
+A five-parameter log-logistic model was fit to the diapause incidence curve (Fig S) to determine the critical date at which the majority of eggs produced by wild mosquitoes were in diapause.
+The critical date occurred during the week of September 7, 2015, corresponding to ordinal day 250.
+The calculated deadline for Manassas, VA in 2015 was 256-263.
+In 2014, the calculated deadline was 247-252.
+These results confirmed that GDD deadlines generally reasonably correspond to field observations and suggested that examining the GDD deadline in years preceding the collection year is a reasonable approach.
+
 ### Statistical analysis
@@
 
+GDD deadlines calculated using historical climate data correlated with latitude, as expected.
+The latitude-specific GDD deadline was earlier in Japan in the US for most years examined (Figure SX), indicating that although the Japanese locations generally experienced a later date of first frost, the limits of the growing seasons in the two regions may be better captured by measures such as GDD that take into account the distribution of temperature conditions over the season.
+
+We examined the relationship between critical photoperiod and two GDD deadline values for each population: the GDD deadline in the year immediately preceding the collection of the 2008 and 2018 populations, and the mean GDD deadline over the preceding 10-year interval.
+These four correlations (2008 CPP vs 2007 GDD deadline; 2008 CPP vs 1998-2007 mean deadline; 2018 CPP vs 2017 deadline; and 2018 CPP vs.
+2008-2017 mean deadline) were analyzed independently.
+In all four models, the country effect was not siginificant, meaning that GDD deadlines predicted critical photoperiods across countries.
+There was also a significant GDD deadline by country interaction effect in the model in which 2017 deadline was used as an explanatory variable for 2018 CPP data, but not when the GDD deadline was averaged over the preceding 10 years, suggesting that year-to-year variation may additionally affect the
+
 ## Supplementary File 1
@@
 The total number of hatched larvae and embryos could differ from the number of eggs initially counted on the paper for the following reasons: 1) presence of inviable eggs or larvae (i.e., eggs that fail to hatch due to developmental abnormalities or larvae that die before being counted), 2) loss of eggs while handling oviposition papers, or 3) inaccurate counting at any step of the scoring process.
-Based on the distribution of percent differences from the initial counts (Figure S1), we considered oviposition papers with over 100% or less than 60% of the initial count to be less reliable data points.
+Based on the distribution of percent differences from the initial counts (Figure S1A), we considered oviposition papers with over 100% or less than 60% of the initial count to be less reliable data points.
 Excluding these data points resulted in 5 cages out of 286 in the total data set having less than 100 remaining eggs, so they were excluded from the photoperiod response curves.

### 2020-02-10 MS.Rmd
combined graphs

--- 2020-02-06 MS.Rmd
@@
 ```{r, echo=FALSE}
+#GDD deadline data
+#importing and reshaping table
+gdd465_us <- read.csv("GDD465_US.csv", check.names=FALSE) %>% gather(year, gdd465, "1998":"2018", factor_key=FALSE)
+gdd465_us$year<-as.numeric(gdd465_us$year)
+gdd465_jp <- read.csv("GDD465_JP.csv", check.names=FALSE) %>% gather(year, gdd465, "1998":"2018", factor_key=FALSE)
+gdd465_jp$year<-as.numeric(gdd465_jp$year)
+gdd465_all<-rbind(gdd465_us, gdd465_jp)
+gdd465_all$pop<-as.factor(gdd465_all$pop)
+gdd_all<-left_join(gdd400_all, gdd465_all)
+gdd_all$gdd<-(gdd_all$gdd400 + gdd_all$gdd465)/2
+```
+
+```{r, echo=FALSE}
 gdd_deadline_plots<-ggplot(filter(gdd400_all, only_2008=="n", year!=2018), aes(x=latitude, y=gdd400, fill=country)) + scale_fill_manual(values=c("black", "gray"), name="Country") + geom_point(shape=21) + facet_wrap(~year) + xlab("Latitude") + ylab("GDD Deadline (ordinal day)") + theme_bw() + theme(legend.position="bottom")
+
 ```
@@
 gdd400_all_new$CPP_YEAR<-as.factor(gdd400_all_new$CPP_YEAR)
+
+gdd_all_new<-left_join(gdd_all, cpp_year, by="pop")
+gdd_all_new$CPP_YEAR<-as.factor(gdd_all_new$CPP_YEAR)
+
+```
+
+```{r echo=FALSE}
+#GDD deadlines averaged across all years
+gdd_summaries<-gdd_all_new %>% filter(interval!="NA") %>% group_by(pop) %>% dplyr::summarise(gdd400_mean=mean(gdd400), gdd400_sd=sd(gdd400), gdd465_mean=mean(gdd465), gdd465_sd=sd(gdd465), latitude=mean(latitude), country=first(country), only_2008=first(only_2008))
+
+lat_gdd_average<-ggplot(filter(gdd_summaries, only_2008=="n"), aes(x=latitude, y=gdd400_mean, fill=country)) + geom_point(shape=21) + scale_fill_manual(values=c("black", "gray")) + geom_errorbar(aes(ymin=gdd400_mean-gdd400_sd, ymax=gdd400_mean+gdd400_sd, color=country)) + scale_color_manual(values=c("black","gray")) + theme_bw() + theme(legend.position="bottom") + xlab("Latitude (N)") + ylab("GDD Deadline (ordinal day)")
 
@@
 
-gdd_cpp_plot<-plot_grid(gdd_08, gdd_18, labels="AUTO")
+```
+
+```{r}
+#plotting CPP by GDD deadlines averaged over intervals
+interval_summaries<-gdd_all_new %>% filter(interval!="NA") %>% group_by(pop, interval, CPP_YEAR) %>% dplyr::summarise(gdd400_mean=mean(gdd400), gdd400_sd=sd(gdd400), cpp=mean(Estimate), latitude=mean(latitude), country=first(country), only_2008=first(only_2008))
+
+gdd_cpp_interval_18<-ggplot(filter(interval_summaries, interval=="y2008-2017", CPP_YEAR==2018), aes(x=gdd400_mean, y=cpp, fill=country)) + geom_point(size=3, shape=21) + scale_fill_manual(values=c("black","gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") + theme_bw() + theme(legend.position="none", text=element_text(size=14)) + xlab("Mean GDD deadline, 2008-2017 (ordinal day)") + ylab("Critical photoperiod, 2018 (h)")
+
+gdd_cpp_interval_08<-ggplot(filter(interval_summaries, interval=="y1998-2007", CPP_YEAR==2008), aes(x=gdd400_mean, y=cpp, fill=country)) + geom_point(size=3, shape=21) + scale_fill_manual(values=c("black","gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") + theme_bw() + theme(legend.position="none", text=element_text(size=14)) + xlab("Mean GDD deadline, 1998-2007 (ordinal day)") + ylab("Critical photoperiod, 2008 (h)")
+
+```
+
+```{r}
+gdd_cpp_plot<-plot_grid(gdd_08, gdd_18, gdd_cpp_interval_08, gdd_cpp_interval_18, nrow=2, labels="AUTO")
+
 ```
@@
 model1<-lm(Estimate~gdd400*country, data=data_18)
-anova(model1)
 
@@
 model2<-lm(Estimate~gdd400*country, data=data_08)
+
+model3<-lm(cpp~gdd400_mean*country, data=filter(interval_summaries, interval=="y2008-2017", CPP_YEAR==2018))
+model4<-lm(cpp~gdd400_mean*country, data=filter(interval_summaries, interval=="y1998-2007", CPP_YEAR==2008))
+
+anova(model1)
 anova(model2)
+anova(model3)
+anova(model4)
+
+```
+
+```{r}
+write.csv(filter(gdd400_all_new, year %in% c(2007,2017)), "gdd_cpp_07_17.csv")
+prev_year<-read.csv("gdd_prevyear.csv")
+prev_year$CPP_YEAR<-as.factor(prev_year$CPP_YEAR)
+
+write.csv(interval_summaries, "interval_summaries.csv")
+prev_interval<-read.csv("gdd_previnterval.csv")
+prev_interval$CPP_YEAR<-as.factor(prev_interval$CPP_YEAR)
+
+```
+
+```{r}
+prevyear_all<-ggplot(prev_year, aes(x=prev_year_gdd400, y=Estimate, fill=country, shape=CPP_YEAR)) + 
+  geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") +
+  stat_smooth(method=lm, se=FALSE, aes(linetype=CPP_YEAR, color=country)) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + 
+  scale_color_manual(values=c("black","gray"), name="Country") +
+  guides(fill=guide_legend(override.aes=list(shape=22)), linetype=guide_legend(override.aes=list(color="black")),
+  color=guide_legend(override.aes=list(linetype=0)), color=guide_legend(override.aes=list(linetype=0)), linetype=guide_legend(legend.key.width=unit(3,"line"))) +
+  theme_bw() + theme(legend.position="none", text=element_text(size=14)) + xlab("Previous year's GDD deadline (ordinal day)") + ylab("Critical photoperiod (h)")
+
+previnterval_all<-ggplot(prev_interval, aes(x=gdd400_mean_previnterval, y=cpp, fill=country, shape=CPP_YEAR)) + 
+  geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") +
+  stat_smooth(method=lm, se=FALSE, aes(linetype=CPP_YEAR, color=country)) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + 
+  scale_color_manual(values=c("black","gray"), name="Country") +
+  guides(fill=guide_legend(override.aes=list(shape=22)), linetype=guide_legend(override.aes=list(color="black")),
+  color=guide_legend(override.aes=list(linetype=0)), color=guide_legend(override.aes=list(linetype=0)), linetype=guide_legend(legend.key.width=unit(3,"line"))) +
+  theme_bw() + theme(legend.position=c(.8,.75), text=element_text(size=14)) + xlab("Previous decade's GDD deadline (ordinal day)") + ylab("Critical photoperiod (h)")
+
+plot_grid(prevyear_all, previnterval_all, labels="AUTO")
+```
+```{r}
+all_prevyear_model<-lm(Estimate~prev_year_gdd400*country*CPP_YEAR, data=prev_year)
+all_previnterval_model<-lm(cpp~gdd400_mean_previnterval*country*CPP_YEAR, data=prev_interval)
 ```
@@
 Japanese population samples were confirmed to be Ae.
-albopictus via morphological identification following [@tanakaRevisionAdultLarval1979] and sent to Georgetown as F1 eggs (except for TAK, for which F2 eggs were provided).
+albopictus via morphological identification following [@tanakaRevisionAdultLarval1979] and sent to Georgetown as F1 or F2 eggs.
 At least 50 male and 50 female adult mosquitoes were used to initiate all populations in the laboratory at Georgetown, except for SEN, which had 37 females and 39 males.
@@
 
+### Growing degree day models
+
 ### Statistical analysis
@@
 
-## Results
+## Results and Discussion (mostly Results)
 
@@
 DI was high (generally greater than 90%) at latitudes north of ~32.5 degrees in both countries.
-Substantial fractions of constitutively non-diapausing eggs were observed in populations south of South Carolina in the US and in Okinawa in Japan (latitude effect p=<0.001).
-Latitude-specific DI was lower in the US than in Japan (country effect p=0.026), though there were fewer sites sampled at lower latitudes in Japan than the US due to lack of land between Okinawa and Kagashima.
+Substantial fractions of constitutively non-diapausing eggs were observed in populations south of South Carolina in the US and in Okinawa in Japan (latitude p=<0.001).
+Latitude-specific DI was lower in the US than in Japan (country p=0.026), though there were fewer sites sampled at lower latitudes in Japan than the US due to lack of land between Okinawa and Kagashima.
 
@@
 This implies earlier entry into diapause at the same latitude in Japan than in the US, as well as earlier entry into diapause in the same locations in 2018 than in 2008.
+Both of these differences are contrary to expectation based on global warming trends and the consistently milder winters in Japan compared to US.
 
-```{r fig.cap="Figure 3: Critical photoperiod (CPP) as a function of latitude in the US and Japan in 1988, 2008, and 2018.
-Error bars represent standard errors of CPP estimates.
-Errors around 1988 CPP estimates are absent because experiment and calculations were conducted using different methods and raw data were not available.", echo=FALSE, message=FALSE, include=FALSE}
-cpp_japan_us_years
-```
-
-```{r fig.cap="Figure 4: Shifts in critical photoperiod latitudinal clines. Lines represent linear regressions. Solid lines, 2018; dotted lines, 2008.", echo=FALSE}
+```{r fig.cap="Figure : Shifts in critical photoperiod latitudinal clines. Lines represent linear regressions. Solid lines, 2018; dotted lines, 2008.", echo=FALSE}
 cpp_08_18_shifts_bw
@@
 ```{r table3, echo=FALSE}
-kable(prettify(anova(cppmodel)), caption="Table 3: Effects of latitude, country and year on critical photoperiod in 2008 and 2018.")
+kable(prettify(anova(cppmodel)), digits=3, caption="Table 3: Effects of latitude, country and year on critical photoperiod in 2008 and 2018.")
 ```
+
+We reasoned that the optimal time for entering diapause depended on the amount of time required to complete a full generation under particular climatic conditions.
+Insect phenology studies often estimate generation times using growing degree day (GDD) models.
+Based on a survey of literature of Aedes albopictus growth rates under various temperature conditions, we estimated that a mosquito generation - from the egg stage through ovipostion - required 400-465 GDD.
+We used climate records over the past 20 years from our sampling locations to calculate a yearly GDD-based "deadline" for diapause at each location: the date after which eggs laid would not have enough growing degree days left in the year to complete another generation.
+
+Although field data about Aedes albopictus diapause egg production are scarce, we were able to use field observations collected in 2015 at site in Washington, DC, USA as preliminary verification of the GDD approach.
+Briefly, oviposition cups were placed outdoors and Aedes albopictus eggs collected weekly were scored for diapause incidence using the same methods described here.
+The diapause incidence curve (Fig S) was used to determine the critical date at which the majority of eggs produced by wild mosquitoes were in diapause.
+The critical date occurred during the week of September 7, 2015, corresponding to ordinal day 250.
+The calculated deadline for Manassas, VA (~50 km from DC) in 2015 was 256-263.
+In 2014, the calculated deadline was 247-252.
+These results confirmed that GDD deadlines generally correspond to field observations and suggested that examining the GDD deadline in years preceding the collection year is a reasonable approach.
+
+GDD deadlines calculated using historical climate data correlated with latitude, as expected.
+The latitude-specific GDD deadline was earlier in Japan in the US for most years examined (Figure SX), indicating that although the Japanese locations generally experienced a later date of first frost, the limits of the growing seasons in the two regions may be better captured by measures such as GDD that take into account the distribution of temperature conditions over the season.
+
+We examined the relationship between critical photoperiod and two GDD deadline values for each population: the 400 GDD deadline in the year immediately preceding the 2008 and 2018 studies, and the mean 400 GDD deadline over the preceding 10-year interval.
+[NB: 2008 populations were actually collected over the period 2005-2008...] These four correlations (2008 CPP vs 2007 GDD deadline; 2008 CPP vs 1998-2007 mean deadline; 2018 CPP vs 2017 deadline; and 2018 CPP vs.
+2008-2017 mean deadline) were analyzed independently.
+In all four models, the country effect was not significant, suggesting that GDD deadlines predicted critical photoperiods across countries.
+There was also a significant GDD deadline by country interaction effect in the model in which 2017 deadline was used as an explanatory variable for 2018 CPP data, but not when the GDD deadline was averaged over the preceding 10 years, suggesting that year-to-year variation may additionally affect the relationship between local GDD conditions and local CPP.
+
+```{r}
+gdd_cpp_plot
+```
+
+```{r}
+kable(rbind(prettify(anova(model1)), prettify(anova(model2)), prettify(anova(model3)), prettify(anova(model4))), digits=3)  %>% pack_rows("2018 CPP vs 2017 GDD Deadline", 1, 3) %>% pack_rows("2008 CPP vs 2007 GDD Deadline", 4, 6) %>% pack_rows("2018 CPP vs 2008-2017 Average GDD Deadline", 7, 9) %>% pack_rows("2008 CPP vs 1998-2007 Average GDD Deadline", 10, 12)
+
+```
+The GDD-based estimate of diapause deadline date does not explain the shift to earlier critical photoperiod between 2008 and 2018. Although deadline dates vary from year to year
+
+We cannot exclude the possibility that the temporal shift in CPP between the 2008 and 2018 studies was due to experimental variability.
+For example, some of the populations in the 2008 study had been maintained in the laboratory for up to 14 generations, whereas in the current experiment they were reared for 3-4 laboratory generations; it is unknown how quickly acclimation or adaptation to laboratory conditions occurs in this species.
+In addition, there were likely cryptic differences in laboratory rearing conditions between 2008 and 2018, as evidenced by the fact that in the 2008 study more cages required multiple blood meals to produce sufficient eggs for analysis, hinting at higher mortality, lower fecundity, or both.
 
@@
 The total number of hatched larvae and embryos could differ from the number of eggs initially counted on the paper for the following reasons: 1) presence of inviable eggs or larvae (i.e., eggs that fail to hatch due to developmental abnormalities or larvae that die before being counted), 2) loss of eggs while handling oviposition papers, or 3) inaccurate counting at any step of the scoring process.
-Based on the distribution of percent differences from the initial counts (Figure S1), we considered oviposition papers with over 100% or less than 60% of the initial count to be less reliable data points.
+Based on the distribution of percent differences from the initial counts (Figure S1A), we considered oviposition papers with more than 100% or less than 60% of the initial count to be less reliable data points; these papers made up less than 10% of both individual papers and total eggs quantified in the experiment.
 Excluding these data points resulted in 5 cages out of 286 in the total data set having less than 100 remaining eggs, so they were excluded from the photoperiod response curves.
@@
 ```
+```{r}
+gdd_deadline_plots
+```
+```{r}
+plot_grid(ggplot(filter(gdd400_jp, only_2008=="n"), aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + theme_bw() + stat_smooth(method=lm, se=FALSE),ggplot(gdd400_us, aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + theme_bw() + stat_smooth(method=lm, se=FALSE), labels="AUTO")
+
+```
 

### 2020-02-11 MS.Rmd
all figures

@@
 output:
+  html_document: default
   word_document: default
-  html_document: default
   pdf_document: default
@@
 install.packages("cowplot")
+install.packages("flextable")
 ```
@@
 library("papeR")
+library("flextable")
 library("drc") #3.0.1
@@
   color=guide_legend(override.aes=list(linetype=0)), color=guide_legend(override.aes=list(linetype=0)), linetype=guide_legend(legend.key.width=unit(3,"line"))) +
-  theme_bw() + theme(legend.position=c(.8,.25), text=element_text(size=14)) + xlab("Latitude") + ylab("Critical photoperiod (h)")
+  theme_bw() + theme(legend.position=c(.8,.25), text=element_text(size=12)) + xlab("Latitude") + ylab("Critical photoperiod (h)")
 ```
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, include=FALSE}
 #GDD deadline data
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, include=FALSE}
 gdd_deadline_plots<-ggplot(filter(gdd400_all, only_2008=="n", year!=2018), aes(x=latitude, y=gdd400, fill=country)) + scale_fill_manual(values=c("black", "gray"), name="Country") + geom_point(shape=21) + facet_wrap(~year) + xlab("Latitude") + ylab("GDD Deadline (ordinal day)") + theme_bw() + theme(legend.position="bottom")
@@
 
-```{r}
+```{r echo=FALSE, include=FALSE}
 #adding 2008 and 2018 CPP values 
@@
 
-```{r echo=FALSE}
+```{r echo=FALSE, include=FALSE}
 #GDD deadlines averaged across all years
@@
 
-```{r}
+```{r echo=FALSE, include=FALSE}
 #plotting GDD deadlines and CPP
 gdd_08 <- ggplot(filter(gdd400_all_new, year==2007, CPP_YEAR==2008), aes(x=gdd400, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=3, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
-  theme_bw() + theme(legend.position="none", text=element_text(size=14)) +
-  xlab("GDD Deadline (ordinal day), 2007") + ylab("Critical photoperiod (h), 2008")
+  theme_bw() + theme(legend.position="none", text=element_text(size=12)) +
+  xlab("GDD Deadline (ordinal day), 2007") + ylab("Critical photoperiod (h), 2008") + ylim(12.3,14.2)
 
 gdd_18 <- ggplot(filter(gdd400_all_new, year==2017, CPP_YEAR==2018), aes(x=gdd400, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=3, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
-theme_bw() + theme(legend.position=c(.8,.8), text=element_text(size=14)) +
-xlab("GDD Deadline (ordinal day), 2017") + ylab("Critical photoperiod (h), 2018")
+theme_bw() + theme(legend.position=c(.8,.8), text=element_text(size=12)) +
+xlab("GDD Deadline (ordinal day), 2017") + ylab("Critical photoperiod (h), 2018") + ylim(12.3, 14.2)
 
@@
 
-```{r}
+```{r echo=FALSE, include=FALSE}
 #plotting CPP by GDD deadlines averaged over intervals
@@
 
-gdd_cpp_interval_18<-ggplot(filter(interval_summaries, interval=="y2008-2017", CPP_YEAR==2018), aes(x=gdd400_mean, y=cpp, fill=country)) + geom_point(size=3, shape=21) + scale_fill_manual(values=c("black","gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") + theme_bw() + theme(legend.position="none", text=element_text(size=14)) + xlab("Mean GDD deadline, 2008-2017 (ordinal day)") + ylab("Critical photoperiod, 2018 (h)")
+gdd_cpp_interval_18<-ggplot(filter(interval_summaries, interval=="y2008-2017", CPP_YEAR==2018), aes(x=gdd400_mean, y=cpp, fill=country)) + geom_point(size=3, shape=21) + scale_fill_manual(values=c("black","gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") + theme_bw() + theme(legend.position="none", text=element_text(size=12)) + xlab("Mean GDD deadline, 2008-2017 (ordinal day)") + ylab("Critical photoperiod, 2018 (h)") + ylim(12.3, 14.2)
 
-gdd_cpp_interval_08<-ggplot(filter(interval_summaries, interval=="y1998-2007", CPP_YEAR==2008), aes(x=gdd400_mean, y=cpp, fill=country)) + geom_point(size=3, shape=21) + scale_fill_manual(values=c("black","gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") + theme_bw() + theme(legend.position="none", text=element_text(size=14)) + xlab("Mean GDD deadline, 1998-2007 (ordinal day)") + ylab("Critical photoperiod, 2008 (h)")
+gdd_cpp_interval_08<-ggplot(filter(interval_summaries, interval=="y1998-2007", CPP_YEAR==2008), aes(x=gdd400_mean, y=cpp, fill=country)) + geom_point(size=3, shape=21) + scale_fill_manual(values=c("black","gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") + theme_bw() + theme(legend.position="none", text=element_text(size=12)) + xlab("Mean GDD deadline, 1998-2007 (ordinal day)") + ylab("Critical photoperiod, 2008 (h)") + ylim(12.3, 14.2)
 
@@
 
-```{r}
+```{r echo=FALSE, include=FALSE}
 gdd_cpp_plot<-plot_grid(gdd_08, gdd_18, gdd_cpp_interval_08, gdd_cpp_interval_18, nrow=2, labels="AUTO")
@@
 
-```{r}
+```{r echo=FALSE, include=FALSE}
 #2015 Diapause Incidence Field Trial
@@
 
-```{r}
+```{r echo=FALSE, include=FALSE}
 data_18<-gdd400_all_new %>% filter(CPP_YEAR==2018) %>% filter(year==2017)
@@
 
-```{r}
+```{r echo=FALSE, include=FALSE}
 write.csv(filter(gdd400_all_new, year %in% c(2007,2017)), "gdd_cpp_07_17.csv")
@@
 
-```{r}
+```{r echo=FALSE, include=FALSE}
 prevyear_all<-ggplot(prev_year, aes(x=prev_year_gdd400, y=Estimate, fill=country, shape=CPP_YEAR)) + 
@@
   color=guide_legend(override.aes=list(linetype=0)), color=guide_legend(override.aes=list(linetype=0)), linetype=guide_legend(legend.key.width=unit(3,"line"))) +
-  theme_bw() + theme(legend.position="none", text=element_text(size=14)) + xlab("Previous year's GDD deadline (ordinal day)") + ylab("Critical photoperiod (h)")
+  theme_bw() + theme(legend.position="none", text=element_text(size=12)) + xlab("Previous year's GDD deadline (ordinal day)") + ylab("Critical photoperiod (h)") + ylim(12.3,14.2)
 
@@
   color=guide_legend(override.aes=list(linetype=0)), color=guide_legend(override.aes=list(linetype=0)), linetype=guide_legend(legend.key.width=unit(3,"line"))) +
-  theme_bw() + theme(legend.position=c(.8,.75), text=element_text(size=14)) + xlab("Previous decade's GDD deadline (ordinal day)") + ylab("Critical photoperiod (h)")
+  theme_bw() + theme(legend.position=c(.8,.75), text=element_text(size=12)) + xlab("Previous decade's GDD deadline (ordinal day)") + ylab("Critical photoperiod (h)") + ylim(12.3,14.2)
 
-plot_grid(prevyear_all, previnterval_all, labels="AUTO")
+prev_year_plots<-plot_grid(prevyear_all, previnterval_all, labels="AUTO")
 ```
-```{r}
+
+```{r echo=FALSE, include=FALSE}
 all_prevyear_model<-lm(Estimate~prev_year_gdd400*country*CPP_YEAR, data=prev_year)
@@
 At least 50 male and 50 female adult mosquitoes were used to initiate all populations in the laboratory at Georgetown, except for SEN, which had 37 females and 39 males.
+
+```{r echo=FALSE}
+map_panels_bw
+
+```
 
@@
 
-### Growing degree day models
+### Climate data and growing degree day calculations
+Daily maximum and minimum temperatures for each location were obtained from the Climate Prediction Center of the National Oceanic and Atmospheric Administration (CPC NOAA, https://www.esrl.noaa.gov/psd/).
+These observations are available at 0.5º spatial and daily temporal resolution.
+Data were missing in the CPC dataset for three locations in Japan (Okinawa, Tanegashima, and Shimonoseki); for these we used the European Center’s Medium Range Weather Forecasting/ ERAI reanalysis [@deeERAInterimReanalysisConfiguration2011].
+This reanalysis is available at 0.125º resolution from 1979 – present.
+
+These temperatures were to calculate daily GDD values, with minimum and maximum thresholds of 11º C and 33º C, respectively, as described in [@baskervilleRapidEstimationHeat1969].
+For each year, daily GDD values were accumulated backward beginning from 31 December.
+We then extracted the day of the year when 400 or 465 GDD is reached to obtain the deadline date.
 
@@
 
+[To discuss: could it be that the steep gradient in diapause incidence over a relatively small geographic scale in Florida suggests that it's the result of a gradient of admixture from tropical populations rather than an adaptive gradient?]
+
 ```{r fig.cap="Figure 2B: Diapause incidence by country and year.
@@
 
-country_DI_bw<-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("black","gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3), text=element_text(size=14)) + guides(fill = guide_legend(override.aes=list(shape=22))) + xlab("Latitude") + ylab("Short-day diapause incidence") + stat_smooth(aes(linetype=year, color=country), se=FALSE) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + scale_color_manual(values=c("black","gray"), name="Country")
+country_DI_bw<-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("black","gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3), text=element_text(size=12)) + guides(fill = guide_legend(override.aes=list(shape=22))) + xlab("Latitude") + ylab("Short-day diapause incidence") + stat_smooth(aes(linetype=year, color=country), se=FALSE) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + scale_color_manual(values=c("black","gray"), name="Country")
+
+country_DI_bw
+```
+```{r echo=FALSE, warning=FALSE}
+ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("black","gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3), text=element_text(size=12)) + guides(fill = guide_legend(override.aes=list(shape=22))) + xlab("Latitude") + ylab("Short-day diapause incidence") + geom_smooth(aes(linetype=year, color=country), se=FALSE, method="glm", method.args=list(family="binomial")) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + scale_color_manual(values=c("black","gray"), name="Country")
 
@@
 
+We next examined critical photoperiod among the diapausing fraction of each population.
 CPP values correlated strongly with latitude in the US and Japan (p<0.001, Table 3).
@@
 This implies earlier entry into diapause at the same latitude in Japan than in the US, as well as earlier entry into diapause in the same locations in 2018 than in 2008.
-Both of these differences are contrary to expectation based on global warming trends and the consistently milder winters in Japan compared to US.
+Both of these differences are contrary to expectations based on global warming trends and the consistently milder winters in Japan compared to US.
 
@@
 In 2014, the calculated deadline was 247-252.
-These results confirmed that GDD deadlines generally correspond to field observations and suggested that examining the GDD deadline in years preceding the collection year is a reasonable approach.
+These results confirmed that GDD deadlines broadly correspond to field observations and suggested that examining the GDD deadline in years preceding the collection year is a reasonable approach.
 
@@
 
-```{r}
+```{r echo=FALSE}
 gdd_cpp_plot
@@
 
-```{r}
-kable(rbind(prettify(anova(model1)), prettify(anova(model2)), prettify(anova(model3)), prettify(anova(model4))), digits=3)  %>% pack_rows("2018 CPP vs 2017 GDD Deadline", 1, 3) %>% pack_rows("2008 CPP vs 2007 GDD Deadline", 4, 6) %>% pack_rows("2018 CPP vs 2008-2017 Average GDD Deadline", 7, 9) %>% pack_rows("2008 CPP vs 1998-2007 Average GDD Deadline", 10, 12)
+```{r echo=FALSE}
+kable(rbind(prettify(anova(model1)), prettify(anova(model2)), prettify(anova(model3)), prettify(anova(model4))), digits=3)  
+```
+```{r echo=FALSE}
+prev_year_plots
+```
+
+```{r echo=FALSE}
+kable(prettify(anova(all_prevyear_model))) 
+kable(prettify(anova(all_previnterval_model)))
 
 ```
-The GDD-based estimate of diapause deadline date does not explain the shift to earlier critical photoperiod between 2008 and 2018. Although deadline dates vary from year to year
+
+The GDD-based estimate of diapause deadline date does not explain the shift to earlier critical photoperiod between 2008 and 2018.
+Although deadline dates vary from year to year, there is no discernible directional trend over the 20-year period examined, except for very slight trends
 
@@
 ```
-```{r}
+```{r echo=FALSE}
 gdd_deadline_plots
 ```
-```{r}
+```{r echo=FALSE}
 plot_grid(ggplot(filter(gdd400_jp, only_2008=="n"), aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + theme_bw() + stat_smooth(method=lm, se=FALSE),ggplot(gdd400_us, aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + theme_bw() + stat_smooth(method=lm, se=FALSE), labels="AUTO")
 
+```
+```{r echo=FALSE, warning=FALSE}
+gdd_by_interval<-ggplot(filter(interval_summaries, CPP_YEAR==2018), aes(x=latitude, y=gdd400_mean, fill=country, shape=interval)) + 
+  geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Period") + geom_errorbar(aes(ymin=gdd400_mean-gdd400_sd, ymax=gdd400_mean+gdd400_sd, color=country)) +
+  stat_smooth(method=lm, se=FALSE, aes(linetype=interval, color=country)) + scale_linetype_manual(values=c("dotted","solid"), name="Period") + 
+  scale_color_manual(values=c("black","gray"), name="Country") +
+  guides(fill=guide_legend(override.aes=list(shape=22)), linetype=guide_legend(override.aes=list(color="black")),
+  color=guide_legend(override.aes=list(linetype=0)), color=guide_legend(override.aes=list(linetype=0)), linetype=guide_legend(legend.key.width=unit(3,"line"))) +
+  theme_bw() + theme(legend.position=c(.8,.75), text=element_text(size=12)) + xlab("Latitude") + ylab("Mean GDD Deadline")
+
+gdd_by_interval
 ```

### 2020-02-20 bioclim_data.Rmd
Bioclim

---
title: "R Notebook"
output: html_notebook
---

```{r}
install.packages("raster")
install.packages("sp")
```

```{r}
library("tidyverse")
library("raster")
library("sp")
```

```{r}
#Download worldclim data 
clim = getData('worldclim', var='bio', res=2.5) 
gain(clim) = 0.1
```

```{r}
#look up SpatialPoints 
# https://gis.stackexchange.com/questions/227585/using-r-to-extract-data-from-worldclim
```

### 2020-02-20 MS.Rmd
Bioclim

@@
 cpp_year<-dplyr::select(read.csv("cpp_calcs_filtered.csv"), pop, Estimate, year) %>% filter(year!="1988") %>% rename("CPP_YEAR"="year")
-gdd400_all_new<-left_join(gdd400_all, cpp_year, by="pop")
+gdd400_all_new<-left_join(gdd400_all, cpp_year, by=c("pop"))
 gdd400_all_new$CPP_YEAR<-as.factor(gdd400_all_new$CPP_YEAR)
@@
 We reasoned that the optimal time for entering diapause depended on the amount of time required to complete a full generation under particular climatic conditions.
-Insect phenology studies often estimate generation times using growing degree day (GDD) models.
-Based on a survey of literature of Aedes albopictus growth rates under various temperature conditions, we estimated that a mosquito generation - from the egg stage through ovipostion - required 400-465 GDD.
+Insect phenology studies often estimate generation times using growing degree day (GDD) models, which take advantage of the fact that insect development time is linearly temperature-dependent under moderate climate conditions.
+Based on a survey of literature of Aedes albopictus growth rates under various temperature conditions, we estimated that a mosquito generation - from the egg stage through ovipostion - required 400-465 growing degree-days.
 We used climate records over the past 20 years from our sampling locations to calculate a yearly GDD-based "deadline" for diapause at each location: the date after which eggs laid would not have enough growing degree days left in the year to complete another generation.
@@
 ```{r echo=FALSE}
-kable(rbind(prettify(anova(model1)), prettify(anova(model2)), prettify(anova(model3)), prettify(anova(model4))), digits=3)  
+kable(rbind(prettify(anova(model2)), prettify(anova(model1)), prettify(anova(model4)), prettify(anova(model3))), digits=3)  
 ```
+
 ```{r echo=FALSE}
@@
 ```{r echo=FALSE}
-kable(prettify(anova(all_prevyear_model))) 
-kable(prettify(anova(all_previnterval_model)))
+kable(prettify(anova(all_prevyear_model)), digits=3) 
+kable(prettify(anova(all_previnterval_model)), digits=3)
 
@@
 Based on the distribution of percent differences from the initial counts (Figure S1A), we considered oviposition papers with more than 100% or less than 60% of the initial count to be less reliable data points; these papers made up less than 10% of both individual papers and total eggs quantified in the experiment.
-Excluding these data points resulted in 5 cages out of 286 in the total data set having less than 100 remaining eggs, so they were excluded from the photoperiod response curves.
+Excluding these data points resulted in 5 cages out of 286 in the total data set (1.7%) having less than 100 remaining eggs, so they were excluded from the photoperiod response curves.
 These filtering steps and exclusions did not affect CPP estimates for any population except PBS; for this population, excluding the unreliable counts eliminated an outlier in the photoperiod response curve and brought the CPP estimate into the range of CPP values for the rest of the populations (Figure S2).
@@
 ```
+```{r echo=FALSE, warning=FALSE}
+field_data_plot
+```
+
 ```{r echo=FALSE}
@@
 ```{r echo=FALSE}
-plot_grid(ggplot(filter(gdd400_jp, only_2008=="n"), aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + theme_bw() + stat_smooth(method=lm, se=FALSE),ggplot(gdd400_us, aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + theme_bw() + stat_smooth(method=lm, se=FALSE), labels="AUTO")
+plot_grid(ggplot(filter(gdd400_jp, only_2008=="n"), aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + theme_bw() + stat_smooth(method=lm, se=FALSE) + xlab("Year") + ylab("GDD Deadline"),ggplot(gdd400_us, aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + theme_bw() + stat_smooth(method=lm, se=FALSE) + xlab("Year") + ylab("GDD Deadline"), labels="AUTO")
 
 ```
+
 ```{r echo=FALSE, warning=FALSE}

### 2020-02-20 bioclim_data.Rmd
Bioclim

@@
 clim = getData('worldclim', var='bio', res=2.5) 
-gain(clim) = 0.1
+gain(clim) = 0.1 #values in bioclim dataset are scaled by 10
 ```
@@
 ```{r}
-#look up SpatialPoints 
 # https://gis.stackexchange.com/questions/227585/using-r-to-extract-data-from-worldclim
+
+coords <- data.frame(x=c(-80.2, -80.4, -80.9, -81.8, -81.5, -80, -78.1, -78.4, -77.1, -77.4, -74.9, -74.1, 127.8, 130.6, 130.6, 130.3, 131.5, 135.8, 136, 136.7, 139, 140.9),
+                     y=c(26.67, 27.58, 28.85, 30.33, 31.13, 33.75, 35.03, 36.35, 37.05, 38.63, 39.77, 40.73, 26.2, 31.6, 32.65, 33.25, 34.18, 35, 35.35, 36.56, 37.91, 38.29))
+points <- SpatialPoints(coords)
+values <- extract(clim,points)
+
 ```
+The bioclimate variables used were: minimum temperature of coldest month

### 2020-02-24 MS.Rmd
Better maps

@@
 library("cowplot")
+library("ggrepel")
 ```
@@
 dimodel_result<-data.frame(anova(dimodel))
+#checking for normality of residuals
+dimodel.stdres<-rstandard(dimodel)
+qqnorm(dimodel.stdres, ylab="Standardized Residuals", xlab="Normal Scores", main="Normal Probability Plot")
+qqline(dimodel.stdres)
 ```
@@
 us_map_bw<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="black") + 
-  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 1) + xlim(c(-87, -65)) + ylim(c(24,48)) +
-  geom_label(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) + theme_bw() + xlab("") + ylab("Latitude")
+  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 2) + xlim(c(-87, -65)) + ylim(c(24,48)) +
+  geom_label_repel(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines"), nudge_x=1.5) + theme_bw() + xlab("") + ylab("")
 
 japan_map_bw<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="black") + 
-  coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 1)  + xlim(c(125, 147)) + ylim(c(24,48)) +
-  geom_label_repel(data=filter(pops, country=="JP", only_2008=="n"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) + theme_bw()  + xlab("") + ylab("")
+  coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 2)  + xlim(c(125, 147)) + ylim(c(24,48)) +
+  geom_label_repel(data=filter(pops, country=="JP", only_2008=="n"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines"), nudge=-4) + theme_bw()  + xlab("") + ylab("Latitude")
 

### 2020-02-24 MS.Rmd
Better maps

@@
 ```{r table1, echo=FALSE}
-kable(filter(pops, only_2008=="n")[,1:6], col.names=c("Population", "City", "Latitude", "Longitude", "Country", "Life Stage Sampled"), caption="Table 1: 2018 Sampling Locations")
+kable(filter(pops, only_2008=="n")[,1:7], col.names=c("Population", "City", "Latitude", "Longitude", "Country", "Generation*", "Life Stage Sampled"), caption="Table 1: 2018 Sampling Locations")
 ```
@@
 ```{r echo=FALSE, warning=FALSE}
-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("black","gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3), text=element_text(size=12)) + guides(fill = guide_legend(override.aes=list(shape=22))) + xlab("Latitude") + ylab("Short-day diapause incidence") + geom_smooth(aes(linetype=year, color=country), se=FALSE, method="glm", method.args=list(family="binomial")) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + scale_color_manual(values=c("black","gray"), name="Country")
+di_plot_binomial<-ggplot(filter(DI_short_2, year!=1988), aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("black","gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") + geom_point(size=3) + theme_bw() + theme(legend.position=c(0.9,0.3), text=element_text(size=12)) + guides(fill = guide_legend(override.aes=list(shape=22))) + xlab("Latitude") + ylab("Short-day diapause incidence") + geom_smooth(aes(linetype=year, color=country), se=FALSE, method="glm", method.args=list(family="binomial")) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + scale_color_manual(values=c("black","gray"), name="Country")
 

### 2020-03-31 MS.Rmd
Updated GDD deadline data with correct baseline
GDD figure

@@
 install.packages("stringr")
-install.packages("geosphere")
 install.packages("lubridate")
@@
 install.packages("flextable")
+install.packages("stats")
+install.packages("ggbiplot")
 ```
@@
 library("mapdata")
-library("scatterpie")
 library("ggedit")
@@
 library("ggrepel")
+library("stats")
+library("ggbiplot")
+
 ```
@@
 cpp_2008<-right_join(cpp_2008, popdata_2008, by="pop")
+#1988 data not used in present study but can load it anyway
 cpp_1988<-data.frame(read.csv("CPP_1988.csv")) 
@@
 cpp_1988$StdError<-na_if(cpp_1988$StdError, "NA")
-cpp_2<-rbind(cpp_1988, cpp_2008, cpp_2018) 
-write.csv(cpp_2, "cpp_calcs_filtered.csv")
-```
 
-```{r Japan_US_years, echo=FALSE, include=FALSE}
-cpp_japan_us_years<-ggplot(cpp_2, aes(x=lat, y=Estimate, fill=year, color=year)) + 
-  scale_color_manual(values=c("white","gray","black")) + 
-  scale_fill_manual(values=c("white","gray","black"))+ 
-  geom_point(shape=21, color="black") + 
-  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
-  facet_grid(~country) + stat_smooth(method=lm, se=FALSE)
+#2008 and 2018 data
+cpp_all<-rbind(cpp_2008, cpp_2018) 
+write.csv(cpp_all, "cpp_calcs_filtered.csv")
 ```
@@
 ```{r cppmodel, echo=FALSE, include=FALSE}
-cppmodel<-lm(Estimate~lat*country*year, data=filter(cpp_2, year!=1988))
+cppmodel<-lm(Estimate~lat*country*year, data=cpp_all)
 cppmodel_result<-data.frame(anova(cppmodel))
@@
 ```{r, echo=FALSE, include=FALSE}
-ggplot(cpp_2, aes(x=lat, y=Estimate, color=country)) + 
+#basic breakdown
+ggplot(cpp_all, aes(x=lat, y=Estimate, color=country)) + 
   geom_point() + 
@@
 ```{r, echo=FALSE, include=FALSE}
-cpp_08_18_shifts<-ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, color=country, shape=year)) + 
-  geom_point(size=3) + scale_color_manual(values=c("red", "blue")) +
-  stat_smooth(method=lm, se=FALSE, aes(linetype=year)) + scale_linetype_manual(values=c("dotted","solid")) + theme_bw() + theme(legend.position=c(.8,.25)) + xlab("Latitude") + ylab("Critical photoperiod")
-  # + geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE)
-
-cpp_08_18_shifts_bw<-ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, fill=country, shape=year)) + 
+#MAIN FIGURE VERSIONS
+cpp_08_18_shifts_bw<-ggplot(cpp_all, aes(x=lat, y=Estimate, fill=country, shape=year)) + 
   geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") +
@@
   theme_bw() + theme(legend.position=c(.8,.25), text=element_text(size=12)) + xlab("Latitude") + ylab("Critical photoperiod (h)")
+
+#version with only one shape, gray/black for years
+ggplot(cpp_all, aes(x=lat, y=Estimate, alpha=year, fill=country)) +
+  geom_point(size=3, shape=21, stroke=1) + scale_fill_manual(values=c("white", "black"), name="Country") +
+  scale_alpha_manual(values=c(0.3, 1), name="Year") +
+  stat_smooth(method=lm, se=FALSE, aes(linetype=country, color=year)) + scale_color_manual(values=c("gray", "black"), name="Year") +  scale_linetype_manual(values=c("dotted","solid"), name="Country") +
+  theme_bw() + theme(legend.position=c(.8,.25), text=element_text(size=12)) + xlab("Latitude") + ylab("Critical photoperiod (h)") + guides(linetype=guide_legend(override.aes=list(color="black"), legend.key.width=unit(5,"line")))
+
 ```
@@
 ```{r dimodel, echo=FALSE, include=FALSE}
+#DI by lat/country/year comparison
+dimodel_poly_us<-lm(DI~lat^2 + lat, data=filter(DI_short_2, country=="US"))
+dimodel_lat_us<-lm(DI~lat, data=filter(DI_short_2, country=="US"))
 dimodel<-lm(DI~lat*country*year, data=DI_short_2)
@@
 qqline(dimodel.stdres)
+
+#Spearman rank correlation coefficients
+cor.test(filter(DI_short_2, country=="US", year==2018)$lat, filter(DI_short_2, country=="US", year==2018)$DI, alternative="two.sided", method="spearman")
+cor.test(filter(DI_short_2, country=="JP", year==2018)$lat, filter(DI_short_2, country=="JP", year==2018)$DI, alternative="two.sided", method="spearman")
+
+wilcox.test(filter(DI_short_2, country=="US", year==2018, pop!="PBS")$DI, (filter(DI_short_2, country=="US", year==2008) %>% arrange(lat))$DI, paired=FALSE)
+
 ```
@@
   theme_bw() + theme(axis.title.x=element_blank(), axis.title.y=element_blank())
-DI_data<-data.frame(read.csv("diapause_incidence_8h_2008_2018.csv"))
-DI_data$ND<-1-DI_data$DI
-DI_info<-full_join(DI_data, pops)
-DI_info<-rename(DI_info, Diapause=DI, Nondiapause=ND)
-us_DI_pies<-remove_geom(us_map, "label") + geom_scatterpie(data=filter(DI_info, country=="US"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + facet_grid(~year) + theme(axis.title.x=element_blank(), axis.title.y=element_blank())
-jp_DI_pies<-remove_geom(japan_map, "label") + geom_scatterpie(data=filter(DI_info, country=="JP"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + 
-  facet_grid(~year) + theme(axis.title.x=element_blank(), axis.title.y=element_blank())
-DI_pies_2018<-plot_grid(remove_geom(us_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="US", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "null")),
-
-remove_geom(japan_map, "label") %>% remove_geom("point") + ylim(24,48) + geom_scatterpie(data=filter(DI_info, country=="JP", year=="2018"), aes(x=longitude, y=latitude, r=.7, alpha=0.95), cols=c("Diapause", "Nondiapause")) + scale_fill_manual(values=c("black","white")) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.title.y=element_blank(), legend.position="none", plot.margin = unit(c(0, 0, 0, 0), "null")), ncol=2, align="hv")
 
@@
 ```{r, echo=FALSE, include=FALSE}
-#GDD deadline data
+#GDD 400 deadline data
 #importing and reshaping table
@@
                                                   year<2008 ~ 'y1998-2007'))
-```
 
-```{r, echo=FALSE, include=FALSE}
-#GDD deadline data
-#importing and reshaping table
 gdd465_us <- read.csv("GDD465_US.csv", check.names=FALSE) %>% gather(year, gdd465, "1998":"2018", factor_key=FALSE)
@@
 gdd465_jp$year<-as.numeric(gdd465_jp$year)
+locations<-read.csv("mapping_pops_cpp_expt.csv")
+gdd465_us<-left_join(gdd465_us, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
+gdd465_jp<-left_join(gdd465_jp, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
 gdd465_all<-rbind(gdd465_us, gdd465_jp)
 gdd465_all$pop<-as.factor(gdd465_all$pop)
+gdd465_all<-mutate(gdd465_all, interval=case_when(year<=2017 & year>=2008 ~ 'y2008-2017',
+                                                  year<2008 ~ 'y1998-2007'))
 gdd_all<-left_join(gdd400_all, gdd465_all)
 gdd_all$gdd<-(gdd_all$gdd400 + gdd_all$gdd465)/2
+
 ```
@@
 ```{r, echo=FALSE, include=FALSE}
-gdd_deadline_plots<-ggplot(filter(gdd400_all, only_2008=="n", year!=2018), aes(x=latitude, y=gdd400, fill=country)) + scale_fill_manual(values=c("black", "gray"), name="Country") + geom_point(shape=21) + facet_wrap(~year) + xlab("Latitude") + ylab("GDD Deadline (ordinal day)") + theme_bw() + theme(legend.position="bottom")
+gdd_deadline_plots<-ggplot(filter(gdd465_all, only_2008=="n", year!=2018), aes(x=latitude, y=gdd465, fill=country)) + scale_fill_manual(values=c("black", "gray"), name="Country") + geom_point(shape=21) + facet_wrap(~year) + xlab("Latitude") + ylab("GDD Deadline (ordinal day)") + theme_bw() + theme(legend.position="bottom")
 
@@
 #adding 2008 and 2018 CPP values 
-cpp_year<-dplyr::select(read.csv("cpp_calcs_filtered.csv"), pop, Estimate, year) %>% filter(year!="1988") %>% rename("CPP_YEAR"="year")
-gdd400_all_new<-left_join(gdd400_all, cpp_year, by=c("pop"))
-gdd400_all_new$CPP_YEAR<-as.factor(gdd400_all_new$CPP_YEAR)
+cpp_year<-dplyr::select(read.csv("cpp_calcs_filtered.csv"), pop, Estimate, year) %>% dplyr::rename("CPP_YEAR"="year")
+gdd465_all_new<-left_join(gdd465_all, cpp_year, by=c("pop"))
+gdd465_all_new$CPP_YEAR<-as.factor(gdd465_all_new$CPP_YEAR)
 
@@
 #GDD deadlines averaged across all years
-gdd_summaries<-gdd_all_new %>% filter(interval!="NA") %>% group_by(pop) %>% dplyr::summarise(gdd400_mean=mean(gdd400), gdd400_sd=sd(gdd400), gdd465_mean=mean(gdd465), gdd465_sd=sd(gdd465), latitude=mean(latitude), country=first(country), only_2008=first(only_2008))
+gdd_summaries<-gdd_all_new %>% filter(interval!="NA") %>% group_by(pop) %>% dplyr::summarise(gdd465_mean=mean(gdd465), gdd465_sd=sd(gdd465), gdd465_mean=mean(gdd465), gdd465_sd=sd(gdd465), latitude=mean(latitude), country=first(country), only_2008=first(only_2008))
 
-lat_gdd_average<-ggplot(filter(gdd_summaries, only_2008=="n"), aes(x=latitude, y=gdd400_mean, fill=country)) + geom_point(shape=21) + scale_fill_manual(values=c("black", "gray")) + geom_errorbar(aes(ymin=gdd400_mean-gdd400_sd, ymax=gdd400_mean+gdd400_sd, color=country)) + scale_color_manual(values=c("black","gray")) + theme_bw() + theme(legend.position="bottom") + xlab("Latitude (N)") + ylab("GDD Deadline (ordinal day)")
+lat_gdd_average<-ggplot(filter(gdd_summaries, only_2008=="n"), aes(x=latitude, y=gdd465_mean, fill=country)) + geom_point(shape=21) + scale_fill_manual(values=c("black", "gray")) + geom_errorbar(aes(ymin=gdd465_mean-gdd465_sd, ymax=gdd465_mean+gdd465_sd, color=country)) + scale_color_manual(values=c("black","gray")) + theme_bw() + theme(legend.position="bottom") + xlab("Latitude (N)") + ylab("GDD Deadline (ordinal day)")
 
@@
 #plotting GDD deadlines and CPP
-gdd_08 <- ggplot(filter(gdd400_all_new, year==2007, CPP_YEAR==2008), aes(x=gdd400, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=3, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
+gdd_08 <- ggplot(filter(gdd465_all_new, year==2007, CPP_YEAR==2008), aes(x=gdd465, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=3, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
   theme_bw() + theme(legend.position="none", text=element_text(size=12)) +
@@
 
-gdd_18 <- ggplot(filter(gdd400_all_new, year==2017, CPP_YEAR==2018), aes(x=gdd400, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=3, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
+gdd_18 <- ggplot(filter(gdd465_all_new, year==2017, CPP_YEAR==2018), aes(x=gdd465, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=3, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
 theme_bw() + theme(legend.position=c(.8,.8), text=element_text(size=12)) +
@@
 #plotting CPP by GDD deadlines averaged over intervals
-interval_summaries<-gdd_all_new %>% filter(interval!="NA") %>% group_by(pop, interval, CPP_YEAR) %>% dplyr::summarise(gdd400_mean=mean(gdd400), gdd400_sd=sd(gdd400), cpp=mean(Estimate), latitude=mean(latitude), country=first(country), only_2008=first(only_2008))
+interval_summaries<-gdd_all_new %>% filter(interval!="NA") %>% group_by(pop, interval, CPP_YEAR) %>% dplyr::summarise(gdd465_mean=mean(gdd465), gdd465_sd=sd(gdd465), cpp=mean(Estimate), latitude=mean(latitude), country=first(country), only_2008=first(only_2008))
 
-gdd_cpp_interval_18<-ggplot(filter(interval_summaries, interval=="y2008-2017", CPP_YEAR==2018), aes(x=gdd400_mean, y=cpp, fill=country)) + geom_point(size=3, shape=21) + scale_fill_manual(values=c("black","gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") + theme_bw() + theme(legend.position="none", text=element_text(size=12)) + xlab("Mean GDD deadline, 2008-2017 (ordinal day)") + ylab("Critical photoperiod, 2018 (h)") + ylim(12.3, 14.2)
+gdd_cpp_interval_18<-ggplot(filter(interval_summaries, interval=="y2008-2017", CPP_YEAR==2018), aes(x=gdd465_mean, y=cpp, fill=country)) + geom_point(size=3, shape=21) + scale_fill_manual(values=c("black","gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") + theme_bw() + theme(legend.position="none", text=element_text(size=12)) + xlab("Mean GDD deadline, 2008-2017 (ordinal day)") + ylab("Critical photoperiod, 2018 (h)") + ylim(12.3, 14.2)
 
-gdd_cpp_interval_08<-ggplot(filter(interval_summaries, interval=="y1998-2007", CPP_YEAR==2008), aes(x=gdd400_mean, y=cpp, fill=country)) + geom_point(size=3, shape=21) + scale_fill_manual(values=c("black","gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") + theme_bw() + theme(legend.position="none", text=element_text(size=12)) + xlab("Mean GDD deadline, 1998-2007 (ordinal day)") + ylab("Critical photoperiod, 2008 (h)") + ylim(12.3, 14.2)
+gdd_cpp_interval_08<-ggplot(filter(interval_summaries, interval=="y1998-2007", CPP_YEAR==2008), aes(x=gdd465_mean, y=cpp, fill=country)) + geom_point(size=3, shape=21) + scale_fill_manual(values=c("black","gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") + theme_bw() + theme(legend.position="none", text=element_text(size=12)) + xlab("Mean GDD deadline, 1998-2007 (ordinal day)") + ylab("Critical photoperiod, 2008 (h)") + ylim(12.3, 14.2)
 
@@
 ```{r echo=FALSE, include=FALSE}
-#2015 Diapause Incidence Field Trial
-field_data<-read.csv("DI_2015_FIELD.csv")
-field_data$WeekDate <- reorder(field_data$WeekDate, field_data$Week)
-field_data_plot<-ggplot(filter(field_data, Eggs=="Field"), aes(x=WeekDate, y=DI)) + geom_point(size=3, shape=21, fill="gray") + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-xlab("Week") + ylab("Diapause incidence") + geom_vline(aes(xintercept=10.4), linetype="dotted")
+data_18<-gdd465_all_new %>% filter(CPP_YEAR==2018) %>% filter(year==2017)
+model1<-lm(Estimate~gdd465*country, data=data_18)
 
-field.model.fit <- drm(DI ~ Week, data = filter(field_data, Eggs=="Field"), fct = LL.5())
-summary(field.model.fit)
-# estimate the 50% intercept
-cpp_field<-data.frame(ED(field.model.fit, 50))
+data_08<-gdd465_all_new %>% filter(CPP_YEAR==2008) %>% filter(year==2007)
+model2<-lm(Estimate~gdd465*country, data=data_08)
 
-```
-
-```{r echo=FALSE, include=FALSE}
-data_18<-gdd400_all_new %>% filter(CPP_YEAR==2018) %>% filter(year==2017)
-model1<-lm(Estimate~gdd400*country, data=data_18)
-
-data_08<-gdd400_all_new %>% filter(CPP_YEAR==2008) %>% filter(year==2007)
-model2<-lm(Estimate~gdd400*country, data=data_08)
-
-model3<-lm(cpp~gdd400_mean*country, data=filter(interval_summaries, interval=="y2008-2017", CPP_YEAR==2018))
-model4<-lm(cpp~gdd400_mean*country, data=filter(interval_summaries, interval=="y1998-2007", CPP_YEAR==2008))
+model3<-lm(cpp~gdd465_mean*country, data=filter(interval_summaries, interval=="y2008-2017", CPP_YEAR==2018))
+model4<-lm(cpp~gdd465_mean*country, data=filter(interval_summaries, interval=="y1998-2007", CPP_YEAR==2008))
 
@@
 ```{r echo=FALSE, include=FALSE}
-write.csv(filter(gdd400_all_new, year %in% c(2007,2017)), "gdd_cpp_07_17.csv")
-prev_year<-read.csv("gdd_prevyear.csv")
-prev_year$CPP_YEAR<-as.factor(prev_year$CPP_YEAR)
+#2015 Diapause Incidence Field Trial
+field_data<-read.csv("DI_2015_FIELD.csv")
+field_data$WeekDate <- reorder(field_data$WeekDate, field_data$Week)
 
-write.csv(interval_summaries, "interval_summaries.csv")
-prev_interval<-read.csv("gdd_previnterval.csv")
-prev_interval$CPP_YEAR<-as.factor(prev_interval$CPP_YEAR)
+#field eggs only
+field.model.fit <- drm(DI ~ Week, data = dc_field_eggs, fct = LL.5())
+summary(field.model.fit)
+# estimate the 50% intercept
+cpp_field<-data.frame(ED(field.model.fit, 50))
 
-```
+field_data_plot<-ggplot(field_data, aes(x=WeekDate, y=DI)) + geom_point(size=3, shape=21, fill="gray") + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+xlab("Week") + ylab("Diapause incidence") + geom_vline(aes(xintercept=cpp_field$Estimate), linetype="dotted")
 
-```{r echo=FALSE, include=FALSE}
-prevyear_all<-ggplot(prev_year, aes(x=prev_year_gdd400, y=Estimate, fill=country, shape=CPP_YEAR)) + 
-  geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") +
-  stat_smooth(method=lm, se=FALSE, aes(linetype=CPP_YEAR, color=country)) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + 
-  scale_color_manual(values=c("black","gray"), name="Country") +
-  guides(fill=guide_legend(override.aes=list(shape=22)), linetype=guide_legend(override.aes=list(color="black")),
-  color=guide_legend(override.aes=list(linetype=0)), color=guide_legend(override.aes=list(linetype=0)), linetype=guide_legend(legend.key.width=unit(3,"line"))) +
-  theme_bw() + theme(legend.position="none", text=element_text(size=12)) + xlab("Previous year's GDD deadline (ordinal day)") + ylab("Critical photoperiod (h)") + ylim(12.3,14.2)
+#2014 GDD diagram
+man2014<-read.csv("2014_Mean_Temp_MAN.csv") %>% gather(degreetype, degrees, "CountDegrees":"UncountDegrees")
 
-previnterval_all<-ggplot(prev_interval, aes(x=gdd400_mean_previnterval, y=cpp, fill=country, shape=CPP_YEAR)) + 
-  geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") +
-  stat_smooth(method=lm, se=FALSE, aes(linetype=CPP_YEAR, color=country)) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + 
-  scale_color_manual(values=c("black","gray"), name="Country") +
-  guides(fill=guide_legend(override.aes=list(shape=22)), linetype=guide_legend(override.aes=list(color="black")),
-  color=guide_legend(override.aes=list(linetype=0)), color=guide_legend(override.aes=list(linetype=0)), linetype=guide_legend(legend.key.width=unit(3,"line"))) +
-  theme_bw() + theme(legend.position=c(.8,.75), text=element_text(size=12)) + xlab("Previous decade's GDD deadline (ordinal day)") + ylab("Critical photoperiod (h)") + ylim(12.3,14.2)
+man2014_gdd_plot<-ggplot(filter(man2014, DOY_2014>187), aes(x=DOY_2014, y=degrees, fill=degreetype)) + geom_bar(stat="identity", width=1) + 
+  scale_fill_manual(values=c("dark gray","light gray")) + 
+  theme_bw() + 
+  geom_hline(yintercept=10, linetype="dotted") + geom_vline(xintercept=251) + theme(legend.position="none") +
+  ylab("Mean Temp (C)")
 
-prev_year_plots<-plot_grid(prevyear_all, previnterval_all, labels="AUTO")
-```
-
-```{r echo=FALSE, include=FALSE}
-all_prevyear_model<-lm(Estimate~prev_year_gdd400*country*CPP_YEAR, data=prev_year)
-all_previnterval_model<-lm(cpp~gdd400_mean_previnterval*country*CPP_YEAR, data=prev_interval)
 ```
@@
 
+```{r}
+summary(cppmodel)
+```
+
 We reasoned that the optimal time for entering diapause depended on the amount of time required to complete a full generation under particular climatic conditions.
@@
 kable(rbind(prettify(anova(model2)), prettify(anova(model1)), prettify(anova(model4)), prettify(anova(model3))), digits=3)  
-```
-
-```{r echo=FALSE}
-prev_year_plots
-```
-
-```{r echo=FALSE}
-kable(prettify(anova(all_prevyear_model)), digits=3) 
-kable(prettify(anova(all_previnterval_model)), digits=3)
-
 ```
@@
 ```
-```{r echo=FALSE}
-plot_grid(ggplot(filter(gdd400_jp, only_2008=="n"), aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + theme_bw() + stat_smooth(method=lm, se=FALSE) + xlab("Year") + ylab("GDD Deadline"),ggplot(gdd400_us, aes(x=year, y=gdd400, color=pop)) + geom_point() + geom_line() + theme_bw() + stat_smooth(method=lm, se=FALSE) + xlab("Year") + ylab("GDD Deadline"), labels="AUTO")
+```{r echo=FALSE, include=FALSE}
+plot_grid(ggplot(filter(gdd465_jp, only_2008=="n"), aes(x=year, y=gdd465, color=pop)) + geom_point() + geom_line() + theme_bw() + stat_smooth(method=lm, se=FALSE) + xlab("Year") + ylab("GDD Deadline"),ggplot(gdd465_us, aes(x=year, y=gdd465, color=pop)) + geom_point() + geom_line() + theme_bw() + stat_smooth(method=lm, se=FALSE) + xlab("Year") + ylab("GDD Deadline"), labels="AUTO")
 
@@
 ```{r echo=FALSE, warning=FALSE}
-gdd_by_interval<-ggplot(filter(interval_summaries, CPP_YEAR==2018), aes(x=latitude, y=gdd400_mean, fill=country, shape=interval)) + 
-  geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Period") + geom_errorbar(aes(ymin=gdd400_mean-gdd400_sd, ymax=gdd400_mean+gdd400_sd, color=country)) +
+gdd_by_interval<-ggplot(filter(interval_summaries, CPP_YEAR==2018), aes(x=latitude, y=gdd465_mean, fill=country, shape=interval)) + 
+  geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Period") + geom_errorbar(aes(ymin=gdd465_mean-gdd465_sd, ymax=gdd465_mean+gdd465_sd, color=country)) +
   stat_smooth(method=lm, se=FALSE, aes(linetype=interval, color=country)) + scale_linetype_manual(values=c("dotted","solid"), name="Period") + 
@@
 ```
+```{r}
+annualmeantemp_US<-read.csv("AnnualMeanTemp_US.csv", check.names=FALSE) %>% gather(year, annualmeantemp, "1998":"2018", factor_key=FALSE)
+annualmeantemp_US$year<-as.numeric(annualmeantemp_US$year)
+#ggplot(annualmeantemp_US, aes(x=year, y=annualmeantemp, color=pop)) + geom_point() + geom_line()
+
+annualmeantemp_JP<-read.csv("AnnualMeanTemp_JP.csv", check.names=FALSE) %>% gather(year, annualmeantemp, "1998":"2018", factor_key=FALSE)
+annualmeantemp_JP$year<-as.numeric(annualmeantemp_JP$year)
+#ggplot(annualmeantemp_JP, aes(x=year, y=annualmeantemp, color=pop)) + geom_point() + geom_line()
+
+plot_grid(ggplot(annualmeantemp_US, aes(x=year, y=annualmeantemp, color=pop)) + geom_point() + geom_line() + stat_smooth(method=lm, se=FALSE), ggplot(annualmeantemp_JP, aes(x=year, y=annualmeantemp, color=pop)) + geom_point() + geom_line() + stat_smooth(method=lm, se=FALSE))
+```
 
 ```{r figs_for_poster, include=FALSE, echo=FALSE}
-ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, color=country, shape=year)) + 
+ggplot(cpp_all, aes(x=lat, y=Estimate, color=country, shape=year)) + 
      geom_point(size=6) + scale_color_manual(values=c("red", "blue")) +
@@
 
+```{r}
+ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, fill=country, shape=year)) + 
+  geom_point(size=5) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") +
+  stat_smooth(method=lm, se=FALSE, aes(linetype=year, color=country)) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + 
+  scale_color_manual(values=c("black","gray"), name="Country") +
+  guides(fill=guide_legend(override.aes=list(shape=22)), linetype=guide_legend(override.aes=list(color="black")),
+  color=guide_legend(override.aes=list(linetype=0)), color=guide_legend(override.aes=list(linetype=0)), linetype=guide_legend(legend.key.width=unit(3,"line"))) +
+  theme(legend.position=c(.8,.25), text=element_text(size=16),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
+  xlab("Latitude") + ylab("Critical photoperiod (h)")
+
+ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, fill=country, shape=year)) + 
+  geom_point(size=5) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,22), name="Year") +
+  stat_smooth(method=lm, se=FALSE, aes(linetype=year, color=country)) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + 
+  scale_color_manual(values=c("black","gray"), name="Country") +
+  guides(fill=guide_legend(override.aes=list(shape=23)), linetype=guide_legend(override.aes=list(color="black"), legend.key.width=unit(3,"line")),
+  color=guide_legend(override.aes=list(linetype=0))) +
+  theme(legend.position=c(.8,.25), text=element_text(size=16),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
+  xlab("Latitude") + ylab("Critical photoperiod (h)")
+
+ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, fill=country, shape=year)) + 
+  geom_point(size=5) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(21,24), name="Year") +
+  stat_smooth(method=lm, se=FALSE, aes(linetype=year, color=country)) + scale_linetype_manual(values=c("dotted","solid"), name="Year") + 
+  scale_color_manual(values=c("black","gray"), name="Country") +
+  guides(fill=guide_legend(override.aes=list(shape=22)), linetype=guide_legend(override.aes=list(color="black")),
+  color=guide_legend(override.aes=list(linetype=0)), color=guide_legend(override.aes=list(linetype=0)), linetype=guide_legend(legend.key.width=unit(3,"line"))) +
+  theme(legend.position=c(.8,.25), text=element_text(size=16),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
+  xlab("Latitude") + ylab("Critical photoperiod (h)")
+
+ggplot(filter(cpp_2, year!=1988), aes(x=lat, y=Estimate, fill=year, shape=country)) + 
+  geom_point(size=5) + scale_fill_manual(values=c("gray", "black"), name="Year") + scale_shape_manual(values=c(21,22), name="Country") +
+  stat_smooth(method=lm, se=FALSE, aes(linetype=country, color=year)) + scale_linetype_manual(values=c("dotted","solid"), name="Country") + 
+  scale_color_manual(values=c("gray","black"), name="year") +
+  guides(fill=guide_legend(override.aes=list(shape=22)), linetype=guide_legend(override.aes=list(color="black")),
+  color=guide_legend(override.aes=list(linetype=0)), color=guide_legend(override.aes=list(linetype=0)), linetype=guide_legend(legend.key.width=unit(3,"line"))) +
+  theme(legend.position=c(.8,.25), text=element_text(size=16),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
+  xlab("Latitude") + ylab("Critical photoperiod (h)")
+
+#version with only one shape, gray/black for years
+ggplot(cpp_all, aes(x=lat, y=Estimate, alpha=year, fill=country)) +
+  geom_point(size=3, shape=21, stroke=1) + scale_fill_manual(values=c("white", "black"), name="Country") +
+  scale_alpha_manual(values=c(0.3, 1), name="Year") +
+  stat_smooth(method=lm, se=FALSE, aes(linetype=country, color=year), size=1.5) + scale_color_manual(values=c("gray", "black"), name="Year") +  scale_linetype_manual(values=c("longdash","solid"), name="Country") +
+  theme(legend.position=c(.8,.25), text=element_text(size=16),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
+  xlab("Latitude") + ylab("Critical photoperiod (h)") + 
+  guides(linetype=guide_legend(override.aes=list(color="black"), legend.key.width=unit(5,"line")))
+
+ggplot(cpp_all, aes(x=lat, y=Estimate, alpha=year, color=country)) +
+  geom_point(size=3) + scale_color_manual(values=c("blue", "black"), name="Country") +
+  scale_alpha_manual(values=c(0.3, 1), name="Year") +
+  stat_smooth(method=lm, se=FALSE, aes(linetype=year, color=country), size=2) + scale_linetype_manual(values=c("longdash","solid"), name="Country") + 
+  theme(legend.position=c(.8,.25), text=element_text(size=16),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
+  xlab("Latitude") + ylab("Critical photoperiod (h)") + 
+  guides(linetype=guide_legend(override.aes=list(color="black"), legend.key.width=unit(5,"line")))
+
+```
+
 # References

### 2020-04-06 _albopictus_cpp_MS.Rmd
analysis and files for Apr1 draft

---
title: "Ae. albopictus CPP manuscript"
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
output:
  html_document: default
  word_document: default
  pdf_document: default
bibliography: MyLibrary.bib
---

```{r eval=FALSE, message=FALSE, include=FALSE}
install.packages("tidyverse") #1.2.1
install.packages("drc") #3.0.1
install.packages("stringr")
install.packages("lubridate")
install.packages(c("maps", "mapdata"))
devtools::install_github("dkahle/ggmap")
install.packages("scatterpie")
install.packages("ggedit")
install.packages("ggrepel")
install.packages("papeR")
install.packages("cowplot")
install.packages("flextable")
install.packages("stats")
install.packages("ggbiplot")
```

```{r message=FALSE, echo=FALSE, include=FALSE}
library("papeR")
library("flextable")
library("drc") #3.0.1
library("knitr") #1.25
library("tidyverse") #1.2.1
library("stringr") #1.4.0
library("geosphere") #1.5.10
library("lubridate") #1.7.4
library("maps")
library("mapdata")
library("ggedit")
library("ggrepel")
library("cowplot")
library("ggrepel")
library("stats")
library("ggbiplot")

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300)
```

```{r datasheet, echo=FALSE, include=FALSE}
all_data<-data.frame(read.csv("CPP_experiment_data_collated.csv"))
#Filter out empty/discarded oviposition papers 
blockAB<-dplyr::filter(all_data, TotalLarvaeEmbryos>0)
```

```{r yields, echo=FALSE, include=FALSE}
#Calculate yield/fraction of initial count
blockAB$yield <- blockAB$TotalLarvaeEmbryos/blockAB$InitialEggs 
hist(blockAB$yield, main="Yield (final/initial count) per oviposition paper")
plot(blockAB$yield ~ blockAB$TotalLarvaeEmbryos)
```

```{r filter papers, echo=FALSE, include=FALSE}
#removing papers with <60% or >100% initial count
blockAB_unfiltered<-blockAB
blockAB_filtered <- filter(blockAB, yield<=1) %>% filter(yield>=0.6) 
```

```{r summarize, echo=FALSE, include=FALSE}
DI_summarize <- function(blockAB) {
DI_summary <- blockAB %>% 
  group_by(CageID) %>% 
  dplyr::summarise(pop=first(Population), pp=mean(Photoperiod), lat=mean(Lat), block=first(Block), country=first(Country), h1=sum(H1Count), h2=sum(H2Count), emb=sum(EmbryoCount)) %>%   filter(h1+h2+emb>100)
DI_summary$total<-DI_summary$h1+DI_summary$h2+DI_summary$emb
DI_summary$DI<-DI_summary$emb/DI_summary$total
return(DI_summary)
}
```

```{r, echo=FALSE, include=FALSE}
DI_summary<-DI_summarize(blockAB_filtered)
```

```{r totals_plot, echo=FALSE, include=FALSE}
hist(DI_summary$total, main="Total number of viable eggs per cage")
```

```{r DI_plot, echo=FALSE, include=FALSE}
#FIG S2
#plot with DI at 8 h shown as TWO DIFFERENT points (corresponding to two cabinets)
c("PBS", "FEM", "OAK", "JAC", "BRU", "ZIO", "FAY", "NVA", "WAV", "MAN", "BER", "NEW", "OKI", "TAN", "KAG", "YAT", "SAG", "SHI", "YAM", "HIR", "KYO", "TAK", "TOK", "UTS", "KAN", "KHO", "AIZ", "NIG", "SEN", "SAK")
DI_summary$pop<-factor(DI_summary$pop, levels=c("PBS", "FEM", "OAK", "JAC", "BRU", "ZIO", "FAY", "NVA", "WAV", "MAN", "BER", "NEW", "OKI", "TAN", "KAG", "YAT", "SAG", "SHI", "YAM", "HIR", "KYO", "TAK", "TOK", "UTS", "KAN", "KHO", "AIZ", "NIG", "SEN", "SAK"))
DI_plot<-ggplot(DI_summary, aes(x=pp, y=DI, group=pop, fill=country)) + geom_point(shape=21) + facet_wrap(~fct_reorder(pop, lat)) + theme_bw() + theme(legend.position="NULL", text=element_text(size=18)) + 
xlab("Photoperiod (hours)") + ylab("Dipause incidence") + scale_fill_manual(values=c("black","gray")) 
```

```{r CPP_calc_2018, echo=FALSE, include=FALSE}
#From Kevin Emerson
# the following fits the models of the given  using data from the
# data.frame d separately for the various treatments using a 5 parameter
# logistic model
d.model.fit <- drm(DI ~ pp, data = DI_summary, curveid = pop, fct = LL.5())
#plot(d.model.fit, log = "")
# output the parameters of the model fits:
summary(d.model.fit)
# estimate the 50% intercept
ED(d.model.fit, 50)
cpp_values<-data.frame(ED(d.model.fit, 50))

pm<-drc:::predict.drc(d.model.fit, type="response")
```

```{r CPP_calc_2008, echo=FALSE, include=FALSE}
DI_summary_2008<-data.frame(read.csv("CPP_2008.csv"))
d.model.fit.2008 <- drm(DI ~ photoperiod, data = DI_summary_2008, curveid = pop, fct = LL.5())

#plot(d.model.fit.2008, log = "")
# output the parameters of the model fits:
summary(d.model.fit)
# estimate the 50% intercept
ED(d.model.fit.2008, 50)
cpp_values_2008<-data.frame(ED(d.model.fit.2008, 50))
```

```{r collate_cpp_values, echo=FALSE, include=FALSE}
cpp_2018<-data.frame(ED(d.model.fit, 50))
cpp_2018<-rownames_to_column(cpp_2018, var="pop") 
cpp_2018$pop<-str_sub(cpp_2018$pop,3,5)
popdata_2018<-DI_summary %>% group_by(pop) %>% 
  dplyr::summarize(lat=first(lat), country=first(country))
cpp_2018<-right_join(cpp_2018, popdata_2018, by="pop")
cpp_2018$year<-rep("2018", nrow(cpp_2018))
colnames(cpp_2018)[colnames(cpp_2018)=="Std..Error"] <- "StdError"
cpp_2008<-data.frame(ED(d.model.fit.2008, 50)) %>% rownames_to_column(var="pop")
cpp_2008$pop<-str_sub(cpp_2008$pop,3,5)
colnames(cpp_2008)[colnames(cpp_2008)=="Std..Error"] <- "StdError"
popdata_2008<-DI_summary_2008 %>% group_by(pop) %>% 
  dplyr::summarize(country=first(country), lat=first(lat), year=first(year)) 
cpp_2008<-right_join(cpp_2008, popdata_2008, by="pop")
#1988 data not used in present study but can load it anyway
cpp_1988<-data.frame(read.csv("CPP_1988.csv")) 
cpp_1988$pop<-na_if(cpp_1988$pop, "NA")
cpp_1988$StdError<-na_if(cpp_1988$StdError, "NA")

#2008 and 2018 data
cpp_all<-rbind(cpp_2008, cpp_2018) 
write.csv(cpp_all, "cpp_calcs_filtered.csv")
```

```{r}
#FIG S1
yield_plot<-ggplot(blockAB, aes(x=TotalLarvaeEmbryos, y=yield)) + geom_point(alpha=0.5) + xlab("Number of viable eggs\n on oviposition paper") + ylab("Viable eggs/initial egg count on paper") + theme_bw() + theme(text=element_text(size=18))

cpp_filtered<-read.csv("cpp_calcs_filtered.csv") %>% dplyr::select(pop, Estimate, country, lat, year) %>% filter(year==2018)
cpp_unfiltered<-read.csv("cpp_calcs_unfiltered.csv") %>% dplyr::select(pop, Estimate, country, lat, year) %>% filter(year==2018)
names(cpp_filtered)[2]<-"CPP_est_filt"
names(cpp_unfiltered)[2]<-"CPP_est_unfilt"
cpp_compare<-left_join(cpp_filtered,cpp_unfiltered,by="pop")

filter_compare_plot<-ggplot(cpp_compare, aes(x=CPP_est_unfilt, y=CPP_est_filt)) + geom_point() + geom_abline(slope=1, intercept=0) + xlab("CPP without filtering criteria") + ylab("CPP calculated with filtering criteria") + theme_bw() + theme(text=element_text(size=18))

quality_control_fig<-plot_grid(yield_plot, filter_compare_plot, labels="AUTO")
```

```{r cppmodel, echo=FALSE, include=FALSE}
cppmodel<-lm(Estimate~lat*country*year, data=cpp_all)
cppmodel_result<-data.frame(anova(cppmodel))
```

```{r, echo=FALSE, include=FALSE}
#basic breakdown
ggplot(cpp_all, aes(x=lat, y=Estimate, color=country)) + 
  geom_point() + 
  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
  facet_grid(~year) + stat_smooth(method=lm, se=FALSE)
```

```{r, echo=FALSE, include=FALSE}
#FIG 2: CPP by latitude, year and country
#black=Japan, gray=US / circle=2018, square=2008 / solid line=2018, dashed line=2008
cpp_08_18_shifts_bw <- ggplot(cpp_all, aes(x=lat, y=Estimate, fill=country, shape=year)) +
  geom_point(size=4, stroke=1) + scale_fill_manual(values=c("black", "gray"), name="Country") +
  scale_shape_manual(values=c(22,21), name="Year") +
  stat_smooth(method=lm, se=FALSE, aes(linetype=year, color=country), size=1.5) + scale_color_manual(values=c("black", "gray"), name="Country") +  scale_linetype_manual(values=c("longdash","solid"), name="Year") + 
  theme(legend.position=c(.8,.25), text=element_text(size=24),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlab("Latitude") + ylab("Critical photoperiod (h)") + 
  guides(linetype=guide_legend(override.aes=list(color="black"), legend.key.width=unit(5,"line")))

```

```{r incidence_lat_year, echo=FALSE, include=FALSE}
DI_short_2008<-filter(DI_summary_2008, photoperiod==8)
DI_short_2018<-filter(DI_summary, pp==8) %>% group_by(pop) %>% 
  dplyr::summarize(country=first(country), lat=first(lat), photoperiod=first(pp), emb=sum(emb), total=sum(total))
DI_short_2018$DI<-DI_short_2018$emb/DI_short_2018$total
DI_short_2018<-DI_short_2018[,c(1:4,7)] 
DI_short_2018$year<-rep("2018",nrow(DI_short_2018))
DI_short<-rbind(DI_short_2008, DI_short_2018)
write_csv(DI_short, "diapause_incidence_8h_2008_2018.csv")
```

```{r dimodel, echo=FALSE, include=FALSE}
#DI by lat/country/year comparison
dimodel_poly_us<-lm(DI~lat^2 + lat, data=filter(DI_short, country=="US"))
dimodel_lat_us<-lm(DI~lat, data=filter(DI_short, country=="US"))
dimodel<-lm(DI~lat*country*year, data=DI_short)
dimodel_result<-data.frame(anova(dimodel))
#checking for normality of residuals
dimodel.stdres<-rstandard(dimodel)
qqnorm(dimodel.stdres, ylab="Standardized Residuals", xlab="Normal Scores", main="Normal Probability Plot")
qqline(dimodel.stdres)

#Spearman rank correlation coefficients
cor.test(filter(DI_short_2, country=="US", year==2018)$lat, filter(DI_short, country=="US", year==2018)$DI, alternative="two.sided", method="spearman")
cor.test(filter(DI_short_2, country=="JP", year==2018)$lat, filter(DI_short, country=="JP", year==2018)$DI, alternative="two.sided", method="spearman")

#comparison of latitude-matched DIs in each country between 2008 and 2018
#exact latitude match in US; matched by rank in JP
wilcox.test((filter(DI_short, country=="US", year==2018, pop!="PBS") %>% arrange(lat))$DI, (filter(DI_short, country=="US", year==2008) %>% arrange(lat))$DI, paired=TRUE)
wilcox.test((filter(DI_short, country=="JP", year==2018) %>% arrange(lat))$DI, (filter(DI_short, country=="JP", year==2008) %>% arrange(lat))$DI, paired=TRUE)

#FIG S3
#black=Japan, gray=US / circle=2018, square=2008 / solid line=2018, dashed line=2008
country_DI_bw<-ggplot(DI_short, aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("gray","black"), name="Country") + scale_shape_manual(values=c(22,21), name="Year") + geom_point(size=3) +   theme(legend.position=c(.8,.25), text=element_text(size=18),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
xlab("Latitude") + ylab("Short-day diapause incidence") + stat_smooth(aes(linetype=year, color=country), se=FALSE) + scale_linetype_manual(values=c("longdash","solid"), name="Year") + scale_color_manual(values=c("gray","black"), name="Country")

```

```{r maps, echo=FALSE, include=FALSE}
usa<-map_data("usa")
jp<-map_data("japan")
pops<-data.frame(read.csv("mapping_pops_cpp_expt.csv"))
us_map<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="blue") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 1) + xlim(c(-85, -65)) +
  geom_label(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
  theme_bw() 
japan_map<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="red") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 1)  + xlim(c(125, 147)) +
  geom_label_repel(data=filter(pops, country=="JP", only_2008=="n"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
  theme_bw() + theme(axis.title.x=element_blank(), axis.title.y=element_blank())

#bw locations map
us_map_bw<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="black") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 2) + xlim(c(-87, -65)) + ylim(c(24,48)) +
  geom_label_repel(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines"), nudge_x=1.5, size=4) + theme_bw() + theme(text=element_text(size=24), plot.margin = unit(c(0,0,0,0), "cm")) + xlab("") + ylab("")

japan_map_bw<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="black") + 
  coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 2)  + xlim(c(125, 147)) + ylim(c(24,48)) +
  geom_label_repel(data=filter(pops, country=="JP", only_2008=="n"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines"), size=4) + theme_bw() + theme(text=element_text(size=24), plot.margin = unit(c(0,0,0,0), "cm")) + xlab("") + ylab("Latitude")

map_panels_bw<-plot_grid(japan_map_bw, us_map_bw)
```

```{r}
#date of first first
#first frost data
#importing and reshaping table
ff_us <- read.csv("FF_US.csv", check.names=FALSE) %>% gather(year, DFF, "1998":"2018", factor_key=FALSE)
ff_us$year<-as.numeric(ff_us$year)
ff_jp <- read.csv("FF_JP.csv", check.names=FALSE) %>% gather(year, DFF, "1998":"2018", factor_key=FALSE)
ff_jp$year<-as.numeric(ff_jp$year)
#adding latitudes
locations<-read.csv("mapping_pops_cpp_expt.csv")
ff_us<-left_join(ff_us, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
ff_jp<-left_join(ff_jp, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
ff_all<-rbind(ff_us, ff_jp)
ff_all$pop<-as.factor(ff_all$pop)
ff_all<-mutate(filter(ff_all, year!=2018), interval=case_when(year<=2017 & year>=2008 ~ 'y2008-2017',
                                                  year<2008 ~ 'y1998-2007'))

ff_interval_summaries<-ff_all %>% group_by(pop, interval) %>% dplyr::summarise(dff_mean=mean(DFF), dff_sd=sd(DFF), latitude=mean(latitude), country=first(country), only_2008=first(only_2008))

ff_all_summary<-ff_all %>% group_by(pop) %>% dplyr::summarise(dff_mean=mean(DFF), dff_sd=sd(DFF), latitude=mean(latitude), country=first(country), only_2008=first(only_2008))

#FIG 1B
#black=Japan, gray=US 
DFF_decade<-ggplot(filter(ff_interval_summaries, only_2008=="n", interval=="y2008-2017"), aes(x=latitude, y=dff_mean, fill=country)) + geom_point(size=3, shape=21, stroke=1) + scale_fill_manual(values=c("black", "gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") +
  theme(legend.position=c(.25,.25), text=element_text(size=22),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +  scale_linetype_manual(values=c("dotted","solid"), name="Country") + geom_errorbar(aes(ymin=dff_mean-dff_sd, ymax=dff_mean+dff_sd)) + xlab("Latitude") + ylab("Date of first frost (ordinal day)")

```

```{r}
locations_dff<-plot_grid(map_panels_bw, DFF_decade, nrow=2, labels="AUTO")
locations_dff
```

```{r, echo=FALSE, include=FALSE}
#GDD 400 deadline data
#importing and reshaping table
gdd400_us <- read.csv("GDD400_US.csv", check.names=FALSE) %>% gather(year, gdd400, "1998":"2018", factor_key=FALSE)
gdd400_us$year<-as.numeric(gdd400_us$year)
gdd400_jp <- read.csv("GDD400_JP.csv", check.names=FALSE) %>% gather(year, gdd400, "1998":"2018", factor_key=FALSE)
gdd400_jp$year<-as.numeric(gdd400_jp$year)
#adding latitudes
locations<-read.csv("mapping_pops_cpp_expt.csv")
gdd400_us<-left_join(gdd400_us, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
gdd400_jp<-left_join(gdd400_jp, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
gdd400_all<-rbind(gdd400_us, gdd400_jp)
gdd400_all$pop<-as.factor(gdd400_all$pop)
gdd400_all<-mutate(gdd400_all, interval=case_when(year<=2017 & year>=2008 ~ 'y2008-2017',
                                                  year<2008 ~ 'y1998-2007'))

gdd465_us <- read.csv("GDD465_US.csv", check.names=FALSE) %>% gather(year, gdd465, "1998":"2018", factor_key=FALSE)
gdd465_us$year<-as.numeric(gdd465_us$year)
gdd465_jp <- read.csv("GDD465_JP.csv", check.names=FALSE) %>% gather(year, gdd465, "1998":"2018", factor_key=FALSE)
gdd465_jp$year<-as.numeric(gdd465_jp$year)
locations<-read.csv("mapping_pops_cpp_expt.csv")
gdd465_us<-left_join(gdd465_us, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
gdd465_jp<-left_join(gdd465_jp, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
gdd465_all<-rbind(gdd465_us, gdd465_jp)
gdd465_all$pop<-as.factor(gdd465_all$pop)
gdd465_all<-mutate(gdd465_all, interval=case_when(year<=2017 & year>=2008 ~ 'y2008-2017',
                                                  year<2008 ~ 'y1998-2007'))
gdd_all<-left_join(gdd400_all, gdd465_all)
gdd_all$gdd<-(gdd_all$gdd400 + gdd_all$gdd465)/2

```

```{r, echo=FALSE, include=FALSE}
#FIG S4
gdd_deadline_plots<-ggplot(filter(gdd465_all, only_2008=="n", year!=2018), aes(x=latitude, y=gdd465, fill=country)) + scale_fill_manual(values=c("black", "gray"), name="Country") + geom_point(shape=21, size=3) + facet_wrap(~year) + xlab("Latitude") + ylab("GDD Deadline (ordinal day)") + theme_bw() + 
theme(legend.position="bottom", text=element_text(size=18)) 
```

```{r echo=FALSE, include=FALSE}
#adding 2008 and 2018 CPP values 
cpp_year<-dplyr::select(read.csv("cpp_calcs_filtered.csv"), pop, Estimate, year) %>% dplyr::rename("CPP_YEAR"="year")
gdd465_all_new<-left_join(gdd465_all, cpp_year, by=c("pop"))
gdd465_all_new$CPP_YEAR<-as.factor(gdd465_all_new$CPP_YEAR)

gdd_all_new<-left_join(gdd_all, cpp_year, by="pop")
gdd_all_new$CPP_YEAR<-as.factor(gdd_all_new$CPP_YEAR)

```

```{r echo=FALSE, include=FALSE}
#GDD deadlines averaged across all years
gdd_summaries<-gdd_all_new %>% filter(interval!="NA") %>% group_by(pop) %>% dplyr::summarise(gdd465_mean=mean(gdd465), gdd465_sd=sd(gdd465), gdd465_mean=mean(gdd465), gdd465_sd=sd(gdd465), latitude=mean(latitude), country=first(country), only_2008=first(only_2008))

lat_gdd_average<-ggplot(filter(gdd_summaries, only_2008=="n"), aes(x=latitude, y=gdd465_mean, fill=country)) + geom_point(shape=21) + scale_fill_manual(values=c("black", "gray")) + geom_errorbar(aes(ymin=gdd465_mean-gdd465_sd, ymax=gdd465_mean+gdd465_sd, color=country)) + scale_color_manual(values=c("black","gray")) + theme_bw() + theme(legend.position="bottom") + xlab("Latitude (N)") + ylab("GDD Deadline (ordinal day)")

```

```{r echo=FALSE, include=FALSE}
#plotting GDD deadlines and CPP
#black=Japan, gray=US / circle=2018, square=2008 / solid line=2018, dashed line=2008

gdd_08 <- ggplot(filter(gdd465_all_new, year==2007, CPP_YEAR==2008), aes(x=gdd465, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=4, stroke=1, shape=22) + stat_smooth(method=lm, se=FALSE, aes(color=country), linetype="longdash") + scale_color_manual(values=c("black","gray"), name="Country") +
  theme(legend.position=c(.25,.25), text=element_text(size=22),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
    xlab("GDD Deadline 2007 (ordinal day)") + ylab("CPP (h), 2008") + ylim(12.3,14.2)

gdd_18 <- ggplot(filter(gdd465_all_new, year==2017, CPP_YEAR==2018), aes(x=gdd465, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=4, stroke=1, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
  theme(legend.position=c(.25,.25), text=element_text(size=22),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
xlab("GDD Deadline 2017 (ordinal day)") + ylab("CPP (h), 2018") + ylim(12.3, 14.2)

```

```{r echo=FALSE, include=FALSE}
#plotting CPP by GDD deadlines averaged over intervals
#black=Japan, gray=US / circle=2018, square=2008 / solid line=2018, dashed line=2008

interval_summaries<-gdd_all_new %>% filter(interval!="NA") %>% group_by(pop, interval, CPP_YEAR) %>% dplyr::summarise(gdd465_mean=mean(gdd465), gdd465_sd=sd(gdd465), cpp=mean(Estimate), latitude=mean(latitude), country=first(country), only_2008=first(only_2008))

gdd_cpp_interval_18<-ggplot(filter(interval_summaries, interval=="y2008-2017", CPP_YEAR==2018), aes(x=gdd465_mean, y=cpp, fill=country)) + geom_point(size=4, stroke=1, shape=21) + scale_fill_manual(values=c("black","gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black", "gray"), name="Country") +   theme(legend.position=c(.25,.25), text=element_text(size=22),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlab("Mean GDD deadline 2008-2017 \n(ordinal day)") + ylab("CPP, 2018 (h)") + ylim(12.3, 14.2)

gdd_cpp_interval_08<-ggplot(filter(interval_summaries, interval=="y1998-2007", CPP_YEAR==2008), aes(x=gdd465_mean, y=cpp, fill=country)) + geom_point(size=4, stroke=1, shape=22) + scale_fill_manual(values=c("black","gray"), name="Country") + stat_smooth(method=lm, se=FALSE, aes(color=country), linetype="longdash") + scale_color_manual(values=c("black", "gray"), name="Country") +   theme(legend.position=c(.25,.25), text=element_text(size=22),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlab("Mean GDD deadline 1998-2007 \n(ordinal day)") + ylab("CPP, 2008 (h)") + ylim(12.3, 14.2)

```

```{r echo=FALSE, include=FALSE}
#FIG 4
gdd_cpp_plot<-plot_grid(gdd_08, gdd_18, gdd_cpp_interval_08, gdd_cpp_interval_18, nrow=2, labels="AUTO")

```

```{r echo=FALSE, warning=FALSE}
#FIG S5A
#black=Japan, gray=US / circle=2018, square=2008 / solid line=2018, dashed line=2008

gdd_by_interval<-ggplot(filter(interval_summaries, CPP_YEAR==2018), aes(x=latitude, y=gdd465_mean, fill=country, shape=interval)) + 
  geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(22,21), name="Period") + geom_errorbar(aes(ymin=gdd465_mean-gdd465_sd, ymax=gdd465_mean+gdd465_sd, color=country)) +
  stat_smooth(method=lm, se=FALSE, aes(linetype=interval, color=country)) + scale_linetype_manual(values=c("longdash","solid"), name="Period") + 
  scale_color_manual(values=c("black","gray"), name="Country") +
  theme(legend.position=c(.8,.8), text=element_text(size=16),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlab("Latitude") + ylab("Mean GDD Deadline")

```

```{r echo=FALSE, include=FALSE}
data_18<-gdd465_all_new %>% filter(CPP_YEAR==2018) %>% filter(year==2017)
model1<-lm(Estimate~gdd465*country, data=data_18)

data_08<-gdd465_all_new %>% filter(CPP_YEAR==2008) %>% filter(year==2007)
model2<-lm(Estimate~gdd465*country, data=data_08)

model3<-lm(cpp~gdd465_mean*country, data=filter(interval_summaries, interval=="y2008-2017", CPP_YEAR==2018))
model4<-lm(cpp~gdd465_mean*country, data=filter(interval_summaries, interval=="y1998-2007", CPP_YEAR==2008))

anova(model1)
anova(model2)
anova(model3)
anova(model4)

```

```{r echo=FALSE, include=FALSE}
#2015 Diapause Incidence Field Trial
field_data<-read.csv("DI_2015_FIELD.csv")
field_data$WeekDate <- reorder(field_data$WeekDate, field_data$Week)

#field eggs only
field.model.fit <- drm(DI ~ DOY, data = field_data, fct = LL.5())
summary(field.model.fit)
# estimate the 50% intercept
cpp_field<-data.frame(ED(field.model.fit, 50))

field_data_plot<-ggplot(field_data, aes(x=DOY, y=DI)) + geom_point(size=3, shape=21, fill="gray") + 
  theme(legend.position=c(.8,.25), text=element_text(size=24),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
xlab("Ordinal day (2015)") + ylab("Diapause incidence") + geom_vline(aes(xintercept=cpp_field$Estimate), linetype="dotted") + coord_cartesian(xlim=c(193,365))

#2014 GDD diagram
man2014<-read.csv("2014_Mean_Temp_MAN.csv") %>% gather(degreetype, degrees, "CountDegrees":"UncountDegrees")

man2014_gdd_plot<-ggplot(filter(man2014, DOY_2014>193), aes(x=DOY_2014, y=degrees, fill=degreetype)) +
    theme(text=element_text(size=24), legend.position="none",
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  geom_bar(stat="identity", width=1) + 
  scale_fill_manual(values=c("dark gray","light gray")) + 
  geom_hline(yintercept=10, linetype="dashed") + geom_vline(xintercept=251) +
  xlab("Ordinal Day (2014)") + ylab("Mean Daily Temp (C)")

#Fig3
field_and_gdd2014<-plot_grid(field_data_plot, man2014_gdd_plot, nrow=2, align="hv", labels="AUTO")

```

```{r}
#climate variation data
tmean_us<-read.csv("Monthly_Tmean_US.csv", check.names=FALSE) %>% gather(year, meantemp, "1998":"2018", factor_key=FALSE)
#In order to look at winter temperatures, need to combine Nov/Dec with Jan/Feb/March of the following year. January and February are reassigned in this table to the previous year to create seasonal years
tmean_us$year<-as.numeric(tmean_us$year) 
tmean_us$seasonalyear<-if_else(tmean_us$month %in% c("Jan","Feb","Mar"), tmean_us$year-1, tmean_us$year)

#make a list of the coldest months by seasonal year and location
coldest_months_US <- tmean_us %>% group_by(pop, seasonalyear) %>% slice(which.min(meantemp)) %>% filter(!(seasonalyear %in% c(1997,2018))) %>% select(pop, seasonalyear, month)

tmin_us<-read.csv("AvgTmin_US.csv", check.names=FALSE) %>% gather(year, mintemp, "1998":"2018", factor_key=FALSE)
tmin_us$year<-as.numeric(tmin_us$year) 
tmin_us$seasonalyear<-if_else(tmin_us$month %in% c("Jan","Feb","Mar"), tmin_us$year-1, tmin_us$year) 
tmin_us<-filter(tmin_us, !(seasonalyear %in% c(1997,2018)))

tmax_us<-read.csv("AvgTmax_US.csv", check.names=FALSE) %>% gather(year, maxtemp, "1998":"2018", factor_key=FALSE)
tmax_us$year<-as.numeric(tmax_us$year) 
tmax_us$seasonalyear<-if_else(tmax_us$month %in% c("Jan","Feb","Mar"), tmax_us$year-1, tmax_us$year) 
tmax_us<-filter(tmax_us, !(seasonalyear %in% c(1997,2018)))

#make a table of the min, mean, and max temperature of the coldest month of each seasonal year in each location
all_temps_us<-left_join(tmin_us, tmean_us)
all_temps_us<-left_join(all_temps_us, tmax_us)
all_temps_us<-select(all_temps_us, pop, month, seasonalyear, mintemp, meantemp, maxtemp)
cold_month_temps_us<-inner_join(all_temps_us, coldest_months_US)
cold_month_temps_us$country=rep("US",nrow(cold_month_temps_us))

#same process for Japan
tmean_jp<-read.csv("Monthly_Tmean_JP.csv", check.names=FALSE) %>% gather(year, meantemp, "1998":"2018", factor_key=FALSE)
#In order to look at winter temperatures, need to combine Nov/Dec with Jan/Feb/March of the following year. January and February are reassigned in this table to the previous year to create seasonal years
tmean_jp$year<-as.numeric(tmean_jp$year) 
tmean_jp$seasonalyear<-if_else(tmean_jp$month %in% c("Jan","Feb","Mar"), tmean_jp$year-1, tmean_jp$year)

#make a list of the coldest months by seasonal year and location
coldest_months_jp <- tmean_jp %>% group_by(pop, seasonalyear) %>% slice(which.min(meantemp)) %>% filter(!(seasonalyear %in% c(1997,2018))) %>% select(pop, seasonalyear, month)

tmin_jp<-read.csv("AvgTmin_JP.csv", check.names=FALSE) %>% gather(year, mintemp, "1998":"2018", factor_key=FALSE)
tmin_jp$year<-as.numeric(tmin_jp$year) 
tmin_jp$seasonalyear<-if_else(tmin_jp$month %in% c("Jan","Feb","Mar"), tmin_jp$year-1, tmin_jp$year) 
tmin_jp<-filter(tmin_jp, !(seasonalyear %in% c(1997,2018)))

tmax_jp<-read.csv("AvgTmax_JP.csv", check.names=FALSE) %>% gather(year, maxtemp, "1998":"2018", factor_key=FALSE)
tmax_jp$year<-as.numeric(tmax_jp$year) 
tmax_jp$seasonalyear<-if_else(tmax_jp$month %in% c("Jan","Feb","Mar"), tmax_jp$year-1, tmax_jp$year) 
tmax_jp<-filter(tmax_jp, !(seasonalyear %in% c(1997,2018)))

#make a table of the min, mean, and max temperature of the coldest month of each seasonal year in each location
all_temps_jp<-left_join(tmin_jp, tmean_jp)
all_temps_jp<-left_join(all_temps_jp, tmax_jp)
all_temps_jp<-select(all_temps_jp, pop, month, seasonalyear, mintemp, meantemp, maxtemp)
cold_month_temps_jp<-inner_join(all_temps_jp, coldest_months_jp)
cold_month_temps_jp$country=rep("JP",nrow(cold_month_temps_jp))

#add latitude information and combine US/Japan
locations<-read.csv("mapping_pops_cpp_expt.csv")
cold_month_temps_us<-left_join(cold_month_temps_us, dplyr::select(locations, pop, latitude, only_2008), by=c("pop"))
cold_month_temps_jp<-left_join(cold_month_temps_jp, dplyr::select(locations, pop, latitude, only_2008), by=c("pop"))
cold_months_all<-rbind(cold_month_temps_us, cold_month_temps_jp)

cold_months_all<-mutate(cold_months_all, interval=case_when(seasonalyear<=2017 & seasonalyear>=2008 ~ 'y2008-2017', seasonalyear<2008 ~ 'y1998-2007'))

#calculate annual range and annual standard deviation
#start with monthly means
us_var <- tmean_us %>% filter(!(seasonalyear %in% c(1997,2018))) %>% group_by(pop, seasonalyear) %>%
  dplyr::summarise(tempannualrange=max(meantemp)-min(meantemp), tempseasonality=sd(meantemp))
jp_var <- tmean_jp %>% filter(!(seasonalyear %in% c(1997,2018))) %>% group_by(pop, seasonalyear) %>%
  dplyr::summarise(tempannualrange=max(meantemp)-min(meantemp), tempseasonality=sd(meantemp))
all_var<-rbind(us_var, jp_var)
all_var<-left_join(all_var, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))

all_var<-mutate(all_var, interval=case_when(seasonalyear<=2017 & seasonalyear>=2008 ~ 'y2008-2017', seasonalyear<2008 ~ 'y1998-2007'))

tf_us <- read.csv("TotalFrostDays_US.csv", check.names=FALSE) %>% gather(year, TotalFrostDays, "1998":"2018", factor_key=FALSE)
tf_us$year<-as.numeric(tf_us$year)

tf_jp <- read.csv("TotalFrostDays_JP.csv", check.names=FALSE) %>% gather(year, TotalFrostDays, "1998":"2018", factor_key=FALSE)
tf_jp$year<-as.numeric(tf_jp$year)
#adding latitudes
locations<-read.csv("mapping_pops_cpp_expt.csv")
tf_us<-left_join(tf_us, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
tf_jp<-left_join(tf_jp, dplyr::select(locations, pop, latitude, country, only_2008), by=c("pop"))
tf_all<-rbind(tf_us, tf_jp)
tf_all$pop<-as.factor(tf_all$pop)
tf_all<-mutate(filter(tf_all, year!=2018), interval=case_when(year<=2017 & year>=2008 ~ 'y2008-2017',
                                                  year<2008 ~ 'y1998-2007'))

frost_all<-select(left_join(ff_all, tf_all), pop, latitude, country, year, DFF, TotalFrostDays, only_2008, interval)
names(frost_all)[names(frost_all) == 'year']<-'seasonalyear'

clim_vars_all<-left_join(cold_months_all, all_var)
names(clim_vars_all)[names(clim_vars_all) == 'month'] <- 'coldest_month'
names(clim_vars_all)[names(clim_vars_all) == 'mintemp'] <- 'mintemp_coldest_month'
names(clim_vars_all)[names(clim_vars_all) == 'meantemp'] <- 'meantemp_coldest_month'
names(clim_vars_all)[names(clim_vars_all) == 'maxtemp'] <- 'maxtemp_coldest_month'
clim_vars_all<-select(clim_vars_all, pop, latitude, country, seasonalyear, coldest_month, mintemp_coldest_month, meantemp_coldest_month, maxtemp_coldest_month, tempannualrange, tempseasonality, interval)

clim_vars_all<-left_join(clim_vars_all, frost_all)

cpp_year<-dplyr::select(read.csv("cpp_calcs_filtered.csv"), pop, Estimate, year) %>% filter(year!="1988") %>% dplyr::rename("CPP_YEAR"="year")

clim_vars_all<-left_join(clim_vars_all, cpp_year, by=c("pop"))
clim_vars_all$CPP_YEAR<-as.factor(clim_vars_all$CPP_YEAR)

```

```{r}
cp<-dplyr::select(filter(clim_vars_all, CPP_YEAR==2018), mintemp_coldest_month, meantemp_coldest_month, maxtemp_coldest_month, tempannualrange, tempseasonality, DFF, TotalFrostDays)
names(cp)<-c("ColdestMonthMin","ColdestMonthMean","ColdestMonthMax","TempAnnualRange","TempSeasonality","DateFirstFrost","TotalFrostDays")
cv<-dplyr::select(filter(clim_vars_all, CPP_YEAR==2018), pop, latitude, country, seasonalyear, interval)

```

```{r}
clim.pca <- prcomp(cp,
                 center = TRUE,
                 scale. = TRUE)
pca_by_country<-ggbiplot(clim.pca, groups=cv$country) + 
        theme(text=element_text(size=16), legend.position=c(.8,.8),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlim(values=c(-3,3)) + ylim(values=c(-3,3)) +
  scale_color_manual(name="Country", values=c("black","gray"))

pca_by_interval<-ggbiplot(clim.pca, groups=cv$interval) + 
        theme(text=element_text(size=16), legend.position=c(.8,.8),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlim(values=c(-3,3)) + ylim(values=c(-3,3)) +
  scale_color_manual(name="Interval", values=c("black","gray"))

pca_by_year<-ggbiplot(clim.pca, groups=cv$seasonalyear) + 
        theme(text=element_text(size=18), legend.position="bottom",
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlim(values=c(-3,3)) + ylim(values=c(-3,3)) +
  scale_color_gradient(name="Year")

pca_by_latitude<-ggbiplot(clim.pca, groups=cv$latitude) +         
  theme(text=element_text(size=18), legend.position="bottom",
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  xlim(values=c(-3,3)) + ylim(values=c(-3,3)) +
  scale_color_gradient(name="Latitude")

pcas_for_fig<-plot_grid(pca_by_country, pca_by_interval, nrow=1)
```

```{r}
#FIG S5
plot_grid(gdd_by_interval, pcas_for_fig, nrow=2, labels="AUTO")
```

### 2020-04-06 maps.R

@@ -1,37 +1,0 @@
-install.packages(c("maps", "mapdata"))
-devtools::install_github("dkahle/ggmap")
-install.packages("tidyverse")
-install.packages("scatterpie")
-install.packages("ggedit")
-
-library(ggplot2)
-library(ggmap)
-library(maps)
-library(mapdata)
-library(scatterpie)
-library(ggedit)
-
-usa<-map_data("usa")
-jp<-map_data("japan")
-pops<-data.frame(read.csv("mapping_pops_cpp_expt.csv"))
-
-us_map<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="black") + 
-  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 1) + xlim(c(-90, -65)) +
-  geom_label(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
-  theme_bw() 
-
-japan_map<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="black") + 
-  coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 1)  +
-  geom_label(data=filter(pops, country=="JP"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
-  theme_bw() + theme(axis.title.x=element_blank(), axis.title.y=element_blank())
-
-DI_data<-data.frame(read.csv("diapause_incidence_8h_2008_2018.csv"))
-DI_data$ND<-1-DI_data$DI
-DI_info<-full_join(DI_data, pops)
-DI_info<-rename(DI_info, Diapause=DI, Nondiapause=ND)
-
-remove_geom(us_map, "label") + geom_scatterpie(data=filter(DI_info, country=="US"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + 
-  facet_grid(~year) + theme(axis.title.x=element_blank(), axis.title.y=element_blank())
-
-remove_geom(japan_map, "label") + geom_scatterpie(data=filter(DI_info, country=="JP"), aes(x=longitude, y=latitude, r=0.5), cols=c("Diapause", "Nondiapause")) + 
[…]

### 2020-04-13 Albopictus_cpp_MS.Rmd

--- 2020-04-06 _albopictus_cpp_MS.Rmd
@@
 #Spearman rank correlation coefficients
-cor.test(filter(DI_short_2, country=="US", year==2018)$lat, filter(DI_short, country=="US", year==2018)$DI, alternative="two.sided", method="spearman")
-cor.test(filter(DI_short_2, country=="JP", year==2018)$lat, filter(DI_short, country=="JP", year==2018)$DI, alternative="two.sided", method="spearman")
+cor.test(filter(DI_short, country=="US", year==2018)$lat, filter(DI_short, country=="US", year==2018)$DI, alternative="two.sided", method="spearman")
+cor.test(filter(DI_short, country=="JP", year==2018)$lat, filter(DI_short, country=="JP", year==2018)$DI, alternative="two.sided", method="spearman")
 
@@
 #bw locations map
-us_map_bw<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="black") + 
-  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 2) + xlim(c(-87, -65)) + ylim(c(24,48)) +
-  geom_label_repel(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines"), nudge_x=1.5, size=4) + theme_bw() + theme(text=element_text(size=24), plot.margin = unit(c(0,0,0,0), "cm")) + xlab("") + ylab("")
+us_map_bw<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="dark gray") + 
+  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 2, shape=21, color="black", fill="gray", stroke=1) + xlim(c(-87, -65)) + ylim(c(24,48)) +
+  geom_text_repel(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines"), nudge_x=1.5, size=4) + theme_bw() + theme(text=element_text(size=24), plot.margin = unit(c(0,0,0,0), "cm"), panel.grid=element_line(size=0.1)) + xlab("") + ylab("")
 
-japan_map_bw<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="black") + 
+japan_map_bw<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="dark gray") +
   coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 2)  + xlim(c(125, 147)) + ylim(c(24,48)) +
-  geom_label_repel(data=filter(pops, country=="JP", only_2008=="n"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines"), size=4) + theme_bw() + theme(text=element_text(size=24), plot.margin = unit(c(0,0,0,0), "cm")) + xlab("") + ylab("Latitude")
+  geom_text_repel(data=filter(pops, country=="JP", only_2008=="n"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines"), size=4) + theme_bw() + theme(text=element_text(size=24), plot.margin = unit(c(0,0,0,0), "cm"), panel.grid=element_line(size=0.1)) + xlab("") + ylab("Latitude")
 

### 2020-04-30 Albopictus_cpp_MS.Rmd
Figure update

@@
 usa<-map_data("usa")
-jp<-map_data("japan")
+jp<-map_data("world") %>% filter(region=="Japan")
 pops<-data.frame(read.csv("mapping_pops_cpp_expt.csv"))
@@
   coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 2, shape=21, color="black", fill="gray", stroke=1) + xlim(c(-87, -65)) + ylim(c(24,48)) +
-  geom_text_repel(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines"), nudge_x=1.5, size=4) + theme_bw() + theme(text=element_text(size=24), plot.margin = unit(c(0,0,0,0), "cm"), panel.grid=element_line(size=0.1)) + xlab("") + ylab("")
+  geom_text_repel(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines"), nudge_x=1.5, size=6) + theme_bw() + theme(text=element_text(size=24), plot.margin = unit(c(0,0,0,0), "cm"), panel.grid=element_line(size=0.1)) + xlab("") + ylab("")
 
@@
   coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 2)  + xlim(c(125, 147)) + ylim(c(24,48)) +
-  geom_text_repel(data=filter(pops, country=="JP", only_2008=="n"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines"), size=4) + theme_bw() + theme(text=element_text(size=24), plot.margin = unit(c(0,0,0,0), "cm"), panel.grid=element_line(size=0.1)) + xlab("") + ylab("Latitude")
-
-map_panels_bw<-plot_grid(japan_map_bw, us_map_bw)
+  geom_text_repel(data=filter(pops, country=="JP", only_2008=="n"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines"), size=6) + theme_bw() + theme(text=element_text(size=24), plot.margin = unit(c(0,0,0,0), "cm"), panel.grid=element_line(size=0.1)) + xlab("") + ylab("Latitude") +
+  labs(x = NULL)
+map_panels_bw<-plot_grid(japan_map_bw, us_map_bw, align="hv")
 ```
@@
 #FIG S4
-gdd_deadline_plots<-ggplot(filter(gdd465_all, only_2008=="n", year!=2018), aes(x=latitude, y=gdd465, fill=country)) + scale_fill_manual(values=c("black", "gray"), name="Country") + geom_point(shape=21, size=3) + facet_wrap(~year) + xlab("Latitude") + ylab("GDD Deadline (ordinal day)") + theme_bw() +
+gdd_deadline_plots<-ggplot(filter(gdd465_all, only_2008=="n", year!=2018), aes(x=latitude, y=gdd465, fill=country)) + scale_fill_manual(values=c("black", "gray"), name="Country") + geom_point(shape=21, size=2) + facet_wrap(~year) + xlab("Latitude") + ylab("GDD Deadline (ordinal day)") + theme_bw() +
 theme(legend.position="bottom", text=element_text(size=18)) 
@@
         panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
-    xlab("GDD Deadline 2007 (ordinal day)") + ylab("CPP (h), 2008") + ylim(12.3,14.2)
+    xlab("GDD Deadline 2007 \n(ordinal day)") + ylab("CPP (h), 2008") + ylim(12.3,14.2)
+
+gdd_08_sameyear <- ggplot(filter(gdd465_all_new, year==2008, CPP_YEAR==2008), aes(x=gdd465, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=4, stroke=1, shape=22) + stat_smooth(method=lm, se=FALSE, aes(color=country), linetype="longdash") + scale_color_manual(values=c("black","gray"), name="Country") +
+  theme(legend.position=c(.25,.25), text=element_text(size=22),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
+    xlab("GDD Deadline 2008 \n(ordinal day)") + ylab("CPP (h), 2008") + ylim(12.3,14.2)
 
@@
         panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
-xlab("GDD Deadline 2017 (ordinal day)") + ylab("CPP (h), 2018") + ylim(12.3, 14.2)
+xlab("GDD Deadline 2017 \n(ordinal day)") + ylab("CPP (h), 2018") + ylim(12.3, 14.2)
+
+gdd_18_sameyear <- ggplot(filter(gdd465_all_new, year==2018, CPP_YEAR==2018), aes(x=gdd465, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=4, stroke=1, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
+  theme(legend.position=c(.25,.25), text=element_text(size=22),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
+xlab("GDD Deadline 2018 (ordinal day)") + ylab("CPP (h), 2018") + ylim(12.3, 14.2)
 
@@
 #FIG 4
-gdd_cpp_plot<-plot_grid(gdd_08, gdd_18, gdd_cpp_interval_08, gdd_cpp_interval_18, nrow=2, labels="AUTO")
-
+gdd_cpp_plot_decade<-plot_grid(gdd_cpp_interval_08, gdd_cpp_interval_18, nrow=2, labels="AUTO")
+#FIG S5
+gdd_cpp_plot_year<-plot_grid(gdd_08, gdd_18, nrow=2, labels="AUTO")
 ```
@@
 ```{r echo=FALSE, warning=FALSE}
-#FIG S5A
+#FIG S6A
 #black=Japan, gray=US / circle=2018, square=2008 / solid line=2018, dashed line=2008
@@
 ```{r}
-#FIG S5
+#FIG S6
 plot_grid(gdd_by_interval, pcas_for_fig, nrow=2, labels="AUTO")

### 2020-05-27 Albopictus_cpp_MS.Rmd
color correction

@@
 #black=Japan, gray=US / circle=2018, square=2008 / solid line=2018, dashed line=2008
-country_DI_bw<-ggplot(DI_short, aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("gray","black"), name="Country") + scale_shape_manual(values=c(22,21), name="Year") + geom_point(size=3) +   theme(legend.position=c(.8,.25), text=element_text(size=18),
+country_DI_bw<-ggplot(DI_short, aes(x=lat, y=DI, fill=country, shape=year)) + scale_fill_manual(values=c("black","gray"), name="Country") + scale_shape_manual(values=c(22,21), name="Year") + geom_point(size=3) +   theme(legend.position=c(.8,.25), text=element_text(size=18),
         panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
         panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
-xlab("Latitude") + ylab("Short-day diapause incidence") + stat_smooth(aes(linetype=year, color=country), se=FALSE) + scale_linetype_manual(values=c("longdash","solid"), name="Year") + scale_color_manual(values=c("gray","black"), name="Country")
+xlab("Latitude") + ylab("Short-day diapause incidence") + stat_smooth(aes(linetype=year, color=country), se=FALSE) + scale_linetype_manual(values=c("longdash","solid"), name="Year") + scale_color_manual(values=c("black","gray"), name="Country")
 

### 2020-07-27 Albopictus_cpp_MS.Rmd
more data

@@
 library("papeR")
-library("flextable")
 library("drc") #3.0.1
@@
 gdd400_all$pop<-as.factor(gdd400_all$pop)
-gdd400_all<-mutate(gdd400_all, interval=case_when(year<=2017 & year>=2008 ~ 'y2008-2017',
-                                                  year<2008 ~ 'y1998-2007'))
+gdd400_all<-mutate(gdd400_all, interval=case_when(year<=2017 & year>=2008 ~ 'y2008-2017',year<2008 ~ 'y1998-2007'))
 
@@
 
+```{r}
+#photoperiods on deadline dates
+PP_us <- read.csv("deadline465_pp_US.csv", check.names=FALSE) %>% gather(year, deadline_daylength, "1998":"2018", factor_key=FALSE)
+PP_jp <- read.csv("deadline465_pp_jp.csv", check.names=FALSE) %>% gather(year, deadline_daylength, "1998":"2018", factor_key=FALSE)
+PP_pops<-rbind(PP_us, PP_jp)
+PP_pops$year<-as.numeric(PP_pops$year)
+```
+
 ```{r, echo=FALSE, include=FALSE}
@@
 
+```
+
+```{r}
+gdd_all_new<-left_join(gdd_all_new, PP_pops, by=c("pop","year"))
 ```
@@
 
+```{r}
+#CPP vs. photoperiod on deadline day
+gdd_pp_08 <- ggplot(filter(gdd_all_new, year==2007, CPP_YEAR==2008), aes(x=deadline_daylength, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=4, stroke=1, shape=22) + stat_smooth(method=lm, se=FALSE, aes(color=country), linetype="longdash") + scale_color_manual(values=c("black","gray"), name="Country") +
+  theme(legend.position=c(.75,.25), text=element_text(size=22),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
+    xlab("2007 Deadline Day Length") + ylab("CPP (h), 2008") + ylim(12.3,14.2)
+
+gdd_pp_18 <- ggplot(filter(gdd_all_new, year==2017, CPP_YEAR==2018), aes(x=deadline_daylength, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=4, stroke=1, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
+  theme(legend.position=c(.75,.25), text=element_text(size=22),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
+xlab("2017 Deadline Day Length") + ylab("CPP (h), 2018") + ylim(12.3, 14.2)
+
+```
+
 ```{r echo=FALSE, include=FALSE}
@@
 #FIG S5
-gdd_cpp_plot_year<-plot_grid(gdd_08, gdd_18, nrow=2, labels="AUTO")
+gdd_cpp_plot_year<-plot_grid(gdd_08, gdd_pp_08, gdd_18, gdd_pp_18, nrow=2, labels="AUTO")
 ```
@@
 field_and_gdd2014<-plot_grid(field_data_plot, man2014_gdd_plot, nrow=2, align="hv", labels="AUTO")
+
+```
+
+```{r}
+sample_locations_2017<-read.csv("MeanTempUSAandJapan2017.csv") %>% gather(location, temp, "MAN":"TAK")
+sample_locations_2017$UncountDegrees<-if_else(sample_locations_2017$temp<=10,sample_locations_2017$temp,10)
+sample_locations_2017$CountDegrees<-if_else(sample_locations_2017$temp<=10,0,sample_locations_2017$temp-10)
+
+sample_locations_2017_reshape <- dplyr::select(sample_locations_2017, DOY, location, CountDegrees, UncountDegrees) %>% gather(degreetype, degrees, "CountDegrees":"UncountDegrees")
+
+sample_locations_2017_reshape$location<-factor(sample_locations_2017_reshape$location, levels=c("BRU","FAY","MAN","KAG","TAK","SEN"))
+
+sample_locations_2017_reshape$deadline2017<-case_when(
+  sample_locations_2017_reshape$location == "BRU" ~ 289,
+  sample_locations_2017_reshape$location == "FAY" ~ 276,
+  sample_locations_2017_reshape$location == "MAN" ~ 263,
+  sample_locations_2017_reshape$location == "KAG" ~ 274,
+  sample_locations_2017_reshape$location == "TAK" ~ 257,
+  sample_locations_2017_reshape$location == "SEN" ~ 246,)
+
+location_labels<-c("BRU (31.3 N)","FAY (35.03 N)","MAN (38.63 N)","KAG (31.6 N)","TAK (35.35 N)","SEN (38.29 N)")
+names(location_labels)<-c("BRU","FAY","MAN","KAG","TAK","SEN")
+
+sample_locations_gdd<-ggplot(filter(sample_locations_2017_reshape, DOY>180), aes(x=DOY, y=degrees, fill=degreetype)) +
+    theme(text=element_text(size=24), legend.position="none",
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
+  geom_bar(stat="identity", width=1) + 
+  scale_fill_manual(values=c("dark gray","light gray")) + 
+  geom_hline(yintercept=10, linetype="dashed") + 
+  xlab("Ordinal Day (2017)") + ylab("Mean Daily Temp (C)") + geom_vline(aes(xintercept=deadline2017)) +
+  facet_wrap(~location, labeller=labeller(location=location_labels))
+
+```
+
+```{r}
+#Fig S4 two parts
+gdd_examples<-plot_grid(gdd_deadline_plots, sample_locations_gdd, nrow=2, labels="AUTO")
 

### 2020-07-27 Albopictus_cpp_MS.Rmd
NEE submission

@@
 
-```{r}
-#photoperiods on deadline dates
-PP_us <- read.csv("deadline465_pp_US.csv", check.names=FALSE) %>% gather(year, deadline_daylength, "1998":"2018", factor_key=FALSE)
-PP_jp <- read.csv("deadline465_pp_jp.csv", check.names=FALSE) %>% gather(year, deadline_daylength, "1998":"2018", factor_key=FALSE)
-PP_pops<-rbind(PP_us, PP_jp)
-PP_pops$year<-as.numeric(PP_pops$year)
-```
-
 ```{r, echo=FALSE, include=FALSE}
@@
 gdd_all_new$CPP_YEAR<-as.factor(gdd_all_new$CPP_YEAR)
-
-```
-
-```{r}
-gdd_all_new<-left_join(gdd_all_new, PP_pops, by=c("pop","year"))
 ```
@@
 
-```{r}
-#CPP vs. photoperiod on deadline day
-gdd_pp_08 <- ggplot(filter(gdd_all_new, year==2007, CPP_YEAR==2008), aes(x=deadline_daylength, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=4, stroke=1, shape=22) + stat_smooth(method=lm, se=FALSE, aes(color=country), linetype="longdash") + scale_color_manual(values=c("black","gray"), name="Country") +
-  theme(legend.position=c(.75,.25), text=element_text(size=22),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
-    xlab("2007 Deadline Day Length") + ylab("CPP (h), 2008") + ylim(12.3,14.2)
-
-gdd_pp_18 <- ggplot(filter(gdd_all_new, year==2017, CPP_YEAR==2018), aes(x=deadline_daylength, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=4, stroke=1, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
-  theme(legend.position=c(.75,.25), text=element_text(size=22),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
-xlab("2017 Deadline Day Length") + ylab("CPP (h), 2018") + ylim(12.3, 14.2)
-
-```
-
 ```{r echo=FALSE, include=FALSE}
@@
 #FIG S5
-gdd_cpp_plot_year<-plot_grid(gdd_08, gdd_pp_08, gdd_18, gdd_pp_18, nrow=2, labels="AUTO")
+gdd_cpp_plot_year<-plot_grid(gdd_08, gdd_18, nrow=2, labels="AUTO")
 ```

### 2020-11-11 Albopictus_cpp_MS.Rmd
added alt deadline data

@@
 devtools::install_github("dkahle/ggmap")
-install.packages("scatterpie")
 install.packages("ggedit")
@@
 library("papeR")
+library("knitr") 
 library("drc") #3.0.1
-library("knitr") #1.25
-library("tidyverse") #1.2.1
+library("tidyverse") #1.3.0
 library("stringr") #1.4.0
-library("geosphere") #1.5.10
 library("lubridate") #1.7.4
-library("maps")
-library("mapdata")
-library("ggedit")
+library("maps") #3.3.0
+library("mapdata") #2.3.0
+library("cowplot") #1.0.0
+library("stats") #3.6.2
+library("ggbiplot") #0.55
 library("ggrepel")
-library("cowplot")
-library("ggrepel")
-library("stats")
-library("ggbiplot")
-
 ```
@@
 cpp_2008<-right_join(cpp_2008, popdata_2008, by="pop")
-#1988 data not used in present study but can load it anyway
-cpp_1988<-data.frame(read.csv("CPP_1988.csv")) 
-cpp_1988$pop<-na_if(cpp_1988$pop, "NA")
-cpp_1988$StdError<-na_if(cpp_1988$StdError, "NA")
 
@@
 cppmodel_result<-data.frame(anova(cppmodel))
-```
-
-```{r, echo=FALSE, include=FALSE}
-#basic breakdown
-ggplot(cpp_all, aes(x=lat, y=Estimate, color=country)) + 
-  geom_point() + 
-  geom_errorbar(aes(ymin=Estimate-StdError, ymax=Estimate+StdError), na.rm=TRUE) + 
-  facet_grid(~year) + stat_smooth(method=lm, se=FALSE)
 ```
@@
 pops<-data.frame(read.csv("mapping_pops_cpp_expt.csv"))
-us_map<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="blue") + 
-  coord_fixed(1.3) + geom_point(data = filter(pops, country=="US"), aes(x = longitude, y = latitude), size = 1) + xlim(c(-85, -65)) +
-  geom_label(data=filter(pops, country=="US"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
-  theme_bw() 
-japan_map<-ggplot() + geom_polygon(data = jp, aes(x=long, y = lat, group=group), fill=NA, color="red") + 
-  coord_fixed(1.3) + geom_point(data = filter(pops, country=="JP", only_2008=="n"), aes(x = longitude, y = latitude), size = 1)  + xlim(c(125, 147)) +
-  geom_label_repel(data=filter(pops, country=="JP", only_2008=="n"), aes(x=longitude, y=latitude, label=pop), hjust=0, label.padding=unit(0.1, "lines")) +
-  theme_bw() + theme(axis.title.x=element_blank(), axis.title.y=element_blank())
 
-#bw locations map
 us_map_bw<-ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group=group), fill=NA, color="dark gray") + 
@@
 
-gdd_08_sameyear <- ggplot(filter(gdd465_all_new, year==2008, CPP_YEAR==2008), aes(x=gdd465, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=4, stroke=1, shape=22) + stat_smooth(method=lm, se=FALSE, aes(color=country), linetype="longdash") + scale_color_manual(values=c("black","gray"), name="Country") +
-  theme(legend.position=c(.25,.25), text=element_text(size=22),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
-    xlab("GDD Deadline 2008 \n(ordinal day)") + ylab("CPP (h), 2008") + ylim(12.3,14.2)
-
 gdd_18 <- ggplot(filter(gdd465_all_new, year==2017, CPP_YEAR==2018), aes(x=gdd465, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=4, stroke=1, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
@@
 xlab("GDD Deadline 2017 \n(ordinal day)") + ylab("CPP (h), 2018") + ylim(12.3, 14.2)
-
-gdd_18_sameyear <- ggplot(filter(gdd465_all_new, year==2018, CPP_YEAR==2018), aes(x=gdd465, y=Estimate, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=4, stroke=1, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
-  theme(legend.position=c(.25,.25), text=element_text(size=22),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
-xlab("GDD Deadline 2018 (ordinal day)") + ylab("CPP (h), 2018") + ylim(12.3, 14.2)
 
@@
 
-```{r echo=FALSE, warning=FALSE}
-#FIG S6A
-#black=Japan, gray=US / circle=2018, square=2008 / solid line=2018, dashed line=2008
-
-gdd_by_interval<-ggplot(filter(interval_summaries, CPP_YEAR==2018), aes(x=latitude, y=gdd465_mean, fill=country, shape=interval)) + 
-  geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(22,21), name="Period") + geom_errorbar(aes(ymin=gdd465_mean-gdd465_sd, ymax=gdd465_mean+gdd465_sd, color=country)) +
-  stat_smooth(method=lm, se=FALSE, aes(linetype=interval, color=country)) + scale_linetype_manual(values=c("longdash","solid"), name="Period") + 
-  scale_color_manual(values=c("black","gray"), name="Country") +
-  theme(legend.position=c(.8,.8), text=element_text(size=16),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
-  xlab("Latitude") + ylab("Mean GDD Deadline")
-
-```
-
 ```{r echo=FALSE, include=FALSE}
@@
 anova(model4)
+
+```
+
+```{r echo=FALSE, warning=FALSE}
+#FIG S6A
+#black=Japan, gray=US / circle=2018, square=2008 / solid line=2018, dashed line=2008
+
+gdd_by_interval<-ggplot(filter(interval_summaries, CPP_YEAR==2018), aes(x=latitude, y=gdd465_mean, fill=country, shape=interval)) + 
+  geom_point(size=3) + scale_fill_manual(values=c("black", "gray"), name="Country") + scale_shape_manual(values=c(22,21), name="Period") + geom_errorbar(aes(ymin=gdd465_mean-gdd465_sd, ymax=gdd465_mean+gdd465_sd, color=country)) +
+  stat_smooth(method=lm, se=FALSE, aes(linetype=interval, color=country)) + scale_linetype_manual(values=c("longdash","solid"), name="Period") + 
+  scale_color_manual(values=c("black","gray"), name="Country") +
+  theme(legend.position=c(.8,.8), text=element_text(size=16),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
+  xlab("Latitude") + ylab("Mean GDD Deadline")
 
@@
 gdd_examples<-plot_grid(gdd_deadline_plots, sample_locations_gdd, nrow=2, labels="AUTO")
+```
+
+```{r}
+#range of possible deadlines
+alt_deadlines<-read.csv("gdd_alternate_deadlines.csv")
+#France location in 2009
+filter(alt_deadlines, Country=="FRA", Year==2009)
+```
+
+``` {r}
+#range of possible deadlines w/latitude/CPP info
+deadlines_range<-left_join(gdd_all_new, alt_deadlines, by=c("pop"="Location", "year"="Year")) %>% select(pop, year, latitude, country, only_2008, Estimate, CPP_YEAR, gdd465, Deadline_GDD_300, Deadline_GDD_350, Deadline_GDD_400, Deadline_GDD_450, Deadline_GDD_500, Deadline_GDD_550, Deadline_GDD_600)
+
+ggplot(filter(deadlines_range, year==2017, CPP_YEAR==2018), aes(x=latitude, y=Deadline_GDD_600, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=2, stroke=1, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
+  theme(legend.position=c(.25,.25), text=element_text(size=22),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
+xlab("Latitude") + ylab("Deadline")
+
+data_18_range<-deadlines_range %>% filter(CPP_YEAR==2018) %>% filter(year==2017)
+model300<-lm(Estimate~Deadline_GDD_300*country, data=data_18_range)
+model350<-lm(Estimate~Deadline_GDD_350*country, data=data_18_range)
+model400<-lm(Estimate~Deadline_GDD_400*country, data=data_18_range)
+model450<-lm(Estimate~Deadline_GDD_450*country, data=data_18_range)
+model500<-lm(Estimate~Deadline_GDD_500*country, data=data_18_range)
+model550<-lm(Estimate~Deadline_GDD_550*country, data=data_18_range)
+model600<-lm(Estimate~Deadline_GDD_600*country, data=data_18_range)
+
+rsq<-c(summary(model300)$r.squared, summary(model300)$r.squared, summary(model350)$r.squared, summary(model400)$r.squared, summary(model450)$r.squared, summary(model500)$r.squared, summary(model550)$r.squared, summary(model600)$r.squared)
+
+data_18_range_long<-pivot_longer(data_18_range, Deadline_GDD_300:Deadline_GDD_600, names_to="deadlinetype", values_to="deadline")
+
+ggplot(filter(data_18_range_long, year==2017, CPP_YEAR==2018), aes(x=latitude, y=deadline, fill=country)) + scale_fill_manual(values=c("black","gray"), name="Country") + geom_point(size=2, stroke=1, shape=21) + stat_smooth(method=lm, se=FALSE, aes(color=country)) + scale_color_manual(values=c("black","gray"), name="Country") +
+  theme(legend.position="bottom",
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
+xlab("Latitude") + ylab("Deadline") + facet_wrap(~deadlinetype)
+
+```
+
+```{r}
+
+#add range of possible deadline values to Manassas 2014
+man_deadlines_range<-filter(deadlines_range, pop=="MAN", year==2014, CPP_YEAR==2018) %>% select(Deadline_GDD_300:Deadline_GDD_600)  %>% t() %>% data.frame() %>% pull()
+
+man2014_gdd_alt_plot <-
+  ggplot(filter(man2014, DOY_2014 %in% c(238:278)), aes(x=DOY_2014, y=degrees, fill=degreetype)) +
+    theme(text=element_text(size=24), legend.position="none",
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
+  geom_bar(stat="identity", width=1) + 
+  scale_fill_manual(values=c("dark gray","light gray")) + 
+  geom_hline(yintercept=10, linetype="dashed") + geom_vline(xintercept=251) + geom_vline(xintercept=man_deadlines_range, linetype="dotted") + 
+  annotate("text", x=(man_deadlines_range)-0.5, y=c(1,3,5,7,9,11,13), label=c("GDD300", "GDD350", "GDD400","GDD450","GDD500","GDD550","GDD600"),angle=90) +
+  geom_vline(xintercept=252.5, color="red") + annotate("text",x=250.5, y=20, label="GDD-465 (model used)", fontface="bold", angle=90) + annotate("text",x=253, y=20, label="50% diapause in the field", color="red", angle=90) +
+  xlab("Ordinal Day (2014)") + ylab("Mean Daily Temp (C)")
 
